---
title: "COIN_Analyses"
output: html_document
date: "2023-01-29"
---

```{r read in, include=FALSE}
library(tidyverse)
library(ade4)

original_colour_data <- read.csv('./2013-09-30 MHC_Colour.csv')
full_data <- read.csv('./Full DF 2023-01-23.csv')
```


#### To Do:   
-Isolate all 3 datasets   
-Format into matrices   
-Run PCAs  
-Visualization?

*Replicate COIN(1), 1000 bootstrap replicates*

    Data matrix (1) contained MHC-ST frequencies for each population
    Data matrix (2) contained colour means and Gyrodactylus prevalence for each population
*Replicate COIN(2)*

    Data matrix (1) contained MHC-Ar, and MHC-STr
    Data matrix (2) contained colour means and Gyrodactylus prevalence for each population


For coinertia analysis the row weighting must be matched between the two dudi (duality diagrams), so for each pca the rows must indicate the same samples. Using only fish we have both genetic data and patternize colour data for we have N = 193.
```{r data prep}
data_for_matrices <- full_data %>% 
  filter(!is.na(Orange) & !is.na(Black)) %>% 
  filter(Year == 2010 & Sex == 'Male') 
#excluding all 2009 guppies because we had unbalanced representation across sites and years

#Parasite and colour
#Note: not using InfectedBinary because it should have less explanatory power than Tparasites
#Should we include body length here?
colour_matrix <- data_for_matrices %>% 
  select(Tparasites,Orange:Black) # total parasites, relative body area of each colour (patternize)

#MHC ST frequency
mhc_f_matrix <- data_for_matrices %>% 
  select(30:38) %>% # Supertype columns, some have NAs, not all fish have same number of STs
  mutate_if(is.numeric, ~replace_na(., 0)) #replace NA with 0 for pca

# allelic richness and ST richness
mhc_r_matrix <- data_for_matrices %>% 
  select(Ai, STi) #richness columns, number of different alleles and MHC STs
```

Co-inertia analyses
```{r coin}
#COIN 1 - colour and MHC ST frequency
colour_pca <- dudi.pca(colour_matrix, scannf = FALSE, nf = 3)
mhc_f_pca <- dudi.pca(mhc_f_matrix, scannf = FALSE, nf = 3)

coinertia(mhc_f_pca, colour_pca, scannf = TRUE) #manually put in axes number to keep? 2?

#COIN 2 - colour and MHC ST and allelic richness
mhc_r_pca <- dudi.pca(mhc_r_matrix, scannf = FALSE, nf = 3)

coinertia(mhc_r_pca, colour_pca, scannf = TRUE) #manually put in axes number to keep? 2?
```

