---
title: "MHC analysis"
author: "Dr. Kiyoko Gotanda"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r load packages, echo = FALSE, message = FALSE}
library(car)
library(tidyverse)
library(ade4)
library(adegraphics)
library(stringr)
library(corrplot)
```

### Data Prep
Load dataframes
```{r load DFs, echo = TRUE}
rm(list = ls())
# Original Colour data
#ColDF <- read.csv("./2013-09-30 MHC_Colour.csv", header = T, row.names = NULL, stringsAsFactors = T, na.strings = "")
# Individual immunity data
MHCDF <- read.csv("./Full DF 2023-01-23.csv", header = T, row.names = NULL, stringsAsFactors = T, na.strings = "")
# population immunity data
PopDF <- read.csv("./microAr 2023-03-16.csv", header = F, row.names = NULL, na.strings = "")
```


```{r clean DFs, echo = FALSE}
#renaming the allele and st columns for ease of use
colnames(MHCDF) <- c("Fish","River","Site","Pred","Predation","Sex","Year","Tparasites","Infected",	"InfectedBinary", "X", "Orange","Black","site","drainage","sex","length","n.gyros","colour_notes",	"Drainage",	"gt_1","gt_2","gt_3","gt_4","gt_5","gt_6","gt_7","gt_8","gt_9","st_1","st_2","st_3",	"st_4","st_5","st_6","st_7","st_8","st_9","Ai","STi")

#rename and clean Population df from appendix
colnames(PopDF) <- PopDF[5,] #reassign column names
PopDF <- PopDF[-(1:5), 1:21 , drop = FALSE] # remove empty rows and colour cols
PopDF$Site <- str_replace(PopDF$Site, '-', '') # standardise site names
PopDF$Site <- str_replace(PopDF$Site, 'El Cedro', 'ElCedro') 
string_cols <- colnames(PopDF[,2:4])
PopDF[string_cols] <- lapply(PopDF[string_cols] , factor)
number_cols <- colnames(PopDF[,c(1,5:21)])
PopDF[number_cols] <- lapply(PopDF[number_cols] , as.numeric)

#excluding all 2009 guppies because we had unbalanced representation across sites and years
data_analysis <- MHCDF %>% 
    filter(!is.na(Orange) & !is.na(Black) & Year == '2010' & Sex == 'Male')
```

We have colour data for `r data_analysis %>% tally() %>% pull()` photos with corresponding MHC data, all photos were taken in 2010. These span `r n_distinct(data_analysis$Site)` sites. Below are the population specific means for relative area of both orange and black.

```{r pop means, echo=FALSE, message = FALSE}
pop_means <- data_analysis %>% 
  group_by(Site) %>% 
  summarize(mean_o = mean(Orange),
            mean_b = mean(Black)) %>% 
  add_row(data.frame(Site="All", mean_o=mean(data_analysis$Orange), mean_b=mean(data_analysis$Black)))

#adding to PopDF, removing populations with no colour data
PopDF <- inner_join(PopDF, pop_means, by = 'Site')

knitr::kable(pop_means, col.names = c("Site", "Mean Relative Orange", "Mean Relative Black"), "simple")
```


```{r st freq data, echo=FALSE, message = FALSE}
# We found this data in the population level table from Jackie's thesis, but we don't have individual frequencies.

ST_freq <- MHCDF %>% 
  filter(!is.na(Orange) & !is.na(Black)) %>% 
  filter(Year == 2010 & Sex == 'Male') %>% 
  select(-Pred,-X,-site,-sex,-n.gyros) %>% 
  pivot_longer(cols = gt_1:st_9, names_pattern = "(.t)_(.)", names_to = c('.value','position'), values_drop_na = TRUE) %>% 
  group_by(Fish, st) %>% 
  summarize(freq = n()) %>% 
  pivot_wider(names_from = st, values_from = freq, names_sort = TRUE, names_prefix = "ST_freq_") #no ST 10

data_analysis <- inner_join(data_analysis, ST_freq, by = "Fish") %>% 
  select(Fish:Site, Predation:InfectedBinary, Orange:Black, length, colour_notes, Drainage, Ai:ST_freq_11)
```


### *Reviewers Comments*

* They mentioned replicating analyses with both drainages separately so we can confirm trends across drainages   
  + COIA, I don't think we can add drainage but we could do a drainage specific COIA?   
  + Adding a drainage term in richness models?   

* Why use parametric and non-parametric tests?   
  + by switching to linear models we aren't mixing parametric and non-parametric anymore   

* Adding in the data from the different years too   
  + We did this, it allowed us to move to individual level analysis   

* Changing to Individual level analyses to leverage the new data   
  + COIA (DONE)   
  + Wilcoxon - change to models with individual data   
  + Microsatellite - dependent on whether we have individual microsat data   

### Analysis To Do: 

###### Original Analysis   
* COIA (In Progress)   
  + prepare all 3 datasets (DONE)   
  + Run PCAs (DONE)   
  + bootstrapping for COIA to test significance (DONE)   
  + pearson's correlation coefficient test (maybe???) *I am not sure about this*   
  + Visualization (In progress)   
* richness linear models (micro-Ar as independent, Ai and STi as dependent) (DONE)   
* Wilcoxon rank-sum test on all variables by predation level (DONE)   
* Wilcoxon rank-sum test on all variables by drainage (DONE-Caroni & North)   

###### NEW and IMPROVED Analysis   
* predation and drainage tests (In Progress)   
  + move to linear mixed models that test for predation and drainage   
  + use data_analysis df with individual level data   
  + Not all individuals have all ST, drop NAs? How will this impact model?   
* microsatellite linear models   
  + check if microsatellite data has individual level in Jackie's thesis   
  + *maybe* re-do richness linear models with individual level data if available   
* boxplot and other visualization/figure recreation with individual data   


#### Co-inertia analyses

For coinertia analysis the row weighting must be matched between the two dudi (duality diagrams), so for each pca the rows must indicate the same samples. Using only fish we have both genetic data and patternize colour data for we have N = 193.
```{r data prep}
# check for uninfected populations to exclude from analyses between immunity and colour
no_parasite_sites <- data_analysis %>% 
  group_by(Site) %>% 
  summarise(max = max(Tparasites)) %>% 
  filter(max < 1) %>% 
  pull(Site)

data_matrices <-data_analysis %>% 
  filter(!Site %in% no_parasite_sites) %>% 
  filter(Site != "ElCedro1") #removed in original study, parasite prev an order of magnitude below others

### Parasite and colour
#Note: not using InfectedBinary because it should have less explanatory power than Tparasites
#Should we include body length here?
colour_matrix <- data_matrices %>% 
  select(Tparasites,Orange:Black)

### MHC ST frequency
mhc_f_matrix <- data_matrices %>% 
  select(ST_freq_1:ST_freq_11) %>% # Supertype columns, some have NAs, not all fish have same number of STs
  mutate_if(is.numeric, ~replace_na(., 0)) #replace NA with 0 for pca

### allelic richness and ST richness
mhc_r_matrix <- data_matrices %>% 
  select(Ai, STi)
```

**COIA(1)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-ST frequencies for each population

**COIA(2)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-Ar, and MHC-STr

Permutation test with 1000 bootstrap replications to test significance of co-variation in COIA

```{r coia}
# pcas, currently keeping all factors
pca_colour <- dudi.pca(colour_matrix, scale = TRUE, scannf = FALSE, nf = 3) # because parasite counts and relative area have different scales
pca_mhc_f <- dudi.pca(mhc_f_matrix, scale = FALSE, scannf = FALSE, nf = 10) #same scale
pca_mhc_r <- dudi.pca(mhc_r_matrix, scale = FALSE, scannf = FALSE, nf = 2) #same scale

# for both COIN analyses the screeplot indicates only the first factor is significantly contributing, should we only keep 1?

#COIN 1 - colour and MHC ST frequency
coin_1 <- coinertia(dudiX = pca_colour, dudiY = pca_mhc_f, scannf = FALSE, nf = 2) #manually put axes to keep

#COIN 2 - colour and allelic richness
coin_2 <- coinertia(dudiX = pca_colour, dudiY = pca_mhc_r, scannf = FALSE, nf = 2) #manually put axes to keep

#permutation test
rv1 <- RV.rtest(pca_colour$tab, pca_mhc_f$tab, 1000)
plot(rv1) # simulated p = 0.001, RV = 0.11

rv2 <- RV.rtest(pca_colour$tab, pca_mhc_r$tab, 1000)
plot(rv2) # simulated p = 0.374, RV = 0.014
```


Pearson's Correlations on key interactions   
I am not sure about this at all   
```{r PCC}
#correlation plot with matrices to look at what to test with a PCC?
correlations <- cor(cbind(colour_matrix, mhc_f_matrix, mhc_r_matrix))
corrplot(correlations , method = 'number')

#using 0.20 as a cutoff, St freq 8,11,3,1 vs orange
# none are significant, I am not sure if this is replicating what Jackie did
PCC_1 <- cor.test(data_analysis$Orange, data_analysis$ST_freq_1, method = "pearson")
PCC_3 <- cor.test(data_analysis$Orange, data_analysis$ST_freq_3, method = "pearson")
PCC_8 <- cor.test(data_analysis$Orange, data_analysis$ST_freq_8, method = "pearson")
PCC_11 <- cor.test(data_analysis$Orange, data_analysis$ST_freq_11, method = "pearson")
```


#### Wilcoxon Rank tests

By Predation and drainage (MHC-Ar, MHC-STr, Micro-Ar, MHC-ST frequencies; colour area, and Gyrodactylus prevalence).

These Wilcox tests are unpaired because we aren't directly comparing the high/low pred sites and include all populations from 2010 with both colour and MHC data at population level (N = 19, populations without parasites included).

```{r wilcox pop, warning=FALSE, results = 'hide'}
#specific checks to see if the data matches the paper
#wil_m_ar <- wilcox.test(high_pred_pop$`Micro-Ar`, low_pred_pop$`Micro-Ar`, paired = FALSE, alternative = "two.sided") # p = 0.010 (like in paper yay!)

#wil_para_prev <- wilcox.test(high_pred_pop$`Gyrodactylus prevalence`, low_pred_pop$`Gyrodactylus prevalence`, paired = FALSE, alternative = "two.sided") # p = 0.001 (matches paper)

wilcoxPred <- function(df, col) {
  stopifnot(typeof(col) == "character")
  wilcox.test(df[df$Predation == "High", col], df[df$Predation == "Low",col], paired = FALSE, alternative = "two.sided")
}

test_result <- wilcoxPred(PopDF, "mean_o")

wilcox_output = list()
for (column in colnames(PopDF[,6:23])) {
  test_results <- wilcoxPred(PopDF, column)
  test_results["data.name"] <- column
  print(test_results)
  wilcox_output[[length(wilcox_output)+1]] <- test_results
  #names(wilcox_output)[column] <- names(column)
}

names(wilcox_output) <- colnames(PopDF[,6:23])
#some have ties, wilcox.test cannot compute exact p-value with ties, try wilcox_test from coin package?

# by drainage
wilcoxDrain <- function(df, col) {
  stopifnot(typeof(col) == "character")
  wilcox.test(df[df$Drainage == "North Slope", col], df[df$Drainage == "Caroni",col], paired = FALSE, alternative = "two.sided")
}

test_result <- wilcoxDrain(PopDF, "mean_o")

wilcox_output_d = list()
for (column in colnames(PopDF[,6:23])) {
  test_results <- wilcoxDrain(PopDF, column)
  test_results["data.name"] <- column
  print(test_results)
  wilcox_output_d[[length(wilcox_output_d)+1]] <- test_results
  #names(wilcox_output)[column] <- names(column)
}

names(wilcox_output_d) <- colnames(PopDF[,6:23])
```


Move to linear mixed models that control for both predation and drainage.   
We could use wilcoxon with individual level data, but we have mroe than two drainages so I don't think it's appropriate anymore.   

These give information on differences in traits of interest (colour, parasite, MHC) between key ecological variables (predation and drainage).   
```{r individual level drainage and predation impact}

```

#### Linear models for population level

Using Micro-Ar as the independent variable and MHC-Ar and MHC-STr as dependent variables at the population level.
These could potentially be made individual level if we have individual microsatellite data if we want to.   

Original Results:
MHC-Ar, significant like original work
MHC-STr, insignificant in original work, now 0.06 (marginally significant?)

```{r linear models}
# p=0.0006 matches (paper = 0.001), it's significant
model_ar <- lm(`Micro-Ar` ~ `MHC-Ar`, data = PopDF)
summary(model_ar)

#p=0.06 not an exact match (paper = 0.289), not significant
#this was 0.20 (much closer to original) before I excluded the fish we didn't have colour data for
model_str <- lm(`Micro-Ar` ~ `MHC-STr`, data = PopDF)
summary(model_str)
```


### Figures

- COIA plots
- Boxplots similar to original paper
- Visualizations of correlations to investigate more thoroughly
- whatever else we want

The coin visualization is a work in progress...   
We will need to recreate the plots in the original paper to make sure they all match our data and so we can provide the source code.
```{r coia plots}
# coin 1
g1 <- s.arrow(coin_1$l1, plab.cex = 0.7)
g2 <- s.arrow(coin_1$c1, plab.cex = 0.7)
g3 <- s.corcircle(coin_1$aX, plot = FALSE)
g4 <- s.corcircle(coin_1$aY, plot = FALSE)
cbindADEg(g3, g4, plot = TRUE)
g5 <- plot(coin_1)

# coin 2
g6 <- s.arrow(coin_2$l1, plab.cex = 0.7)
g7 <- s.arrow(coin_2$c1, plab.cex = 0.7)
g8 <- s.corcircle(coin_2$aX, plot = FALSE)
g9 <- s.corcircle(coin_2$aY, plot = FALSE)
cbindADEg(g8, g9, plot = TRUE)
g10 <- plot(coin_2)
```



###