---
title: "MHC analysis"
author: "Dr. Kiyoko Gotanda"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r load packages, echo = FALSE, message = FALSE}
library(car)
library(tidyverse)
library(ade4)
library(adegraphics)
library(stringr)
library(corrplot)
library(lmerTest)
library(factoextra)
```

### Data Prep
Load dataframes
```{r load DFs, echo = TRUE}
rm(list = ls())
# Original Colour data
#ColDF <- read.csv("./2013-09-30 MHC_Colour.csv", header = T, row.names = NULL, stringsAsFactors = T, na.strings = "")
# Individual immunity data
MHCDF <- read.csv("./Full DF 2023-10-20.csv", header = T, row.names = NULL, stringsAsFactors = T, na.strings = "")
# population immunity data
PopDF <- read.csv("./microAr 2023-03-16.csv", header = F, row.names = NULL, na.strings = "")
```


```{r clean DFs, echo = FALSE}
#renaming the allele and st columns for ease of use
colnames(MHCDF) <- c("Fish","River","Site","Pred","Predation","Sex","Year","Tparasites","Infected",	"InfectedBinary", "Orange","Black","site","drainage","sex","length","n.gyros","colour_notes",	"Drainage",	"gt_1","gt_2","gt_3","gt_4","gt_5","gt_6","gt_7","gt_8","gt_9","st_1","st_2","st_3",	"st_4","st_5","st_6","st_7","st_8","st_9","Ai","STi")

MHCDF$Drainage <- as.factor(str_replace(MHCDF$Drainage, 'North slope', 'North Slope'))

#rename and clean Population df from appendix
colnames(PopDF) <- PopDF[5,] #reassign column names
PopDF <- PopDF[-(1:5), 1:21 , drop = FALSE] # remove empty rows and colour cols
PopDF$Site <- str_replace(PopDF$Site, '-', '') # standardise site names
PopDF$Site <- str_replace(PopDF$Site, 'El Cedro', 'ElCedro') 
string_cols <- colnames(PopDF[,2:4])
PopDF[string_cols] <- lapply(PopDF[string_cols] , factor)
number_cols <- colnames(PopDF[,c(1,5:21)])
PopDF[number_cols] <- lapply(PopDF[number_cols] , as.numeric)

#excluding all 2009 guppies because we had unbalanced representation across sites and years
data_analysis <- MHCDF %>% 
    filter(!is.na(Orange) & !is.na(Black) & Year == '2010' & Sex == 'Male')
```

We have colour data for `r data_analysis %>% tally() %>% pull()` photos with corresponding MHC data, all photos were taken in 2010. These span `r n_distinct(data_analysis$Site)` sites. Below are the population specific means for relative area of both orange and black.

```{r pop means, echo=FALSE, message = FALSE}
pop_means <- data_analysis %>% 
  group_by(Site) %>% 
  summarize(mean_o = mean(Orange),
            mean_b = mean(Black)) %>% 
  add_row(data.frame(Site="All", mean_o=mean(data_analysis$Orange), mean_b=mean(data_analysis$Black)))

#adding to PopDF, removing populations with no colour data
PopDF <- inner_join(PopDF, pop_means, by = 'Site')

knitr::kable(pop_means, col.names = c("Site", "Mean Relative Orange", "Mean Relative Black"), "simple")
```


```{r st freq data, echo=FALSE, message = FALSE}
# We found this data in the population level table from Jackie's thesis, but we don't have individual frequencies.

ST_freq <- MHCDF %>% 
  filter(!is.na(Orange) & !is.na(Black)) %>% 
  filter(Year == 2010 & Sex == 'Male') %>% 
  select(-Pred,-site,-sex,-n.gyros) %>% 
  pivot_longer(cols = gt_1:st_9, names_pattern = "(.t)_(.)", names_to = c('.value','position'), values_drop_na = TRUE) %>% 
  group_by(Fish, st) %>% 
  summarize(freq = n()) %>% 
  pivot_wider(names_from = st, values_from = freq, names_sort = TRUE, names_prefix = "ST_freq_") #no ST 10

data_analysis <- inner_join(data_analysis, ST_freq, by = "Fish") %>% 
  select(Fish:Site, Predation:InfectedBinary, Orange:Black, length, colour_notes, Drainage, Ai:ST_freq_15)
```


### *Reviewers Comments*

* They mentioned replicating analyses with both drainages separately so we can confirm trends across drainages   
  + COIA, I don't think we can add drainage but we could do a drainage specific COIA?   
  + Adding a drainage term in richness models?   

* Why use parametric and non-parametric tests?   
  + by switching to linear models we aren't mixing parametric and non-parametric anymore   

* Adding in the data from the different years too   
  + We did this, it allowed us to move to individual level analysis   

* Changing to Individual level analyses to leverage the new data   
  + COIA (DONE)   
  + Wilcoxon - change to models with individual data (DONE)   
  + Microsatellite - dependent on whether we have individual microsat data   

### Analysis To Do: 

###### Original Analysis   
* COIA (In Progress)   
  + prepare all 3 datasets (DONE)   
  + Run PCAs (DONE)   
  + bootstrapping for COIA to test significance (DONE)   
  + pearson's correlation coefficient test (maybe???) *I am not sure about this*   
  + Visualization (In progress)   
* richness linear models (micro-Ar as independent, Ai and STi as dependent) (DONE)   
* Wilcoxon rank-sum test on all variables by predation level (DONE)   
* Wilcoxon rank-sum test on all variables by drainage (DONE-Caroni & North)   

###### NEW and IMPROVED Analysis   
* predation and drainage tests (In Progress)   
  + move to linear mixed models that test for predation and drainage   
  + use data_analysis df with individual level data   
  + Not all individuals have all ST, drop NAs? How will this impact model?   
* microsatellite linear models   
  + check if microsatellite data has individual level in Jackie's thesis   
  + *maybe* re-do richness linear models with individual level data if available   
* boxplot and other visualization/figure recreation with individual data   


#### Co-inertia analyses

For coinertia analysis the row weighting must be matched between the two dudi (duality diagrams), so for each pca the rows must indicate the same samples. Using only fish we have both genetic data and patternize colour data for we have N = 193.
```{r data prep}
# check for uninfected populations to exclude from analyses between immunity and colour
no_parasite_sites <- data_analysis %>% 
  group_by(Site) %>% 
  summarise(max = max(Tparasites)) %>% 
  filter(max < 1) %>% 
  pull(Site)

data_matrices <-data_analysis %>% 
  filter(!Site %in% no_parasite_sites) %>% 
  filter(Site != "ElCedro1") #removed in original study, parasite prev an order of magnitude below others

### Parasite and colour
#Note: not using InfectedBinary because it should have less explanatory power than Tparasites
#Should we include body length here?
colour_matrix <- data_matrices %>% 
  select(Tparasites,Orange:Black)

### MHC ST frequency
mhc_f_matrix <- data_matrices %>% 
  select(ST_freq_1:ST_freq_11) %>% # Supertype columns, some have NAs, not all fish have same number of STs
  mutate_if(is.numeric, ~replace_na(., 0)) #replace NA with 0 for pca

### allelic richness and ST richness
mhc_r_matrix <- data_matrices %>% 
  select(Ai, STi)
```

**COIA(1)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-ST frequencies for each population

**COIA(2)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-Ar, and MHC-STr

Permutation test with 1000 bootstrap replications to test significance of co-variation in COIA

```{r coia}
# pcas, currently keeping all factors
pca_colour <- dudi.pca(colour_matrix, scale = TRUE, scannf = FALSE, nf = 3) # because parasite counts and relative area have different scales
pca_mhc_f <- dudi.pca(mhc_f_matrix, scale = FALSE, scannf = FALSE, nf = 10) #same scale
pca_mhc_r <- dudi.pca(mhc_r_matrix, scale = FALSE, scannf = FALSE, nf = 2) #same scale

# for both COIN analyses the screeplot indicates only the first factor is significantly contributing, should we only keep 1?

#COIN 1 - colour and MHC ST frequency
coin_1 <- coinertia(dudiX = pca_colour, dudiY = pca_mhc_f, scannf = FALSE, nf = 2) #manually put axes to keep

#COIN 2 - colour and allelic richness
coin_2 <- coinertia(dudiX = pca_colour, dudiY = pca_mhc_r, scannf = FALSE, nf = 2) #manually put axes to keep

#permutation test
rv1 <- RV.rtest(pca_colour$tab, pca_mhc_f$tab, 1000)
plot(rv1) # simulated p = 0.001, RV = 0.11

rv2 <- RV.rtest(pca_colour$tab, pca_mhc_r$tab, 1000)
plot(rv2) # simulated p = 0.374, RV = 0.014
```


Pearson's Correlations on key interactions   
I am not sure about this at all   
```{r PCC}
#correlation plot with matrices to look at what to test with a PCC?
correlations <- cor(cbind(colour_matrix, mhc_f_matrix, mhc_r_matrix))
corrplot(correlations , method = 'number')

#using 0.20 as a cutoff, St freq 8,11,3,1 vs orange
# none are significant, I am not sure if this is replicating what Jackie did
PCC_1 <- cor.test(data_analysis$Orange, data_analysis$ST_freq_1, method = "pearson")
PCC_3 <- cor.test(data_analysis$Orange, data_analysis$ST_freq_3, method = "pearson")
PCC_8 <- cor.test(data_analysis$Orange, data_analysis$ST_freq_8, method = "pearson")
PCC_11 <- cor.test(data_analysis$Orange, data_analysis$ST_freq_11, method = "pearson")
```


#### Wilcoxon Rank tests

By Predation and drainage (MHC-Ar, MHC-STr, Micro-Ar, MHC-ST frequencies; colour area, and Gyrodactylus prevalence).

These Wilcox tests are unpaired because we aren't directly comparing the high/low pred sites and include all populations from 2010 with both colour and MHC data at population level (N = 19, populations without parasites included).

```{r wilcox pop, warning=FALSE, results = 'hide'}
#specific checks to see if the data matches the paper
#wil_m_ar <- wilcox.test(high_pred_pop$`Micro-Ar`, low_pred_pop$`Micro-Ar`, paired = FALSE, alternative = "two.sided") # p = 0.010 (like in paper yay!)

#wil_para_prev <- wilcox.test(high_pred_pop$`Gyrodactylus prevalence`, low_pred_pop$`Gyrodactylus prevalence`, paired = FALSE, alternative = "two.sided") # p = 0.001 (matches paper)

wilcoxPred <- function(df, col) {
  stopifnot(typeof(col) == "character")
  wilcox.test(df[df$Predation == "High", col], df[df$Predation == "Low",col], paired = FALSE, alternative = "two.sided")
}

test_result <- wilcoxPred(PopDF, "mean_o")

wilcox_output = list()
for (column in colnames(PopDF[,6:23])) {
  test_results <- wilcoxPred(PopDF, column)
  test_results["data.name"] <- column
  print(test_results)
  wilcox_output[[length(wilcox_output)+1]] <- test_results
  #names(wilcox_output)[column] <- names(column)
}

names(wilcox_output) <- colnames(PopDF[,6:23])
#some have ties, wilcox.test cannot compute exact p-value with ties, try wilcox_test from coin package?

# by drainage
wilcoxDrain <- function(df, col) {
  stopifnot(typeof(col) == "character")
  wilcox.test(df[df$Drainage == "North Slope", col], df[df$Drainage == "Caroni",col], paired = FALSE, alternative = "two.sided")
}

test_result <- wilcoxDrain(PopDF, "mean_o")

wilcox_output_d = list()
for (column in colnames(PopDF[,6:23])) {
  test_results <- wilcoxDrain(PopDF, column)
  test_results["data.name"] <- column
  print(test_results)
  wilcox_output_d[[length(wilcox_output_d)+1]] <- test_results
  #names(wilcox_output)[column] <- names(column)
}

names(wilcox_output_d) <- colnames(PopDF[,6:23])
```


#### Individual level mixed models

Move to linear mixed models that control for both predation and drainage, these are just updated versions of the wilcoxon. However, we can do something nicer with a mixed model (second chunk).      
(Not in MS currently)   

```{r individual level drainage and predation only, include=FALSE, eval=FALSE}
#these are analagous to the wilcox, however we can do better, not currently used or run in this file
model_o_simple <- lmer(Orange ~ Predation + Drainage + (1|Site), data = data_analysis)
model_b_simple <- lmer(Black ~ Predation + Drainage + (1|Site), data = data_analysis)
model_para_simple <- lmer(Tparasites ~ Predation + Drainage + (1|Site), data = data_analysis)
model_mhcSTr_simple <- lmer(STi ~ Predation + Drainage + (1|Site), data = data_analysis)
model_mhcAr_simple <- lmer(Ai ~ Predation + Drainage + (1|Site), data = data_analysis)
```

These give information on differences in traits of interest (colour, parasite, MHC) between key variables (MHC, predation, and drainage).   
I've done these up for now but have not done model validation or interpretation.

Cannot do individual level microsatellite data, and population means are of more interest? We only have population estimates of Microstellite richness

*make 3 models, orange, black, and parasite as responses (MHCSTr, MHCAr, MHCSTfreq, Pred as fixes, and random with site nested within drainage)*
*Make 3 models*
- 2 for colour (black and orange as response variables)
- 1 for the impact of MHC traits on parasite abundance
This means we need a PCA for the MHC variables, this should also address the correlation between STi and Ai (significant correlation)

```{r MHC PCA}
hist(data_analysis$STi)
hist(data_analysis$Ai) # both have a slight left skew, all the ST freq will too as it's count data

#gather the variables for the PCA (numerical MHC data)
MHC_matrix<-data_analysis[,15:30] %>% 
  mutate_all(~replace(., is.na(.), 0)) # replaces all NA values with 0s in frequency data

mhc_pca<-prcomp(MHC_matrix) # when scaled the variance is much lower for the first few PCs
summary(mhc_pca)
#based on this the first one or maybe 2 PCs are worth including (PC1 54% and PC2 is 10% of variation)

## Need to arcsin transform and join pca data to analysis data
data_models<-data_analysis %>% 
  mutate(Orange_arcsin = asin(sqrt(data_analysis$Orange)),
         Black_arcsin = asin(sqrt(data_analysis$Black))) %>% # arcsin transform both Orange and Black
  cbind(as.data.frame(mhc_pca[["x"]])) # add the PC variables from mhc_pca
```

*next steps and thoughts*
try to fit models with the arcsin transformed data, still bounded between 0 and 1 but normal distribution
Moving to lmer since normal distribution
Need to do different model structures and generate model table like parrot fish paper

```{r individual level model orange, eval = FALSE}
#significant effect when Site is not a random effect, site is contributing most of the variation
model_o <- lmer(Orange_arcsin ~ Tparasites + PC1 + PC2 + Predation + (1|Drainage/Site), data = data_models)
summary(model_o)

model_o2 <- lmer(Orange_arcsin ~ PC1 + PC2 + Predation + (1|Drainage/Site), data = data_models)
summary(model_o2)

model_o3 <- lmer(Orange_arcsin ~ Tparasites + PC1 + Predation + (1|Drainage/Site), data = data_models)
summary(model_o3)

model_o4 <- lmer(Orange_arcsin ~ Tparasites + PC1 + PC2+ (1|Drainage/Site), data = data_models)
summary(model_o4)

ggplot(data = data_models, aes(x = PC1, y = Orange, colour = Predation)) + geom_point() + facet_wrap(~Drainage)
ggplot(data = data_models, aes(x = Tparasites, y = Orange, colour = Predation)) + geom_point() + facet_wrap(~Drainage)
```


```{r individual model black, eval = FALSE}
model_b <- lmer(Black_arcsin ~ Tparasites + PC1 + PC2 + Predation + (1|Drainage/Site), data = data_models)
summary(model_b)
```


parasite model is overdispersed and singular rn
```{r individual model parasites, eval = FALSE}
#Poisson dist, if overdispersed use negative binomial or correct for overdispersion in poisson dist

plot(mean(data_analysis$Tparasites), var(data_analysis$Tparasites)) #variance > mean, overdispersed

model_para <- glmer.nb(Tparasites ~ PC1 + PC2 + Predation + (1|Drainage/Site), data = data_models) #singular fit

#suggestion present this but do drainage specific models in supplementary? in MS we can note that the random effects explain very little variation but this is the appropriate model structure for the ecology of the system

model_para2 <- glmer(Tparasites ~ PC1 + PC2 + Predation + (1|Drainage/Site), data = data_models, family = poisson(link = "log")) #singular fit, pred and PC1 is significant
model_para3 <- glmer(Tparasites ~ PC1 + PC2 + Predation + (1|Site), data = data_models, family = poisson(link = "log")) #not singular when I exclude Drainage, pred and PC1 is significant

# but also take a look at the infection status rather than number of parasites?? 
#This is less ecologically relevant and we may not include it

model_infect <- glmer(InfectedBinary ~ PC1 + PC2 + Predation + (1|Drainage/Site), data = data_models, family = binomial(link = "logit")) #singular

```

#### Linear models for population level

Using Micro-Ar as the independent variable and MHC-Ar and MHC-STr as dependent variables at the population level.
These could potentially be made individual level if we have individual microsatellite data if we want to.   

Results:
MHC-Ar, significant like original work
MHC-STr, insignificant like original work

```{r linear models}
# p=0.0006 matches (paper = 0.001), it's significant
model_ar <- lm(`Micro-Ar` ~ `MHC-Ar`, data = PopDF)
summary(model_ar)

#p=0.12 not an exact match (paper = 0.289), not significant
model_str <- lm(`Micro-Ar` ~ `MHC-STr`, data = PopDF)
summary(model_str)
```


### Figures

- COIA plots
- Boxplots similar to original paper
- Visualizations of correlations to investigate more thoroughly
- whatever else we want

The coin visualization is a work in progress...   
We will need to recreate the plots in the original paper to make sure they all match our data and so we can provide the source code.
```{r coia plots}
# coin 1
g1 <- s.arrow(coin_1$l1, plab.cex = 0.7)
g2 <- s.arrow(coin_1$c1, plab.cex = 0.7)
g3 <- s.corcircle(coin_1$aX, plot = FALSE)
g4 <- s.corcircle(coin_1$aY, plot = FALSE)
cbindADEg(g3, g4, plot = TRUE)
g5 <- plot(coin_1)

# coin 2
g6 <- s.arrow(coin_2$l1, plab.cex = 0.7)
g7 <- s.arrow(coin_2$c1, plab.cex = 0.7)
g8 <- s.corcircle(coin_2$aX, plot = FALSE)
g9 <- s.corcircle(coin_2$aY, plot = FALSE)
cbindADEg(g8, g9, plot = TRUE)
g10 <- plot(coin_2)
```



###