---
title: "MHC analysis"
author: "Dr. Kiyoko Gotanda"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r load packages, echo = FALSE, message = FALSE}
library(car)
library(tidyverse)
library(ade4)
library(adegraphics)
library(stringr)
library(corrplot)
library(lmerTest)
library(factoextra)
library(MuMIn)
library(lme4)
library(performance)
library(ggplot2)
library(ggiraphExtra)
```

### Data Prep
Load dataframes
```{r load DFs, echo = TRUE}
rm(list = ls())
# Original Colour data
#ColDF <- read.csv("./2013-09-30 MHC_Colour.csv", header = T, row.names = NULL, stringsAsFactors = T, na.strings = "")
# Individual immunity data
MHCDF <- read.csv("./Full DF 2023-10-20.csv", header = T, row.names = NULL, stringsAsFactors = T, na.strings = "")
# population immunity data
PopDF <- read.csv("./microAr 2023-03-16.csv", header = F, row.names = NULL, na.strings = "")
```


```{r clean DFs, echo = FALSE}
#renaming the allele and st columns for ease of use
colnames(MHCDF) <- c("Fish","River","Site","Pred","Predation","Sex","Year","Tparasites","Infected",	"InfectedBinary", "Orange","Black","site","drainage","sex","length","n.gyros","colour_notes",	"Drainage",	"gt_1","gt_2","gt_3","gt_4","gt_5","gt_6","gt_7","gt_8","gt_9","st_1","st_2","st_3",	"st_4","st_5","st_6","st_7","st_8","st_9","Ai","STi")

MHCDF$Drainage <- as.factor(str_replace(MHCDF$Drainage, 'North slope', 'North Slope'))

#rename and clean Population df from appendix
colnames(PopDF) <- PopDF[5,] #reassign column names
PopDF <- PopDF[-(1:5), 1:21 , drop = FALSE] # remove empty rows and colour cols
PopDF$Site <- str_replace(PopDF$Site, '-', '') # standardise site names
PopDF$Site <- str_replace(PopDF$Site, 'El Cedro', 'ElCedro') 
string_cols <- colnames(PopDF[,2:4])
PopDF[string_cols] <- lapply(PopDF[string_cols] , factor)
number_cols <- colnames(PopDF[,c(1,5:21)])
PopDF[number_cols] <- lapply(PopDF[number_cols] , as.numeric)

#excluding all 2009 guppies because we had unbalanced representation across sites and years
data_analysis <- MHCDF %>% 
    filter(!is.na(Orange) & !is.na(Black) & Year == '2010' & Sex == 'Male')
```

We have colour data for `r data_analysis %>% tally() %>% pull()` photos with corresponding MHC data, all photos were taken in 2010. These span `r n_distinct(data_analysis$Site)` sites. Below are the population specific means for relative area of both orange and black.

```{r pop means, echo=FALSE, message = FALSE}
pop_means <- data_analysis %>% 
  group_by(Site) %>% 
  summarize(mean_o = mean(Orange),
            mean_b = mean(Black)) %>% 
  add_row(data.frame(Site="All", mean_o=mean(data_analysis$Orange), mean_b=mean(data_analysis$Black)))

#adding to PopDF, removing populations with no colour data
PopDF <- inner_join(PopDF, pop_means, by = 'Site')

PopDF <- PopDF %>% #create river column 
  mutate(River = as.factor(gsub('[0-9]+', '', Site))) %>% 
  mutate(Drainage_simple = gsub(' \\(Caroni type\\)', '', Drainage)) #creates a simplified column for drainage that includes turure

#knitr::kable(pop_means, col.names = c("Site", "Mean Relative Orange", "Mean Relative Black"), "simple")
```


```{r st freq data, echo=FALSE, message = FALSE}
# We found this data in the population level table from Jackie's thesis, but we don't have individual frequencies.

ST_freq <- MHCDF %>% 
  filter(!is.na(Orange) & !is.na(Black)) %>% 
  filter(Year == 2010 & Sex == 'Male') %>% 
  select(-Pred,-site,-sex,-n.gyros) %>% 
  pivot_longer(cols = gt_1:st_9, names_pattern = "(.t)_(.)", names_to = c('.value','position'), values_drop_na = TRUE) %>% 
  group_by(Fish, st) %>% 
  summarize(freq = n()) %>% 
  pivot_wider(names_from = st, values_from = freq, names_sort = TRUE, names_prefix = "ST_freq_") #no ST 10

data_analysis <- inner_join(data_analysis, ST_freq, by = "Fish") %>% 
  select(Fish:Site, Predation:InfectedBinary, Orange:Black, length, colour_notes, Drainage, Ai:ST_freq_15)
```


### *Reviewers Comments*

* They mentioned replicating analyses with both drainages separately so we can confirm trends across drainages   
  + COIA, I don't think we can add drainage but we could do a drainage specific COIA?   
  + Adding a drainage term in richness models?   

* Why use parametric and non-parametric tests?   
  + by switching to linear models we aren't mixing parametric and non-parametric anymore   

* Adding in the data from the different years too   
  + We did this, it allowed us to move to individual level analysis   

* Changing to Individual level analyses to leverage the new data   
  + COIA (DONE, but is this best?)   *look into COIA interpretation more*
  + Wilcoxon - change to models with individual data (DONE)   
  + Microsatellite - keep at pop level 

### Analysis To Do: 

###### Original Analysis   
* COIA (In Progress)   
  + prepare all 3 datasets (DONE)   
  + Run PCAs (DONE)   
  + bootstrapping for COIA to test significance (DONE)   
  + pearson's correlation coefficient test (maybe???) *I am not sure about this*   
  + Visualization (In progress)   
* richness linear models (micro-Ar as independent, Ai and STi as dependent) (DONE)   
* Wilcoxon rank-sum test on all variables by predation level (DONE)   
* Wilcoxon rank-sum test on all variables by drainage (DONE-Caroni & North)   

###### NEW and IMPROVED Analysis   
* predation and drainage tests (In Progress)   
  + move to linear mixed models that test for predation and drainage   
  + use data_analysis df with individual level data   
  + Not all individuals have all ST, drop NAs? How will this impact model?   
* microsatellite linear models   
  + check if microsatellite data has individual level in Jackie's thesis   
  + *maybe* re-do richness linear models with individual level data if available   
* boxplot and other visualization/figure recreation with individual data   
* COIA per drainage?
  + do we want to address this comment?



#### Co-inertia analyses

For coinertia analysis the row weighting must be matched between the two dudi (duality diagrams), so for each pca the rows must indicate the same samples. Using only fish we have both genetic data and patternize colour data for we have N = 193.

currently has turure 2 which was originally a caroni introduction (turure 3 was too but has no parsites)
```{r data prep}
# check for uninfected populations to exclude from analyses between immunity and colour
no_parasite_sites <- data_analysis %>% 
  group_by(Site) %>% 
  summarise(max = max(Tparasites)) %>% 
  filter(max < 1) %>% 
  pull(Site)

data_matrices <-data_analysis %>% 
  filter(!Site %in% no_parasite_sites) %>% 
  filter(Site != "ElCedro1") #removed in original study, parasite prev an order of magnitude below others

### Parasite and colour
#Note: not using InfectedBinary because it should have less explanatory power than Tparasites
#Should we include body length here?
colour_matrix <- data_matrices %>% 
  select(Tparasites,Orange:Black)

### MHC ST frequency
mhc_f_matrix <- data_matrices %>% 
  select(ST_freq_1:ST_freq_11) %>% # Supertype columns, some have NAs, not all fish have same number of STs
  mutate_if(is.numeric, ~replace_na(., 0)) #replace NA with 0 for pca

### allelic richness and ST richness
mhc_r_matrix <- data_matrices %>% 
  select(Ai, STi)
```

**COIA(1)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-ST frequencies for each population

**COIA(2)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-Ar, and MHC-STr

Permutation test with 1000 bootstrap replications to test significance of co-variation in COIA

If we were to address the drainage comment from first reviewers we would want to split these data frames where possible to be drainage specific, not sure if there's sufficient sites for that

```{r coia}
# pcas, currently keeping all factors
pca_colour <- dudi.pca(colour_matrix, scale = TRUE, scannf = FALSE, nf = 3) # because parasite counts and relative area have different scales
pca_mhc_f <- dudi.pca(mhc_f_matrix, scale = FALSE, scannf = FALSE, nf = 10) #same scale
pca_mhc_r <- dudi.pca(mhc_r_matrix, scale = FALSE, scannf = FALSE, nf = 2) #same scale

# for both COIN analyses the screeplot indicates only the first factor is significantly contributing, should we only keep 1?

#COIN 1 - colour and MHC ST frequency
coin_1 <- coinertia(dudiX = pca_colour, dudiY = pca_mhc_f, scannf = FALSE, nf = 2) #manually put axes to keep

#COIN 2 - colour and allelic richness
coin_2 <- coinertia(dudiX = pca_colour, dudiY = pca_mhc_r, scannf = FALSE, nf = 2) #manually put axes to keep

#permutation test
rv1 <- RV.rtest(pca_colour$tab, pca_mhc_f$tab, 1000)
plot(rv1) # simulated p = 0.001, RV = 0.11

rv2 <- RV.rtest(pca_colour$tab, pca_mhc_r$tab, 1000)
plot(rv2) # simulated p = 0.374, RV = 0.014
```


Pearson's Correlations on key interactions   
I am not sure about this at all   
```{r PCC}
#correlation plot with matrices to look at what to test with a PCC?
correlations <- cor(cbind(colour_matrix, mhc_f_matrix, mhc_r_matrix))
corrplot(correlations , method = 'number')

#using 0.20 as a cutoff, St freq 8,11,3,1 vs orange
# none are significant, I am not sure if this is replicating what Jackie did
cor.test(data_analysis$Orange, data_analysis$ST_freq_1, method = "pearson")
cor.test(data_analysis$Orange, data_analysis$ST_freq_3, method = "pearson")
cor.test(data_analysis$Orange, data_analysis$ST_freq_8, method = "pearson")
cor.test(data_analysis$Orange, data_analysis$ST_freq_11, method = "pearson")
```


#### individual level mixed models

Move to linear mixed models that control for both predation and drainage, these are replacements for the wilcoxon but with some more thought into model structure and questions

Edit axes so they go to the maximum predicted values by increments of 1 (12, that's the highest parasite load)
```{r individual level, include=FALSE, eval=FALSE}
data_analysis_parasites <- data_analysis %>% 
  filter(!Site %in% no_parasite_sites) %>% 
  filter(Site != "ElCedro1") %>%  #removed in original study, parasite prev an order of magnitude below others
  mutate(Site = as.factor(as.character(Site))) 

model_o <- lmer(Orange ~ Tparasites + Ai + STi + Predation + Drainage + Tparasites:Drainage + (1|Site), data = data_analysis_parasites) #north drainage and low pred almost sig
model_b <- lmer(Black ~ Tparasites + Ai + STi + Predation + Drainage + Tparasites:Drainage + (1|Site), data = data_analysis_parasites) #oropouche drainage sig

check_model(model_o)
check_model(model_b)

summary(model_o)
summary(model_b)

#visualization

#ggplot orange
 data.frame(ggeffects::ggpredict(model_o, terms = c("Tparasites [0:12]", "Drainage"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_jitter(inherit.aes = FALSE,
             data = data_analysis_parasites,
             aes(x = Tparasites, y = Orange, colour = Drainage),
             width = 0.1, height = 0.0,
             alpha = 0.5) +
  xlab("Individual Parasite Load") +
  ylab("Relative Area Orange")
 
#ggplot black
 data.frame(ggeffects::ggpredict(model_b, terms = c("Tparasites [0:12]", "Drainage"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_jitter(inherit.aes = FALSE,
             data = data_analysis_parasites,
             aes(x = Tparasites, y = Black, colour = Drainage),
             width = 0.1, height = 0.0,
             alpha = 0.5) +
  xlab("Individual Parasite Load") +
  ylab("Relative Area Black")

```

individual level no colour
```{r parasite mm}
model_para <- glmer(Tparasites ~ Ai + STi + Predation + Drainage + (1|Site), data = data_analysis_parasites, family = "poisson")

check_model(model_para)

summary(model_para)

#visualize this doesn't look right yet
data.frame(ggeffects::ggpredict(model_para, terms = c("Ai", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_point(inherit.aes = FALSE,
             data = data_analysis_parasites,
             aes(x = Ai, y = Tparasites, colour = Predation),
             alpha = 0.5) +
  xlab("MHC Allelic Richness") +
  ylab("Individual Parasite Load")
```


#### Linear models for population level

Using Micro-Ar as the independent variable and MHC-Ar and MHC-STr as dependent variables at the population level.

Results:
MHC-Ar, significant like original work
MHC-STr, insignificant like original work

```{r microsat models old, eval = false}
# p=0.0006 matches (paper = 0.001), it's significant
model_ar <- lm(`Micro-Ar` ~ `MHC-Ar`, data = PopDF)
summary(model_ar)

#p=0.12 not an exact match (paper = 0.289), not significant
model_str <- lm(`Micro-Ar` ~ `MHC-STr`, data = PopDF)
summary(model_str)
```

moving to a more appropriate mixed model to account for drainage differences. Run these with and without Turure (as it is an introduction site). 

```{r microsat mixed model}

#Predation regime may contribute variance but only has two levels and will be added as a fixed factor
model_micro <- lmer(`Micro-Ar` ~ `MHC-STr` + `MHC-Ar` + Predation + (1|Drainage/River), data = PopDF %>% filter(!Drainage == "Oropuche (Caroni type)"))

check_model(model_micro)
summary(model_micro)

#run with and without turure (since it's an introduction), conclusions are the same so use that (larger sample size)
model_micro_all <- lmer(`Micro-Ar` ~ `MHC-STr` + `MHC-Ar` + Predation + (1|Drainage_simple/River), data = PopDF)
check_model(model_micro_all)
summary(model_micro_all)

#effect size
library("r2glmm")

model_micro_all_r2 <- r2beta(model_micro_all) #marginal r2
plot(model_micro_all_r2)

#plots of these factors

data.frame(ggeffects::ggpredict(model_micro_all, terms = c("MHC-Ar", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_point(inherit.aes = FALSE,
             data = PopDF,
             aes(x = `MHC-Ar`, y = `Micro-Ar`, colour = Predation),
             alpha = 0.5) +
  xlab("Microsatellite Richness") +
  ylab("MHC Allelic Richness")

```


Population level colour models, with prevalence

There's just too few data points to do mized models at this level I think.
```{r pop mm}
PopDF_parasites <- PopDF %>% 
  filter(!Site %in% no_parasite_sites) %>% 
  filter(Site != "ElCedro1") %>%  #removed in original study, parasite prev an order of magnitude below others
  mutate(Site = as.factor(as.character(Site))) %>% 
  rename(prevalence = `Gyrodactylus prevalence`, mhc_Ai = `MHC-Ar`, mhc_STi = `MHC-STr`)

model_pop_o <- lm(mean_o ~ prevalence + mhc_Ai + mhc_STi + Predation, data = PopDF_parasites) 

model_pop_b <- lm(mean_b ~ prevalence + mhc_Ai + mhc_STi + Predation, data = PopDF_parasites) 

check_model(model_pop_o) #not good
check_model(model_pop_b)

summary(model_pop_o)
summary(model_pop_b)

#visualization
#ggplot orange
 data.frame(ggeffects::ggpredict(model_pop_o, terms = c("mhc_Ai"))) %>% 
  ggplot(aes(x = x, y = predicted)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error),fill = "orange", alpha = 0.35) +
  geom_line(aes(x = x, y = predicted), colour = "orange") +
  geom_point(inherit.aes = FALSE,
             data = PopDF_parasites,
             aes(x = mhc_Ai, y = mean_o),
             alpha = 0.5, colour = "orange") +
  xlab("MHC Allelic Richness") +
  ylab("Average Relative Area Orange")
 
#ggplot black
 data.frame(ggeffects::ggpredict(model_pop_b, terms = c("mhc_Ai"))) %>% 
  ggplot(aes(x = x, y = predicted)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted)) +
  geom_point(inherit.aes = FALSE,
             data = PopDF_parasites,
             aes(x = mhc_Ai, y = mean_b),
             alpha = 0.5) +
  xlab("MHC Allelic Richness") +
  ylab("Average Relative Area Black")
```



### Figures

- COIA plots
- Boxplots similar to original paper
- Visualizations of correlations to investigate more thoroughly
- whatever else we want

The coin visualization is a work in progress...   
We will need to recreate the plots in the original paper to make sure they all match our data and so we can provide the source code.
```{r coia plots}
# coin 1
g1 <- s.arrow(coin_1$l1, plab.cex = 0.7)
g2 <- s.arrow(coin_1$c1, plab.cex = 0.7)
g3 <- s.corcircle(coin_1$aX, plot = FALSE)
g4 <- s.corcircle(coin_1$aY, plot = FALSE)
cbindADEg(g3, g4, plot = TRUE)
g5 <- plot(coin_1)

# coin 2
g6 <- s.arrow(coin_2$l1, plab.cex = 0.7)
g7 <- s.arrow(coin_2$c1, plab.cex = 0.7)
g8 <- s.corcircle(coin_2$aX, plot = FALSE)
g9 <- s.corcircle(coin_2$aY, plot = FALSE)
cbindADEg(g8, g9, plot = TRUE)
g10 <- plot(coin_2)
```



###