---
title: "MHC analysis"
author: "Dr. Kiyoko Gotanda"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r load packages, echo = FALSE, message = FALSE}
library(car)
library(tidyverse)
library(ade4)
library(adegraphics)
library(stringr)
```

Load dataframes
```{r load DFs, echo = TRUE}
rm(list = ls())
# Original Colour data
#ColDF <- read.csv("./2013-09-30 MHC_Colour.csv", header = T, row.names = NULL, stringsAsFactors = T, na.strings = "")
# Individual immunity data
MHCDF <- read.csv("./Full DF 2023-01-23.csv", header = T, row.names = NULL, stringsAsFactors = T, na.strings = "")
# population immunity data
PopDF <- read.csv("./microAr 2023-03-16.csv", header = F, row.names = NULL, na.strings = "")
```


```{r clean DFs, echo = FALSE}
#renaming the allele and st columns for ease of use
colnames(MHCDF) <- c("Fish","River","Site","Pred","Predation","Sex","Year","Tparasites","Infected",	"InfectedBinary", "X", "Orange","Black","site","drainage","sex","length","n.gyros","colour_notes",	"Drainage",	"gt_1","gt_2","gt_3","gt_4","gt_5","gt_6","gt_7","gt_8","gt_9","st_1","st_2","st_3",	"st_4","st_5","st_6","st_7","st_8","st_9","Ai","STi")

#rename and clean Population df from appendix
colnames(PopDF) <- PopDF[5,] #reassign column names
PopDF <- PopDF[-(1:5), 1:21 , drop = FALSE] # remove empty rows and colour cols
PopDF$Site <- str_replace(PopDF$Site, '-', '') # standardise site names
PopDF$Site <- str_replace(PopDF$Site, 'El Cedro', 'ElCedro') 
string_cols <- colnames(PopDF[,2:4])
PopDF[string_cols] <- lapply(PopDF[string_cols] , factor)
number_cols <- colnames(PopDF[,c(1,5:21)])
PopDF[number_cols] <- lapply(PopDF[number_cols] , as.numeric)

#excluding all 2009 guppies because we had unbalanced representation across sites and years
data_analysis <- MHCDF %>% 
    filter(!is.na(Orange) & !is.na(Black) & Year == '2010' & Sex == 'Male')
```

We have colour data for `r data_analysis %>% tally() %>% pull()` photos with corresponding MHC data, all photos were taken in 2010. These span `r n_distinct(data_analysis$Site)` sites. Below are the population specific means for relative area of both orange and black.

```{r pop means, echo=FALSE, message = FALSE}
pop_means <- data_analysis %>% 
  group_by(Site) %>% 
  summarize(mean_o = mean(Orange),
            mean_b = mean(Black)) %>% 
  add_row(data.frame(Site="All", mean_o=mean(data_analysis$Orange), mean_b=mean(data_analysis$Black)))

#adding to PopDF, removing populations with no colour data
PopDF <- inner_join(PopDF, pop_means, by = 'Site')

knitr::kable(pop_means, col.names = c("Site", "Mean Relative Orange", "Mean Relative Black"), "simple")
```


```{r st freq data, echo=FALSE, message = FALSE}
# We found this data in the population level table from Jackie's thesis, but we don't have individual frequencies.

ST_freq <- MHCDF %>% 
  filter(!is.na(Orange) & !is.na(Black)) %>% 
  filter(Year == 2010 & Sex == 'Male') %>% 
  select(-Pred,-X,-site,-sex,-n.gyros) %>% 
  pivot_longer(cols = gt_1:st_9, names_pattern = "(.t)_(.)", names_to = c('.value','position'), values_drop_na = TRUE) %>% 
  group_by(Fish, st) %>% 
  summarize(freq = n()) %>% 
  pivot_wider(names_from = st, values_from = freq, names_sort = TRUE, names_prefix = "ST_freq_") #no ST 10

data_analysis <- inner_join(data_analysis, ST_freq, by = "Fish") %>% 
  select(Fish:Site, Predation:InfectedBinary, Orange:Black, length, colour_notes, Drainage, Ai:ST_freq_11)
```

### Analysis To Do: 
-COIN (In Progress)   
  -prepare all 3 datasets (DONE)   
  -Run PCAs (DONE)  
  -bootstrapping, and pearson's correlation coefficient?   
  -Visualization (In progress).  

- ~~Where is the microsatellite richness (micro-Ar), how do I calculate this?~~
- richness linear models (micro-Ar as independent, Ai and STi as dependent) (DONE)
- Wilcoxon rank-sum test on all variables by predation level (DONE)
- Wilcoxon rank-sum test on all variables by drainage(Caroni and North only) *I think we might not want this now?*


#### Co-inertia analyses

For coinertia analysis the row weighting must be matched between the two dudi (duality diagrams), so for each pca the rows must indicate the same samples. Using only fish we have both genetic data and patternize colour data for we have N = 193.
```{r data prep}
# check for uninfected populations to exclude from analyses between immunity and colour
no_parasite_sites <- data_analysis %>% 
  group_by(Site) %>% 
  summarise(max = max(Tparasites)) %>% 
  filter(max < 1) %>% 
  pull(Site)

data_matrices <-data_analysis %>% 
  filter(!Site %in% no_parasite_sites) %>% 
  filter(Site != "ElCedro1") #removed in original study, parasite prev an order of magnitude below others

### Parasite and colour
#Note: not using InfectedBinary because it should have less explanatory power than Tparasites
#Should we include body length here?
colour_matrix <- data_matrices %>% 
  select(Tparasites,Orange:Black)

### MHC ST frequency
mhc_f_matrix <- data_matrices %>% 
  select(ST_freq_1:ST_freq_11) %>% # Supertype columns, some have NAs, not all fish have same number of STs
  mutate_if(is.numeric, ~replace_na(., 0)) #replace NA with 0 for pca

### allelic richness and ST richness
mhc_r_matrix <- data_matrices %>% 
  select(Ai, STi)
```

**COIN(1)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-ST frequencies for each population

**COIN(2)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-Ar, and MHC-STr

1000 bootstrap replicates??

```{r coin}
# pcas
pca_colour <- dudi.pca(colour_matrix, scannf = FALSE, nf = 3)
pca_mhc_f <- dudi.pca(mhc_f_matrix, scannf = FALSE, nf = 3)
pca_mhc_r <- dudi.pca(mhc_r_matrix, scannf = FALSE, nf = 3)

#COIN 1 - colour and MHC ST frequency
coin_1 <- coinertia(dudiX = pca_colour, dudiY = pca_mhc_f, scannf = FALSE, nf = 3) #manually put in axes number to keep? 2?

#COIN 2 - colour and allelic richness
coin_2 <- coinertia(dudiX = pca_colour, dudiY = pca_mhc_r, scannf = FALSE, nf = 3) #manually put in axes number to keep? 2?
```

The coin visualization is a work in progress...   
We will need to recreate the plots in the original paper to make sure they all match our data and so we can provide the source code.
```{r coin plots}
# coin 1
g6 <- s.arrow(coin_2$l1, plab.cex = 0.7)
g7 <- s.arrow(coin_2$c1, plab.cex = 0.7)
g8 <- s.corcircle(coin_2$aX, plot = FALSE)
g9 <- s.corcircle(coin_2$aY, plot = FALSE)
cbindADEg(g8, g9, plot = TRUE)
g10 <- plot(coin_2)

# coin 2
g1 <- s.arrow(coin_1$l1, plab.cex = 0.7)
g2 <- s.arrow(coin_1$c1, plab.cex = 0.7)
g3 <- s.corcircle(coin_1$aX, plot = FALSE)
g4 <- s.corcircle(coin_1$aY, plot = FALSE)
cbindADEg(g3, g4, plot = TRUE)
g5 <- plot(coin_1)
```


#### Wilcoxon Rank tests

By Predation (MHC-Ar, MHC-STr, Micro-Ar, MHC-ST frequencies; colour area, and Gyrodactylus prevalence).

These Wilcox tests are unpaired because we aren't directly comparing the high/low pred sites and include all populations from 2010 with both colour and MHC data at population level (N = 19, populations without parasites included).

```{r wilcox predation}
#specific checks to see if the data matches the paper
#wil_m_ar <- wilcox.test(high_pred_pop$`Micro-Ar`, low_pred_pop$`Micro-Ar`, paired = FALSE, alternative = "two.sided") # p = 0.010 (like in paper yay!)

#wil_para_prev <- wilcox.test(high_pred_pop$`Gyrodactylus prevalence`, low_pred_pop$`Gyrodactylus prevalence`, paired = FALSE, alternative = "two.sided") # p = 0.001 (matches paper)

wilcoxPred <- function(df, col) {
  stopifnot(typeof(col) == "character")
  wilcox.test(df[df$Predation == "High", col], df[df$Predation == "Low",col], paired = FALSE, alternative = "two.sided")
}

test_result <- wilcoxPred(PopDF, "mean_o")

wilcox_output = list()
for (column in colnames(PopDF[,6:23])) {
  test_results <- wilcoxPred(PopDF, column)
  test_results["data.name"] <- column
  print(test_results)
  wilcox_output[[length(wilcox_output)+1]] <- test_results
  #names(wilcox_output)[column] <- names(column)
}

names(wilcox_output) <- colnames(PopDF[,6:23])
#some have ties, wilcox.test cannot compute exact p-value with ties, try wilcox_test from coin package?
```


#### Linear models

Using Micro-Ar as the independent variable and MHC-Ar and MHC-STr as dependent variables at the population level.

Original Results:
"Linear models showed a significant positive relationship between MHC-Ar and Micro-Ar across all populations (P=0.001, R2=0.41, F=14.18, df=18), suggesting that variation in MHC-Ar is strongly influenced by demographic processes (i.e. genetic drift). In contrast, the relationship between MHC-STr and Micro-Ar among all populations was not significant (P=0.289, R2=0.01, F=1.194, d.f=18)"

```{r linear models}
# p-value? matches (paper = 0.001), it's significant
model_ar <- lm(`Micro-Ar` ~ `MHC-Ar`, data = PopDF)
summary(model_ar)

#p-value? not an exact match (paper = 0.289), not significant
#this was 0.20 (much closer to original) before I excluded the fish we didn't ahve colour data for
model_str <- lm(`Micro-Ar` ~ `MHC-STr`, data = PopDF)
summary(model_str)
```

####