---
title: "MHC analysis"
author: "Dr. Kiyoko Gotanda"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r load packages, echo = FALSE}
library(car)
library(tidyverse)
library(ade4)
```

Load dataframes
```{r load DFs, echo = TRUE}
rm(list = ls())
ColDF <- read.csv("./2013-09-30 MHC_Colour.csv", header = T, row.names = NULL, stringsAsFactors = T, na.strings = "")
MHCDF <- read.csv("./Full DF 2023-01-23.csv", header = T, row.names = NULL, stringsAsFactors = T, na.strings = "")

#renaming the allele and st columns for ease of use
colnames(MHCDF) <- c("Fish","River","Site","Pred","Predation","Sex","Year","Tparasites","Infected",	"InfectedBinary", "X", "Orange","Black","site","drainage","sex","length","n.gyros","colour_notes",	"Drainage",	"gt_1","gt_2","gt_3","gt_4","gt_5","gt_6","gt_7","gt_8","gt_9","st_1","st_2","st_3",	"st_4","st_5","st_6","st_7","st_8","st_9","Ai","STi")
```

### Analysis To Do: 
-COIN (In Progress)
  -prepare all 3 datasets (DONE)   
  -Run PCAs (DONE)  
  -bootstrapping, and pearson's correlation coefficient?   
  -Visualization?   

- Where is the microsatellite richness (micro-Ar), how do I calculate this?  
- richness linear models (micro-Ar as independent, Ai and STi as dependent)
- Wilcoxon rank-sum test on all variables by predation level
- Wilcoxon rank-sum test on all variables by drainage(?) I think we might not want this now

##### ST Frequency

Need to calculate the frequency of each ST (1-15) within each individual.

```{r ST freq}
# first create a long df that's easier to work with
MHCDF_long <- MHCDF %>% 
  filter(!is.na(Orange) & !is.na(Black)) %>% 
  filter(Year == 2010 & Sex == 'Male') %>% 
  select(-Pred,-X,-site,-sex,-n.gyros) %>% 
  pivot_longer(cols = gt_1:st_9, names_pattern = "(.t)_(.)", names_to = c('.value','position'), values_drop_na = TRUE)

# next calculate the frequency of each unique supertype

```

#### Co-inertia analyses

For coinertia analysis the row weighting must be matched between the two dudi (duality diagrams), so for each pca the rows must indicate the same samples. Using only fish we have both genetic data and patternize colour data for we have N = 193.
```{r data prep}
data_for_analysis <- MHCDF %>% 
  filter(!is.na(Orange) & !is.na(Black)) %>% 
  filter(Year == 2010 & Sex == 'Male') 
#excluding all 2009 guppies because we had unbalanced representation across sites and years

# check for uninfected populations to exclude from analyses between immunity and colour
no_parasite_sites <- data_for_analysis %>% 
  group_by(Site) %>% 
  summarise(max = max(Tparasites)) %>% 
  filter(max < 1)

data_for_matrices <- data_for_analysis %>% 
  filter(!Site %in% no_parasite_sites$Site)
#Parasite and colour
#Note: not using InfectedBinary because it should have less explanatory power than Tparasites
#Should we include body length here?
colour_matrix <- data_for_matrices %>% 
  select(Tparasites,Orange:Black) # total parasites, relative body area of each colour (patternize)

#MHC ST frequency this is wrong, this is not what this is
#mhc_f_matrix <- data_for_matrices %>% 
  #select(30:38) %>% # Supertype columns, some have NAs, not all fish have same number of STs
  #mutate_if(is.numeric, ~replace_na(., 0)) #replace NA with 0 for pca

# allelic richness and ST richness
mhc_r_matrix <- data_for_matrices %>% 
  select(Ai, STi) #richness columns, number of different alleles and MHC STs
```


*COIN(1), 1000 bootstrap replicates??*

    Data matrix (1) contained MHC-ST frequencies for each population
    Data matrix (2) contained colour means and Gyrodactylus prevalence for each population
*COIN(2)*

    Data matrix (1) contained MHC-Ar, and MHC-STr
    Data matrix (2) contained colour means and Gyrodactylus prevalence for each population

```{r coin}
#COIN 1 - colour and MHC ST frequency
colour_pca <- dudi.pca(colour_matrix, scannf = FALSE, nf = 3)
#mhc_f_pca <- dudi.pca(mhc_f_matrix, scannf = FALSE, nf = 3)

#coinertia(mhc_f_pca, colour_pca, scannf = TRUE) #manually put in axes number to keep? 2?

#COIN 2 - colour and MHC ST and allelic richness
mhc_r_pca <- dudi.pca(mhc_r_matrix, scannf = FALSE, nf = 3)

coinertia(mhc_r_pca, colour_pca, scannf = TRUE) #manually put in axes number to keep? 2?
```


#### Wilcoxon Rank tests

By Predation (MHC-Ar, MHC-STr, Micro-Ar, MHC-ST frequencies; colour area, and Gyrodactylus prevalence)

Should we use all available data for non colour analyses? Or use the same dataset restricted to 2020 and fish we have colour data for?

```{r wilcox prep}
data_for_analysis <- MHCDF %>% 
  filter(!is.na(Orange) & !is.na(Black)) %>% 
  filter(Year == 2010 & Sex == 'Male')

high_pred <- data_for_analysis %>% 
  filter(Predation == 'High')
low_pred <- data_for_analysis %>% 
  filter(Predation == 'Low')

```


Actual wilcoxon tests, these are unpaired because we aren't directly comparing the high/low pred sites and include all populations from 2010 (populations without parasites too).
```{r wilcox predation}
#not comparing paired high and low so unpaired test
wil_orange <- wilcox.test(high_pred$Orange, paired = FALSE, alternative = "two.sided")
wil_black <- wilcox.test(high_pred$Black, paired = FALSE, alternative = "two.sided")

wil_para <- wilcox.test(high_pred$Tparasites, paired = FALSE, alternative = "two.sided")

wil_all_rich <- wilcox.test(high_pred$Ai, paired = FALSE, alternative = "two.sided")
wil_st_rich <- wilcox.test(high_pred$STi, paired = FALSE, alternative = "two.sided")

# Where is Micro-allelic richness data? Am I missing that?

#super type frequencies, all 15? need to calculate the frequencies of each

```

