---
title: "MHC analysis"
author: "Dr. Kiyoko Gotanda"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r load packages, echo = FALSE, message = FALSE}
library(car)
library(tidyverse)
library(ade4)
library(adegraphics)
library(stringr)
library(corrplot)
library(lmerTest)
library(factoextra)
library(MuMIn)
library(lme4)
library(performance)
library(effectsize)
library(r2glmm)
library(ggplot2)
library(ggiraphExtra)
library(corrr)
library(gtsummary)
library(gridExtra)
```

### Data Prep
Load dataframes
```{r load DFs, echo = TRUE}
rm(list = ls())
# Original Colour data
#ColDF <- read.csv("./2013-09-30 MHC_Colour.csv", header = T, row.names = NULL, stringsAsFactors = T, na.strings = "")
# Individual immunity data
MHCDF <- read.csv("./Full DF 2023-10-20.csv", header = T, row.names = NULL, stringsAsFactors = T, na.strings = "")
# population immunity data
PopDF <- read.csv("./microAr 2023-03-16.csv", header = F, row.names = NULL, na.strings = "")
```


```{r clean DFs, echo = FALSE}
#renaming the allele and st columns for ease of use
colnames(MHCDF) <- c("Fish","River","Site","Pred","Predation","Sex","Year","Tparasites","Infected",	"InfectedBinary", "Orange","Black","site","drainage","sex","length","n.gyros","colour_notes",	"Drainage",	"gt_1","gt_2","gt_3","gt_4","gt_5","gt_6","gt_7","gt_8","gt_9","st_1","st_2","st_3",	"st_4","st_5","st_6","st_7","st_8","st_9","Ai","STi")

MHCDF$Drainage <- as.factor(str_replace(MHCDF$Drainage, 'North slope', 'North Slope'))

#rename and clean Population df from appendix
colnames(PopDF) <- PopDF[5,] #reassign column names
PopDF <- PopDF[-(1:5), 1:21 , drop = FALSE] # remove empty rows and colour cols
PopDF$Site <- str_replace(PopDF$Site, '-', '') # standardise site names
PopDF$Site <- str_replace(PopDF$Site, 'El Cedro', 'ElCedro') 
string_cols <- colnames(PopDF[,2:4])
PopDF[string_cols] <- lapply(PopDF[string_cols] , factor)
number_cols <- colnames(PopDF[,c(1,5:21)])
PopDF[number_cols] <- lapply(PopDF[number_cols] , as.numeric)

#excluding all 2009 guppies because we had unbalanced representation across sites and years
data_analysis <- MHCDF %>% 
    filter(!is.na(Orange) & !is.na(Black) & Year == '2010' & Sex == 'Male')
```

We have colour data for `r data_analysis %>% tally() %>% pull()` photos with corresponding MHC data, all photos were taken in 2010. These span `r n_distinct(data_analysis$Site)` sites. Below are the sample sizes by population for colour and genetic analyses

```{r pop means, echo=FALSE, message = FALSE}
pop_means <- data_analysis %>% 
  group_by(Site) %>% 
  summarize(mean_o = mean(Orange),
            mean_b = mean(Black),
            colour_n = n()) %>% 
  add_row(data.frame(Site="All", mean_o=mean(data_analysis$Orange), mean_b=mean(data_analysis$Black), colour_n = nrow(data_analysis)))

#adding to PopDF, removing populations with no colour data#adding to PopDF, removing populations with no colour dataOrange
PopDF <- inner_join(PopDF, pop_means, by = 'Site')

PopDF <- PopDF %>% #create river column 
  mutate(River = as.factor(gsub('[0-9]+', '', Site))) %>% 
  mutate(Drainage_simple = gsub(' \\(Caroni type\\)', '', Drainage)) %>%  #creates a simplified column for drainage that includes turure
  mutate(Drainage_simple = as.factor(Drainage_simple))

knitr::kable((PopDF %>% select(c(1,2,4,21,24:26))), 
             col.names = c("Sample Number (MHC)", "Predation", "Site", "Gyrodactylus Prevalence", "Sample Number (Colour)", "River", "Drainage"), "simple")

#for export 
# write.table((PopDF %>% 
#   select(c(1, 2, 4, 21, 24:26)) %>% 
#   rename(
#     "Sample Number (MHC)" = 1,
#     "Predation" = 2,
#     "Site" = 3,
#     "Gyrodactylus Prevalence" = 4,
#     "Sample Number (Colour)" = 5,
#     "River" = 6,
#     "Drainage" = 7)),
#   "output/sample_size_table.csv", quote = FALSE, sep = ",", row.names = FALSE)
```


```{r st freq data, echo=FALSE, message = FALSE}
# We found this data in the population level table from Jackie's thesis, but we don't have individual frequencies.

ST_freq <- MHCDF %>% 
  filter(!is.na(Orange) & !is.na(Black)) %>% 
  filter(Year == 2010 & Sex == 'Male') %>% 
  select(-Pred,-site,-sex,-n.gyros) %>% 
  pivot_longer(cols = gt_1:st_9, names_pattern = "(.t)_(.)", names_to = c('.value','position'), values_drop_na = TRUE) %>% 
  group_by(Fish, st) %>% 
  summarize(freq = n()) %>% 
  pivot_wider(names_from = st, values_from = freq, names_sort = TRUE, names_prefix = "ST_freq_") #no ST 10

data_analysis <- inner_join(data_analysis, ST_freq, by = "Fish") %>% 
  select(Fish:Site, Predation:InfectedBinary, Orange:Black, length, colour_notes, Drainage, Ai:ST_freq_15)
```


currently has turure 2 which was originally a caroni introduction (turure 3 was too but has no parsites)
```{r data prep}
# check for uninfected populations to exclude from analyses between immunity and colour
no_parasite_sites <- data_analysis %>% 
  group_by(Site) %>% 
  summarise(max = max(Tparasites)) %>% 
  filter(max < 1) %>% 
  mutate(Site = droplevels(Site)) %>% # Drop unused levels
  pull(Site)

data_matrices <- data_analysis %>% 
  filter(!Site %in% no_parasite_sites) %>% 
  filter(Site != "ElCedro1") %>% #removed in original, parasite prev order of magnitude below
  mutate(Site = droplevels(Site))

### Parasite and colour
#Note: not using InfectedBinary because it should have less explanatory power than Tparasites
#Should we include body length here?
colour_matrix <- data_matrices %>% 
  select(Tparasites,Orange:Black)

### MHC ST frequency
mhc_f_matrix <- data_matrices %>% 
  select(ST_freq_1:ST_freq_11) %>% # Supertype columns, some have NAs, not all fish have same number of STs
  mutate_if(is.numeric, ~replace_na(., 0)) #replace NA with 0 for pca

### allelic richness and ST richness
mhc_r_matrix <- data_matrices %>% 
  select(Ai, STi)

#Population level
PopDF_parasites <- PopDF %>% 
  filter(!Site %in% no_parasite_sites) %>% 
  filter(Site != "ElCedro1") %>%  #removed in original study, parasite prev an order of magnitude below others
  mutate(Site = as.factor(as.character(Site))) %>% 
  rename(prevalence = `Gyrodactylus prevalence`, mhc_Ai = `MHC-Ar`, mhc_STi = `MHC-STr`)

### Parasite and colour matrices for coia
colour_matrix_pop <- PopDF_parasites %>% 
  select(prevalence,mean_o,mean_b)

### MHC ST frequency
mhc_f_matrix_pop <- PopDF_parasites %>% 
  select(ST1:ST11) %>% # Supertype columns, some have NAs, not all fish have same number of STs
  mutate_if(is.numeric, ~replace_na(., 0)) #replace NA with 0 for pca

### allelic richness and ST richness
mhc_r_matrix_pop <- PopDF_parasites %>% 
  select(mhc_Ai, mhc_STi)
```

#### Co-inertia analyses

**COIA(1)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-ST frequencies for each population

**COIA(2)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-Ar, and MHC-STr

Permutation test with 1000 bootstrap replications to test significance of co-variation in COIA

To address variation from drainage I added a within-class analysis at the drainage level, this decreased the estimate of covariation and increased the p-value (individual level)

At the population level...

```{r coia individual level}
# pcas, currently keeping all factors
pca_colour <- dudi.pca(colour_matrix, scale = TRUE, scannf = FALSE, nf = 3) # because parasite counts and relative area have different scales
pca_mhc_f <- dudi.pca(mhc_f_matrix, scale = FALSE, scannf = FALSE, nf = 10) #same scale
pca_mhc_r <- dudi.pca(mhc_r_matrix, scale = FALSE, scannf = FALSE, nf = 2) #same scale

# for both COIN analyses the screeplot indicates only the first factor is significantly contributing, should we only keep 1?

#COIN 1 - colour and MHC ST frequency
coin_1 <- coinertia(dudiY = pca_colour, dudiX = pca_mhc_f, scannf = FALSE) #manually put axes to keep

#COIN 2 - colour and allelic richness
coin_2 <- coinertia(dudiY = pca_colour, dudiX = pca_mhc_r, scannf = FALSE) #manually put axes to keep

#permutation test
rv1 <- RV.rtest(pca_colour$tab, pca_mhc_f$tab, 1000)
plot(rv1) # simulated p = 0.001, RV = 0.11

rv2 <- RV.rtest(pca_colour$tab, pca_mhc_r$tab, 1000)
plot(rv2) # simulated p = 0.374, RV = 0.014

plot(coin_1)
plot(coin_2)
```


"Within-class analyses take into account a partition of individuals into groups and focus on structures which are common to all groups." Dray & Dufour, 2007

Anlayze this data and account for drainage differences, covariation is no longer significant. Correlations between colour and parasite load are likely explained by drainage level processes
```{r within group}
#doing a within group analysis for drainage level, this only looks at common processes across groups?

wca_colour <- wca(pca_colour, fac = data_matrices$Drainage, scannf = FALSE)
wca_mhc_f <- wca(pca_mhc_f, fac = data_matrices$Drainage, scannf = FALSE) 

plot(wca_mhc_f)
plot(wca_colour)

coin_3 <- coinertia(dudiY = wca_colour, dudiX = wca_mhc_f, scannf = FALSE)
plot(coin_3)

#permutation test
rv3 <- RV.rtest(wca_colour$tab, wca_mhc_f$tab, 1000)
plot(rv3) # p-value = 0.26, rv = 0.035

#with site? Same interpretation
wca_colour_2 <- wca(pca_colour, fac = data_matrices$Site, scannf = FALSE)
wca_mhc_f_2 <- wca(pca_mhc_f, fac = data_matrices$Site, scannf = FALSE) 

plot(wca_mhc_f_2)
plot(wca_colour_2)

coin_4 <- coinertia(dudiY = wca_colour_2, dudiX = wca_mhc_f_2, scannf = FALSE)
plot(coin_4) #saved as coia_plot_2.png, 590 x 640

#permutation test
rv4 <- RV.rtest(wca_colour_2$tab, wca_mhc_f_2$tab, 1000)
plot(rv4) # rv = 0.023, p-value = 0.69

#between class analysis?
# bca_colour <- bca(pca_colour, fac = data_matrices$Drainage, scannf = FALSE)
# bca_mhc_f <- bca(pca_mhc_f, fac = data_matrices$Drainage, scannf = FALSE) 
# 
# plot(bca_mhc_f)
# plot(bca_colour)
# 
# coin_5 <- coinertia(dudiY = bca_colour, dudiX = bca_mhc_f, scannf = FALSE)
# plot(coin_5)
# 
# #permutation test
# rv5 <- RV.rtest(bca_colour$tab, bca_mhc_f$tab, 1000)
# plot(rv5) # rv = 0.70, p-value = 1


#for the other coia (2)
wca_mhc_r <- wca(pca_mhc_r, fac = data_matrices$Drainage, scannf = FALSE) 

plot(wca_mhc_r)
plot(wca_colour)

coin_6 <- coinertia(dudiY = wca_colour, dudiX = wca_mhc_r, scannf = FALSE)
plot(coin_6) #saved as coia_plot_2.png, 590 x 640

#permutation test
rv6 <- RV.rtest(wca_colour$tab, wca_mhc_r$tab, 1000)

plot(rv6) # rv = 0.61, p-value = 0.66
```

Confirmed that the same trend exists at the population level (prevalence), significant relationship when not using the within class analysis
```{r coia population level}
# pcas, currently keeping all factors
pca_colour_p <- dudi.pca(colour_matrix_pop, scale = TRUE, scannf = FALSE) # because parasite counts and relative area have different scales
pca_mhc_f_p <- dudi.pca(mhc_f_matrix_pop, scale = FALSE, scannf = FALSE)
pca_mhc_r_p <- dudi.pca(mhc_r_matrix_pop, scale = FALSE, scannf = FALSE) 

# for both COIN analyses the screeplot indicates only the first factor is significantly contributing, should we only keep 1?

#COIN 1 - colour and MHC ST frequency
coin_7 <- coinertia(dudiY = pca_colour_p, dudiX = pca_mhc_f_p, scannf = FALSE)

#COIN 2 - colour and allelic richness
coin_8 <- coinertia(dudiY = pca_colour_p, dudiX = pca_mhc_r_p, scannf = FALSE) 

#permutation test
rv7 <- RV.rtest(pca_colour_p$tab, pca_mhc_f_p$tab, 1000)
plot(rv7) # simulated p = 0.05, RV = 0.386

rv8 <- RV.rtest(pca_colour_p$tab, pca_mhc_r_p$tab, 1000)
plot(rv8) # simulated p = 0.002, RV = 0.564


## WITHIN CLASS
wca_colour_p <- wca(pca_colour_p, fac = PopDF_parasites$Drainage_simple, scannf = FALSE)
wca_mhc_f_p <- wca(pca_mhc_f_p, fac = PopDF_parasites$Drainage_simple, scannf = FALSE) 
wca_mhc_r_p <- wca(pca_mhc_r_p, fac = PopDF_parasites$Drainage_simple, scannf = FALSE) 

plot(wca_mhc_f_p)
plot(wca_mhc_r_p)
plot(wca_colour_p)

coin_9 <- coinertia(dudiY = wca_colour_p, dudiX = wca_mhc_f_p, scannf = FALSE)
plot(coin_9)

coin_10 <- coinertia(dudiY = wca_colour_p, dudiX = wca_mhc_r_p, scannf = FALSE)
plot(coin_10)

#permutation test
rv9 <- RV.rtest(wca_colour_p$tab, wca_mhc_f_p$tab, 1000)
plot(rv9) # p-value = 0.24, rv = 0.26
rv10 <- RV.rtest(wca_colour_p$tab, wca_mhc_r_p$tab, 1000)
plot(rv10) # p-value = 0.065, rv = 0.30
```

Pearson's Correlations on key interactions at population level
```{r PCC}
#correlation plot
PopDF_parasites %>% 
  correlate() %>% 
  focus(prevalence)


cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST1, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST2, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST3, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST4, method = "pearson") 
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST5, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST6, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST7, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST8, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST9, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST10, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST11, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$mean_o, method = "pearson") # strong but insig p = 0.09

PopDF_parasites %>% 
  correlate() %>% 
  focus(mean_o)

cor.test(PopDF_parasites$mean_o, PopDF_parasites$mhc_STi, method = "pearson") #sig
cor.test(PopDF_parasites$mean_o, PopDF_parasites$mhc_Ai, method = "pearson") #sig

cor.test(PopDF_parasites$prevalence, PopDF_parasites$mhc_STi, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$mhc_Ai, method = "pearson") #sig

#of the ST none correlated with orange
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST1, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST2, method = "pearson") 
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST3, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST4, method = "pearson") 
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST5, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST6, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST7, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST8, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST9, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST10, method = "pearson") 
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST11, method = "pearson") 

```


#### individual level mixed models

Linear mixed models that control for both predation and drainage

Edit axes so they go to the maximum predicted values by increments of 1 (12, that's the highest parasite load)
```{r individual level models colour}
library(gtsummary)
data_analysis_parasites <- data_analysis %>% 
  filter(!Site %in% no_parasite_sites) %>% 
  filter(Site != "ElCedro1") %>%  #removed in original study, parasite prev an order of magnitude below others
  mutate(Site = as.factor(as.character(Site))) 

model_o <- lmer(Orange ~ Tparasites + Ai + STi + Predation + Drainage + Tparasites:Drainage + (1|Site), data = data_analysis_parasites) #north drainage and low pred almost sig
model_b <- lmer(Black ~ Tparasites + Ai + STi + Predation + Drainage + Tparasites:Drainage + (1|Site), data = data_analysis_parasites) #oropouche drainage sig

#check_model(model_o)
#check_model(model_b)

#currently presenting summary statistics in text
summary(model_o)
summary(model_b)
#car::Anova(model_o)
#car::Anova(model_b)

merged_tbl <- tbl_merge(tbls = list(tbl_regression(model_o),tbl_regression(model_b)),tab_spanner = c("**Orange**", "**Black**"))

model_o_r2 <- r2beta(model_o)
model_b_r2 <- r2beta(model_b)
plot(model_o_r2)
plot(model_b_r2)

```


individual level parasite model, no colour.
```{r individual models parasites}
model_para <- glmer(Tparasites ~ Ai + STi + Predation + Drainage + (1|Site), data = data_analysis_parasites, family = "poisson")

summary(model_para)
#car::Anova(model_para)

tbl_regression(model_para, exponentiate = TRUE)

#visualization, not in MS
data.frame(ggeffects::ggpredict(model_para, terms = c("Ai", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_point(inherit.aes = FALSE,
             data = data_analysis_parasites,
             aes(x = Ai, y = Tparasites, colour = Predation),
             alpha = 0.5) +
  xlab("MHC Allelic Richness") +
  ylab("Individual Parasite Intensity")
```


#### Linear models for population level

Using Micro-Ar as the independent variable and MHC-Ar and MHC-STr as dependent variables at the population level.Using a mixed model to account for drainage differences. Run these with and without Turure (as it is an introduction site), however conclusions are the same across both so will report with complete dataset in manuscript. 

Effect size plots are in plots section.

```{r microsat mixed model no turure, eval=FALSE}
model_micro_ar <- lmer(`MHC-Ar` ~ `Micro-Ar` + Predation + (1|Drainage), data = PopDF %>% filter(!Drainage == "Oropuche (Caroni type)"))

#check_model(model_micro_ar)
summary(model_micro_ar)

model_micro_str <- lmer(`MHC-STr` ~ `Micro-Ar` + Predation + (1|Drainage), data = PopDF %>% filter(!Drainage == "Oropuche (Caroni type)"))

#check_model(model_micro_str)
summary(model_micro_str)
```


```{r microsat mixed model}
#Predation regime may contribute variance but only has two levels and will be added as a fixed factor

model_micro_all_ar <- lmer(`MHC-Ar` ~ `Micro-Ar` + Predation + (1|Drainage), data = PopDF)

#check_model(model_micro_all_ar)
summary(model_micro_all_ar)

model_micro_all_str <- lmer(`MHC-STr` ~ `Micro-Ar` + Predation + (1|Drainage), data = PopDF)

#check_model(model_micro_all_str)
summary(model_micro_all_str)

merged_tbl2 <- tbl_merge(tbls = list(tbl_regression(model_micro_all_str),tbl_regression(model_micro_all_ar)),
                        tab_spanner = c("**ST Richness**", "**Allelic Richness**"))

#effect size, plots are in plots section
model_micro_all_ar_r2 <- r2beta(model_micro_all_ar) #marginal r2
model_micro_all_str_r2 <- r2beta(model_micro_all_str) #marginal r2
```

Making other plots for reference. Manuscript has a summary table with both results (Merged_tbl2) and I think that's sufficient.
```{r microsat mixed model plots not in MS, eval=FALSE}

#model plot
data.frame(ggeffects::ggpredict(model_micro_all_ar, terms = c("Micro-Ar", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_point(inherit.aes = FALSE,
             data = PopDF,
             aes(y = `MHC-Ar`, x = `Micro-Ar`, colour = Predation),
             alpha = 0.5) +
  xlab("Microsatellite Richness") +
  ylab("MHC Allelic Richness")



data.frame(ggeffects::ggpredict(model_micro_all_str, terms = c("Micro-Ar", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_point(inherit.aes = FALSE,
             data = PopDF,
             aes(y = `MHC-STr`, x = `Micro-Ar`, colour = Predation),
             alpha = 0.5) +
  xlab("Microsatellite Richness") +
  ylab("MHC Supertype Richness")

```


### Figures

```{r coia plots}
# coin 1
g1 <- s.arrow(coin_1$l1, plab.cex = 0.7)
g2 <- s.arrow(coin_1$c1, plab.cex = 0.7)
g3 <- s.corcircle(coin_1$aX, plot = FALSE)
g4 <- s.corcircle(coin_1$aY, plot = FALSE)
cbindADEg(g3, g4, plot = TRUE)
g5 <- plot(coin_1)

# coin 2
g6 <- s.arrow(coin_2$l1, plab.cex = 0.7)
g7 <- s.arrow(coin_2$c1, plab.cex = 0.7)
g8 <- s.corcircle(coin_2$aX, plot = FALSE)
g9 <- s.corcircle(coin_2$aY, plot = FALSE)
cbindADEg(g8, g9, plot = TRUE)
g10 <- plot(coin_2)
```


Note: drainage specific hex values for figures
Fill (map) (too light for figure): 
Northern - #b9c1ef
Oropouche - #f0e8bb
Caroni - #eecdcd

Fill (figures):
Northern - #8292ed
Oropouche - #ebdb81
Caroni - #ed8a8a

Lines and points (darker):
Northern - #425ced
Oropouche - #edd442
Caroni - #ed3e3e

```{r manuscript_figures}
#Figure 1: map DONE by Allegra

# Figure 2: Drainage specific relationship between individual parasite load and relative area black.DONE by Allegra (orange in supp materials)

op <- par(mfrow = c(1,2))
#ggplot black
bplot <- data.frame(ggeffects::ggpredict(model_b, terms = c("Tparasites [0:12]", "Drainage"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group), show.legend = F) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_jitter(inherit.aes = FALSE,
             data = data_analysis_parasites,
             aes(x = Tparasites, y = Black, colour = Drainage),
             width = 0.1, height = 0.0,
             alpha = 0.5) +
  scale_color_manual(values = c("#ed3e3e","#425ced","#edd442")) +
  scale_fill_manual(values = c("#ed8a8a","#8292ed","#ebdb81")) +
  scale_x_continuous(breaks = seq(0, 12, by = 1)) +
  xlab("Individual Parasite Load") +
  ylab("Relative Area Black") +
  theme(legend.position = "none")

#ggplot orange
oplot <- data.frame(ggeffects::ggpredict(model_o, terms = c("Tparasites [0:12]", "Drainage"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_jitter(inherit.aes = FALSE,
             data = data_analysis_parasites,
             aes(x = Tparasites, y = Black, colour = Drainage),
             width = 0.1, height = 0.0,
             alpha = 0.5) +
  scale_color_manual(values = c("#ed3e3e","#425ced","#edd442")) +
  scale_fill_manual(values = c("#ed8a8a","#8292ed","#ebdb81")) +
  scale_x_continuous(breaks = seq(0, 12, by = 1)) +
  xlab("Individual Parasite Load") +
  ylab("Relative Area Orange")

# jpeg(filename =  "~/McGill/MHC Guppy Stuff/MHC R Project/MHC-colour/Figures/ColourParasite 2025-03-21.jpeg", width = 12, height = 8, units = "in", quality = 75, bg = "white", res = 300)
grid.arrange(bplot, oplot, ncol=2)
# dev.off()


#triangles for high predation (17), coloured by drainage similar to map
drainage_colours <- c(  "Caroni" = "#ed3e3e", 
  "North Slope" = "#425ced",
  "Oropuche" = "#edd442",
  "Oropuche (Caroni Type)" = "#edd442")

# Figure S5 in Jackie’s stuff, remake, and include in main manuscript. Two panels, MicroAR on Y axis, one plot with MHC Ar, and one with MHC STr. different symbols for predation.
# Figure S5. Linear models showing the relationship between MHC population allelic richness (MHC-Ar; black circles), supertype richness (MHC-STr; red circles), and microsatellite population allelic richness (Micro-Ar). A strong and significant relationship is observed among all population estimates of MHC-Ar and Micro-Ar (solid line, P=0.001, R2=0.41), and an insignificant relationship between MHC-STr and Micro-Ar (P= 0.289, R2=0.01). Open circles denote low predation site, and closed circles denote high predation sites.

# jpeg(filename =  "~/McGill/MHC Guppy Stuff/MHC R Project/MHC-colour/Figures/MS Alleles MHC 2025-03-21.jpeg", width = 12, height = 8, units = "in", quality = 75, bg = "white", res = 300)
op <- par(mfrow = c(1,2))
plot(x = PopDF_parasites$`Micro-Ar`, y = PopDF_parasites$mhc_Ai
     , xlab = "Number of Unique Microsattelite Alleles", ylab = "MHC Allelic Richness"
     , pch = ifelse(PopDF_parasites$Predation == "High", 17, 19)
     , cex = 2
     , col = drainage_colours
     )
legend(x = "topleft"
       , c("High Predation", "Low Predation")
       , pch = c(17,19)
       , bty = "n"
       , pt.bg = c("black"), pt.cex = 2, cex = 1.05)
legend(x = "left"
       , legend = names(drainage_colours)[1:3]
       , col = drainage_colours[1:3]
       , pch = 18
       , bty = "n"
       , pt.cex = 2
       , cex = 1.05)
plot(x = PopDF_parasites$`Micro-Ar`, y = PopDF_parasites$mhc_STi
     , xlab = "Number of Unique Microsattelite Alleles"
     , ylab = "Number of Unique MHC supertypes represented by alleles"
     , pch = ifelse(PopDF_parasites$Predation == "High", 17, 19)
     , cex = 2
     , col = drainage_colours
     )
par(op)
# dev.off()

# 
# Effect size visualiztion (lines 339-345) in ind. models manuscript.
#allelic richness
# R2 plot
plot(model_micro_all_ar_r2)

#supertype richness
#R2 plot
plot(model_micro_all_str_r2)
# 
# two panel plot of orange/black separate, and by drainage - multiple palnels with MHC richness, ST richness, etc.

# jpeg(filename =  "~/McGill/MHC Guppy Stuff/MHC R Project/MHC-colour/Figures/Non Effect Colour 2025-04-02.jpeg", width = 12, height = 12, units = "in", quality = 75, bg = "white", res = 300)
op <- par(mfrow = c(2,2))
#orange colours
plot(x = PopDF_parasites$mhc_STi, y = PopDF_parasites$mean_o
     , xlab = "Number of unique supertypes", ylab = "Relative Orange Area"
     , pch = ifelse(PopDF_parasites$Predation == "High", 17, 19), cex = 2
     , col = drainage_colours
     )
plot(x = PopDF_parasites$mhc_STi, y = PopDF_parasites$mean_b
     , xlab = "Number of unique supertypes", ylab = "Relative Black Area"
     , pch = ifelse(PopDF_parasites$Predation == "High", 17, 19), cex = 2
     , col = drainage_colours
     )
plot(x = PopDF_parasites$mhc_Ai, y = PopDF_parasites$mean_o
     , xlab = "Allelic Richness", ylab = "Relative Orange Area"
     , pch = ifelse(PopDF_parasites$Predation == "High", 17, 19), cex = 2
     , col = drainage_colours
     )
plot(x = PopDF_parasites$mhc_Ai, y = PopDF_parasites$mean_b
     , xlab = "Number of Unique Microsattelite Alleles", ylab = "Relative Black Area"
     , pch = ifelse(PopDF_parasites$Predation == "High", 17, 19), cex = 2
     , col = drainage_colours
     )
par(op)
# dev.off()


# 
# Co-intertia plots, two different Co inertia, original and then within class analysis. For now make a figure for both.
# 

# 
# S6 to S8 can be deleted. Redoing S1 map, Allegra has landmarks, keeping S2-S4, remake S5 (move to main manuscript)
```



