---
title: "01-data-prep"
author: "Allegra Love"
date: "2025-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#library(gtsummary)
library(tidyverse)
```


### Data Prep


Load raw data
```{r load DFs, echo = TRUE}
rm(list = ls())
# Individual immunity data
MHCDF <- read.csv("../data/Full DF 2023-10-20.csv", header = T, row.names = NULL, stringsAsFactors = T, na.strings = "")
# population immunity data
PopDF <- read.csv("../data/microAr 2023-03-16.csv", header = F, row.names = NULL, na.strings = "")
#microsatellite data
microsat_raw <- read.csv("../data/microsat_ind.csv", header = T, skip = 2)
```


```{r clean DFs, echo = FALSE}
#renaming the allele and st columns for ease of use
colnames(MHCDF) <- c("Fish","River","Site","Pred","Predation","Sex","Year","Tparasites","Infected",	"InfectedBinary", "Orange","Black","site","drainage","sex","length","n.gyros","colour_notes",	"Drainage",	"gt_1","gt_2","gt_3","gt_4","gt_5","gt_6","gt_7","gt_8","gt_9","st_1","st_2","st_3",	"st_4","st_5","st_6","st_7","st_8","st_9","Ai","STi")

MHCDF$Drainage <- as.factor(str_replace(MHCDF$Drainage, 'North slope', 'North Slope'))

#rename and clean Population df from appendix
colnames(PopDF) <- PopDF[5,] #reassign column names
PopDF <- PopDF[-(1:5), 1:21 , drop = FALSE] # remove empty rows and colour cols
PopDF$Site <- str_replace(PopDF$Site, '-', '') # standardise site names
PopDF$Site <- str_replace(PopDF$Site, 'El Cedro', 'ElCedro') 
string_cols <- colnames(PopDF[,2:4])
PopDF[string_cols] <- lapply(PopDF[string_cols] , factor)
number_cols <- colnames(PopDF[,c(1,5:21)])
PopDF[number_cols] <- lapply(PopDF[number_cols] , as.numeric)

#cleaning microsatellite data
names(microsat_raw) <- gsub(".", "_", names(microsat_raw), fixed = TRUE)
microsat <- microsat_raw %>% 
  rename(sample_simple = Sample_1) %>% 
  pivot_longer(cols = matches("Pre.")) %>% 
  mutate(allele = sub(".*_", "", value)) %>% 
  group_by(sample_simple) %>% 
  mutate(microsat_rich = n_distinct(allele)) %>% 
  distinct(sample_simple, .keep_all=TRUE) %>% 
  select(-c(value, allele, name, Date__site, Drainage, River))
```

create some basic dataframes that will be used for manipulating and creating final dataframe products for analysis.
```{r prep DFs, echo = FALSE}
#Calculating frequencies of each Supertype from genetic data
# population level data available but not individual for this data

ST_freq <- MHCDF %>% 
  filter(!is.na(Orange) & !is.na(Black)) %>% 
  filter(Year == 2010 & Sex == 'Male') %>% 
  select(-Pred,-site,-sex,-n.gyros) %>% 
  pivot_longer(cols = gt_1:st_9, names_pattern = "(.t)_(.)", names_to = c('.value','position'), values_drop_na = TRUE) %>% 
  group_by(Fish, st) %>% 
  summarize(freq = n()) %>% 
  pivot_wider(names_from = st, values_from = freq, names_sort = TRUE, names_prefix = "ST_freq_") #no ST 10

#excluding all 2009 guppies because we had unbalanced representation across sites and years
#add microsatellite data for individuals with colour data in data_analysis set
data_analysis <- MHCDF %>% 
    filter(!is.na(Orange) & !is.na(Black) & Year == '2010' & Sex == 'Male') %>% 
  inner_join(., ST_freq, by = "Fish") %>% 
  select(Fish:Site, Predation:InfectedBinary, Orange:Black, length, colour_notes, Drainage, Ai:ST_freq_15) %>% 
  inner_join(., microsat, by = join_by(Fish == sample_simple)) %>% 
  select(-Sample)

N <- nrow(data_analysis)
data_analysis$Orange_squeezed <- (((data_analysis$Orange*(N - 1)) + 0.5) / N) #squeeze transformation for beta-binomial models
data_analysis$length_sc <- datawizard::standardise(data_analysis$length) #scale length for analysis
```


```{r no parasite list, echo = FALSE}
# create a list of uninfected populations to exclude from analyses between immunity and colour
no_parasite_sites <- data_analysis %>% 
  group_by(Site) %>% 
  summarise(max = max(Tparasites)) %>% 
  filter(max < 1) %>% 
  mutate(Site = droplevels(Site)) %>% # Drop unused levels
  pull(Site)
```


### dataframe for 3. Linear models for population level
```{r pop level linear models}
#adding to pop_means to PopDF, removing populations with no colour data#adding to PopDF, removing populations with no colour dataOrange

pop_level_dat <- data_analysis %>% 
  group_by(Site) %>% 
  summarize(mean_o = mean(Orange),
            mean_b = mean(Black),
            colour_n = n()) %>% 
  add_row(data.frame(Site="All", mean_o=mean(data_analysis$Orange), mean_b=mean(data_analysis$Black), colour_n = nrow(data_analysis))) %>% 
  inner_join(., PopDF, by = 'Site') %>% #create river column 
  mutate(River = as.factor(gsub('[0-9]+', '', Site))) %>% 
  mutate(Drainage_simple = gsub(' \\(Caroni type\\)', '', Drainage)) %>%  #creates a simplified column for drainage that includes turure
  mutate(Drainage_simple = as.factor(Drainage_simple)) %>% 
  rename(gyro_prevalence = `Gyrodactylus prevalence`, mhc_Ai = `MHC-Ar`, mhc_STi = `MHC-STr`, micro_Ai = `Micro-Ar`)
```


### Matrices for 1. Co-inertia analyses

currently has turure 2 which was originally a caroni introduction (turure 3 was too but has no parsites).
```{r coia data prep}

#filter out sites with no parasites
data_matrices <- data_analysis %>% 
  filter(!Site %in% no_parasite_sites) %>% 
  filter(Site != "ElCedro1") %>% #removed in original, parasite prev order of magnitude below
  mutate(Site = droplevels(Site))

## separate into matrices that will be be used in COIA analysis for pop and indiv. level
### Parasite and colour
colour_matrix <- data_matrices %>% 
  select(Tparasites,Orange:Black)

### MHC ST frequency
mhc_f_matrix <- data_matrices %>% 
  select(ST_freq_1:ST_freq_11) %>% # Supertype columns, some have NAs, not all fish have same number of STs
  mutate_if(is.numeric, ~replace_na(., 0)) #replace NA with 0 for pca

### allelic richness and ST richness
mhc_r_matrix <- data_matrices %>% 
  select(Ai, STi)

#Population level
PopDF_parasites <- pop_level_dat %>% 
  filter(!Site %in% no_parasite_sites) %>% 
  filter(Site != "ElCedro1") %>%  #removed in original study, parasite prev an order of magnitude below others
  mutate(Site = as.factor(as.character(Site))) %>% 
  rename(prevalence = gyro_prevalence)

### Parasite and colour matrices for coia
colour_matrix_pop <- PopDF_parasites %>% 
  select(prevalence,mean_o,mean_b)

### MHC ST frequency
mhc_f_matrix_pop <- PopDF_parasites %>% 
  select(ST1:ST11) %>% # Supertype columns, some have NAs, not all fish have same number of STs
  mutate_if(is.numeric, ~replace_na(., 0)) #replace NA with 0 for pca

### allelic richness and ST richness
mhc_r_matrix_pop <- PopDF_parasites %>% 
  select(mhc_Ai, mhc_STi)
```

### dataframe for 2. individual level mixed models
```{r indiv mixed models}
data_analysis_parasites <- data_analysis %>% 
  filter(!Site %in% no_parasite_sites) %>% #remove uninfected sites
  filter(Site != "ElCedro1") %>%  #removed because parasite prev an order of magnitude below others
  mutate(Site = as.factor(as.character(Site))) 
```


### sample size table

We have colour data for `r data_analysis %>% tally() %>% pull()` photos with corresponding MHC data, all photos were taken in 2010. These span `r n_distinct(data_analysis$Site)` sites. Below are the sample sizes by population for colour and genetic analyses

```{r pop level linear models}
knitr::kable((pop_level_dat %>% select(c(1,2,4,21,24:26))), 
             col.names = c("Sample Number (MHC)", "Predation", "Site", "gyro_prevalence", "Sample Number (Colour)", "River", "Drainage"), "simple")

#for export 
# write.table((pop_level_dat %>% 
#   select(c(1, 2, 4, 21, 24:26)) %>% 
#   rename(
#     "Sample Number (MHC)" = 1,
#     "Predation" = 2,
#     "Site" = 3,
#     "Gyrodactylus Prevalence" = 4,
#     "Sample Number (Colour)" = 5,
#     "River" = 6,
#     "Drainage" = 7)),
#   "output/sample_size_table.csv", quote = FALSE, sep = ",", row.names = FALSE)
```


### putative allele counts and summary

Appendix 5 is a superset for the larger dataset used across Jackie's thesis
```{r total population set}
col_names <- c("Sample",	"Site",	"Drainage",	"a_1",	"a_2",	"a_3",	"a_4",	"a_5",	"a_6","a_7",	"a_8",	"a_9", "st_1",	"st_2",	"st_3",	"st_4",	"st_5",	"st_6",	"st_7",	"st_8",	"st_9","Ai","STi")

genotype_dat_wide <- read.csv("../data/Appendix_5.csv", header = FALSE, skip = 5) %>% 
  select(1:12,14:22,24:25) %>% 
  rename_with(~ col_names, everything())
genotype_dat_wide$Site <- str_replace(genotype_dat_wide$Site, ' ', '') # standardise site names
genotype_dat_wide$Site <- str_replace(genotype_dat_wide$Site, 'Elcedro ', 'ElCedro') 

genotype_dat <- read.csv("../data/Appendix_5.csv", header = FALSE, skip = 5) %>% 
  select(1:12,14:22,24:25) %>% 
  rename_with(~ col_names, everything()) %>% 
  pivot_longer(cols = matches("^(a|st)_\\d+$"),names_to = c(".value", "pair_index"),names_pattern = "(a|st)_(\\d+)") %>% 
  rename(allele = a, supertype = st) %>% 
  select(-pair_index) %>% 
  filter(allele != "")

genotype_dat$Site <- str_replace(genotype_dat$Site, ' ', '') # standardise site names
genotype_dat$Site <- str_replace(genotype_dat$Site, 'Elcedro ', 'ElCedro') 

good_sites <- pop_level_dat$Site
good_sites_full <- c(good_sites, "Arima2") # we don't have colour for this site but we do have genetic data?
```


7 samples short from the reported number on this old datasheet
```{r subset used in this paper}
col_names2 <- c("Drainage", "site",	"id_label",	"year",	"depth", "sample",	"final_ai", "sti", "Tparasites",
               "prevalence", "a_1",	"a_2",	"a_3",	"a_4",	"a_5",	"a_6","a_7",	"a_8",	"a_9", "st_1",	"st_2",
               "st_3",	"st_4",	"st_5",	"st_6",	"st_7",	"st_8",	"st_9")

genotype_subset <- read.csv("../data/original_sampleset.csv", header = FALSE, skip = 2) %>% 
  select(1, 3:20, 22:30) %>% 
  rename_with(~ col_names2, everything())

good_ids <- unique(genotype_subset$sample)

genotype_dat_subset <- genotype_dat %>% 
  filter(Sample %in% good_ids)

genotype_dat_wide_subset <- genotype_dat_wide %>% 
  filter(Sample %in% good_ids) %>% 
  filter(!Sample == "") # we're short 7


allele_counts <- genotype_dat_subset %>% 
  group_by(allele) %>% 
  summarize(num_individuals = n_distinct(Sample), num_populations = n_distinct(Site)) %>% 
  # Group by whether they are private or shared
  mutate(status = case_when(
    num_populations == 1 ~ "private", #unique to 1 site
    num_populations >= 2 ~ "shared" #2 or more sites
  ))

#number by population
genotype_dat_wide_subset  %>% 
  group_by(Site) %>% 
  tally()

# find total number of alleles
nrow(allele_counts) #162

#rare putative alleles (in one individual)
allele_counts %>% 
  filter(num_individuals == 1) %>% 
  nrow() #26

# private vs shared alleles
allele_counts %>% 
  count(status) # 95 private and 67 shared

site_summary <- genotype_dat_wide_subset %>% 
  group_by(Site) %>% 
  summarize(mean_Ar = mean(Ai), mean_STr = mean(STi), 
            min_Ar = min(Ai), max_Ar = max(Ai), 
            min_STr = min(STi), max_STr = max(STi))
```


### Save cleaned data for analysis script (02-data-analysis)

Analyses and their data products
1. Co-inertia analyses
mhc_f_matrix
mhc_r_matrix
colour_matrix
mhc_f_matrix_pop
mhc_r_matrix_pop
colour_matrix_pop
PopDF_parasites
data_matrices
2. individual level mixed models
data_analysis_parasites
data_analysis
3. Linear models for population level
PopDF (for population level models)


```{r saving data for analysis}
#analysis 1
save(mhc_f_matrix, mhc_r_matrix, colour_matrix, mhc_f_matrix_pop, mhc_r_matrix_pop, colour_matrix_pop, PopDF_parasites, data_matrices, file = "../output/analysis_1.RData")

#analysis 2
save(data_analysis_parasites, data_analysis, file = "../output/analysis_2.RData")

#analysis 3
save(pop_level_dat, file = "../output/analysis_3.RData")
```
