---
title: "02-data-analysis"
author: "Allegra Love"
date: "2025-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(gtsummary)
library(ade4)
library(corrr)
library(lme4)
library(r2glmm)
library(performance)
library(flextable)
library(officer)
library(glmmTMB)
library(lmerTest)

#read in cleaned data
#analysis 1
load(file = "../output/analysis_1.RData")
#analysis 2
load(file = "../output/analysis_2.RData")
#analysis 3
load(file = "../output/analysis_3.RData")
```


**To Do:**
- non-linear model for MHC diversity and parasitism
- re-run individual models with binary infected/uninfected (done, and colour models but I don't think they requested that)
- test quadratic terms for individual MHC allelic richness and supertype richness(e.g., Herdegen et al. 2014, doi:10.1111/jeb.12290)
- simple null comparison using microsatellite diversity
- add an explicit analysis across all populations, modelling colour as a function of drainage, predation regime, MHC diversity, and neutral genetic diversity, without Gyrodactylus


### 1. Co-inertia analyses

**COIA(1)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-ST frequencies for each population

**COIA(2)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-Ar, and MHC-STr

Permutation test with 1000 bootstrap replications to test significance of co-variation in COIA

To address variation from drainage I added a within-class analysis at the drainage level.

#### At the individual level...

```{r coia individual level}
# pcas, currently keeping all factors
pca_colour <- dudi.pca(colour_matrix, scale = TRUE, scannf = FALSE, nf = 3) # because parasite counts and relative area have different scales
pca_mhc_f <- dudi.pca(mhc_f_matrix, scale = FALSE, scannf = FALSE, nf = 10) #same scale
pca_mhc_r <- dudi.pca(mhc_r_matrix, scale = FALSE, scannf = FALSE, nf = 2) #same scale

# for both COIN analyses the screeplot indicates only the first factor is significantly contributing, should we only keep 1?

#COIN 1 - colour and MHC ST frequency
coin_1 <- coinertia(dudiY = pca_colour, dudiX = pca_mhc_f, scannf = FALSE) #manually put axes to keep

#COIN 2 - colour and allelic richness
coin_2 <- coinertia(dudiY = pca_colour, dudiX = pca_mhc_r, scannf = FALSE) #manually put axes to keep

#permutation test
rv1 <- RV.rtest(pca_colour$tab, pca_mhc_f$tab, 1000)
plot(rv1) # simulated p = 0.001, RV = 0.11

rv2 <- RV.rtest(pca_colour$tab, pca_mhc_r$tab, 1000)
plot(rv2) # simulated p = 0.374, RV = 0.014

plot(coin_1)
plot(coin_2)
```


"Within-class analyses take into account a partition of individuals into groups and focus on structures which are common to all groups." Dray & Dufour, 2007

Analyze this data and account for drainage differences, covariation is no longer significant. Correlations between colour and parasite load are likely explained by drainage level processes
```{r within group}
#doing a within group analysis for drainage level, this only looks at common processes across groups?

wca_colour <- wca(pca_colour, fac = data_matrices$Drainage, scannf = FALSE)
wca_mhc_f <- wca(pca_mhc_f, fac = data_matrices$Drainage, scannf = FALSE) 

plot(wca_mhc_f)
plot(wca_colour)

coin_3 <- coinertia(dudiY = wca_colour, dudiX = wca_mhc_f, scannf = FALSE)
plot(coin_3)

#permutation test
rv3 <- RV.rtest(wca_colour$tab, wca_mhc_f$tab, 1000)
plot(rv3) # p-value = 0.26, rv = 0.035

#with site? Same interpretation
wca_colour_2 <- wca(pca_colour, fac = data_matrices$Site, scannf = FALSE)
wca_mhc_f_2 <- wca(pca_mhc_f, fac = data_matrices$Site, scannf = FALSE) 

plot(wca_mhc_f_2)
plot(wca_colour_2)

coin_4 <- coinertia(dudiY = wca_colour_2, dudiX = wca_mhc_f_2, scannf = FALSE)
plot(coin_4) #saved as coia_plot_2.png, 590 x 640

#permutation test
rv4 <- RV.rtest(wca_colour_2$tab, wca_mhc_f_2$tab, 1000)
plot(rv4) # rv = 0.023, p-value = 0.69


#for the other coia (2)
wca_mhc_r <- wca(pca_mhc_r, fac = data_matrices$Drainage, scannf = FALSE) 

plot(wca_mhc_r)
plot(wca_colour)

coin_6 <- coinertia(dudiY = wca_colour, dudiX = wca_mhc_r, scannf = FALSE)
plot(coin_6) #saved as coia_plot_2.png, 590 x 640

#permutation test
rv6 <- RV.rtest(wca_colour$tab, wca_mhc_r$tab, 1000)

plot(rv6) # rv = 0.61, p-value = 0.66
```

#### At the population level...

Confirmed that the same trend exists at the population level (prevalence), significant relationship when not using the within class analysis
```{r coia population level}
# pcas, currently keeping all factors
pca_colour_p <- dudi.pca(colour_matrix_pop, scale = TRUE, scannf = FALSE) # because parasite counts and relative area have different scales
pca_mhc_f_p <- dudi.pca(mhc_f_matrix_pop, scale = FALSE, scannf = FALSE)
pca_mhc_r_p <- dudi.pca(mhc_r_matrix_pop, scale = FALSE, scannf = FALSE) 

#COIN 1 - colour and MHC ST frequency
coin_7 <- coinertia(dudiY = pca_colour_p, dudiX = pca_mhc_f_p, scannf = FALSE)

#COIN 2 - colour and allelic richness
coin_8 <- coinertia(dudiY = pca_colour_p, dudiX = pca_mhc_r_p, scannf = FALSE) 

#permutation test
rv7 <- RV.rtest(pca_colour_p$tab, pca_mhc_f_p$tab, 1000)
plot(rv7) # simulated p = 0.05, RV = 0.386

rv8 <- RV.rtest(pca_colour_p$tab, pca_mhc_r_p$tab, 1000)
plot(rv8) # simulated p = 0.002, RV = 0.564


## WITHIN CLASS
wca_colour_p <- wca(pca_colour_p, fac = PopDF_parasites$Drainage_simple, scannf = FALSE)
wca_mhc_f_p <- wca(pca_mhc_f_p, fac = PopDF_parasites$Drainage_simple, scannf = FALSE) 
wca_mhc_r_p <- wca(pca_mhc_r_p, fac = PopDF_parasites$Drainage_simple, scannf = FALSE) 

plot(wca_mhc_f_p)
plot(wca_mhc_r_p)
plot(wca_colour_p)

coin_9 <- coinertia(dudiY = wca_colour_p, dudiX = wca_mhc_f_p, scannf = FALSE)
plot(coin_9)

coin_10 <- coinertia(dudiY = wca_colour_p, dudiX = wca_mhc_r_p, scannf = FALSE)
plot(coin_10)

#permutation test
rv9 <- RV.rtest(wca_colour_p$tab, wca_mhc_f_p$tab, 1000)
plot(rv9) # p-value = 0.24, rv = 0.26
rv10 <- RV.rtest(wca_colour_p$tab, wca_mhc_r_p$tab, 1000)
plot(rv10) # p-value = 0.065, rv = 0.30
```

#### Pearson's Correlations on key interactions at population level
```{r PCC}
#correlation plot
PopDF_parasites %>% 
  correlate() %>% 
  focus(prevalence)


cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST1, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST2, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST3, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST4, method = "pearson") 
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST5, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST6, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST7, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST8, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST9, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST10, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST11, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$mean_o, method = "pearson") # strong but insig p = 0.09

PopDF_parasites %>% 
  correlate() %>% 
  focus(mean_o)

cor.test(PopDF_parasites$mean_o, PopDF_parasites$mhc_STi, method = "pearson") #sig
cor.test(PopDF_parasites$mean_o, PopDF_parasites$mhc_Ai, method = "pearson") #sig

cor.test(PopDF_parasites$prevalence, PopDF_parasites$mhc_STi, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$mhc_Ai, method = "pearson") #sig

#of the ST none correlated with orange
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST1, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST2, method = "pearson") 
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST3, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST4, method = "pearson") 
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST5, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST6, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST7, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST8, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST9, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST10, method = "pearson") 
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST11, method = "pearson") 

```


### 2. individual level mixed models

Linear mixed models that control for both predation and drainage


```{r individual level models colour}
#model_o <- lmer(Orange ~ Tparasites + Ai + STi + Predation + Drainage + Tparasites:Drainage + (1|Site), data = data_analysis_parasites) #north drainage and low pred almost sig

#we have one 0 value in the dataset which cannot be tolerated by a betabinomial distribution
N <- nrow(data_analysis_parasites)
data_analysis_parasites$Orange_squeezed <- (((data_analysis_parasites$Orange * (N - 1)) + 0.5) / N)

model_o <- glmmTMB(Orange_squeezed ~ Tparasites + Ai + STi + Predation + Drainage + Tparasites:Drainage + (1 | Site), data = data_analysis_parasites, family = beta_family())

model_b <- lmer(Black ~ Tparasites + Ai + STi + Predation + Drainage + Tparasites:Drainage + (1|Site), data = data_analysis_parasites) #oropouche drainage sig

check_model(model_o)
check_model(model_b)

#currently presenting summary statistics in text
summary(model_o)
summary(model_b)

model_b_r2 <- r2beta(model_b)
plot(model_b_r2)

```


individual level parasite model, no colour.
```{r individual models parasites}
model_para <- glmer(Tparasites ~ Ai  + STi  + Predation + Drainage + (1|Site), data = data_analysis_parasites, family = "poisson")

check_model(model_para)
summary(model_para)
#car::Anova(model_para)

tbl_regression(model_para, exponentiate = TRUE)

#visualization, not in MS
data.frame(ggeffects::ggpredict(model_para, terms = c("Ai", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_point(inherit.aes = FALSE,
             data = data_analysis_parasites,
             aes(x = Ai, y = Tparasites, colour = Predation),
             alpha = 0.5) +
  xlab("MHC Allelic Richness") +
  ylab("Individual Parasite Intensity")
```


with orthogonal polynomial term to assess the relevance of the quadratic term, marginally significant quadratic term

but if we are looking at prevalence vs number of parasites maybe doing intensity is more appropriate.
```{r quad model}
#high collinearity between polynomial terms for Ai and STI so used mdoel selection to proceed with the better model
model_both_quad <- glmer(Tparasites ~ poly(Ai, 2, raw = FALSE) + poly(STi, 2, raw = FALSE) + Predation + Drainage + (1|Site), data = data_analysis_parasites, family = "poisson") #high VIF may be a bug in performance package from using polynomials

model_ai_quad <- glmer(Tparasites ~ poly(Ai, 2, raw = FALSE) + STi + Predation + Drainage + (1|Site), 
                       data = data_analysis_parasites, family = "poisson") #convergence warning

model_sti_quad <- glmer(Tparasites ~ Ai + poly(STi, 2, raw = FALSE) + Predation + Drainage + (1|Site), 
                       data = data_analysis_parasites, family = "poisson") 

# Compare AIC scores
compare_performance(model_both_quad, model_ai_quad, model_sti_quad, rank = TRUE) # model_sti_quad has lowest AIC
AIC(model_both_quad, model_ai_quad, model_sti_quad) # AIC within 2 units but STi quad model is lower

check_model(model_sti_quad)
summary(model_sti_quad)
```


```{r quad model visualization}
tbl_regression(model_sti_quad, exponentiate = TRUE)

data.frame(ggeffects::predict_response(model_sti_quad, terms = c("STi [all]", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2) +
  geom_line(aes(colour = group), linewidth = 1) +
  geom_point(data = data_analysis_parasites,
             aes(x = STi, y = Tparasites, colour = Predation),
             inherit.aes = FALSE,
             alpha = 0.4, 
             position = position_jitter(width = 0.1, height = 0.2)) +
  labs(
    x = "STi", # Or change to your actual variable name
    y = "Individual Parasite Load",
    fill = "Predation",
    colour = "Predation"
  )
```

### 2.2 individual models binary infection variable


```{r individual level infection status colour}
model_o_bin <- glmmTMB(Orange_squeezed ~ Infected + Ai + STi + Predation + Drainage + Infected:Drainage + (1|Site), data = data_analysis_parasites, family = beta_family())
#north drainage and low pred sig
model_b_bin <- lmer(Black ~ Infected + Ai + STi + Predation + Drainage + Infected:Drainage + (1|Site), data = data_analysis_parasites) #oropouche drainage sig

check_model(model_o_bin)
check_model(model_b_bin)

#currently presenting summary statistics in text
summary(model_o_bin)
summary(model_b_bin)
#car::Anova(model_o_bin)
#car::Anova(model_b_bin)

merged_tbl <- tbl_merge(tbls = list(tbl_regression(model_o_bin),tbl_regression(model_b_bin)),tab_spanner = c("**Orange**", "**Black**"))

model_o_r2_bin <- r2beta(model_o_bin)
model_b_r2_bin <- r2beta(model_b_bin)
plot(model_o_r2_bin)
plot(model_b_r2_bin)

```

individual level parasite (infected binary) model, no colour.
```{r individual models parasites}
model_para_bin <- glmer(Infected ~ Ai + STi + Predation + Drainage + (1|Site), data = data_analysis_parasites, family = "binomial") #singular fit due to site variance estimated as 0, model checks are okay

check_model(model_para_bin)
summary(model_para_bin)
#car::Anova(model_para_bin)

tbl_regression(model_para_bin, exponentiate = TRUE)

#visualization, not in MS
data.frame(ggeffects::ggpredict(model_para_bin, terms = c("Ai", "Predation"))) %>% 
ggplot( aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.35) +
  geom_line(aes(colour = group), size = 1) +
  geom_point(inherit.aes = FALSE,
             data = data_analysis_parasites,
             aes(x = Ai, y = InfectedBinary, colour = Predation), 
             alpha = 0.5,
             position = position_jitter(height = 0.02, width = 0)) + 
  xlab("MHC Allelic Richness") +
  ylab("Infection Status")
```
binary model with quadratic term, not sure this is required. Maybe in supplement?

```{r quad binary parasite model}
model_both_quad_bin <- glmer(Infected ~ poly(Ai, 2, raw = FALSE) + poly(STi, 2, raw = FALSE) + Predation + Drainage + (1|Site), data = data_analysis_parasites, family = "binomial") #high VIF may be a bug in performance package from using polynomials

model_ai_quad_bin <- glmer(Infected ~ poly(Ai, 2, raw = FALSE) + STi + Predation + Drainage + (1|Site), 
                       data = data_analysis_parasites, family = "binomial") #convergence warning

model_sti_quad_bin <- glmer(Infected ~ Ai + poly(STi, 2, raw = FALSE) + Predation + Drainage + (1|Site), 
                       data = data_analysis_parasites, family = "binomial") 

# Compare AIC scores
compare_performance(model_both_quad_bin, model_ai_quad_bin, model_sti_quad_bin, rank = TRUE)
AIC(model_both_quad_bin, model_ai_quad_bin, model_sti_quad_bin) # model_sti_quad has similar AIC so use this to parallel intensity model

check_model(model_sti_quad_bin)
summary(model_sti_quad_bin)
```

```{r binary quad model visualization}
tbl_regression(model_sti_quad_bin, exponentiate = TRUE)

data.frame(ggeffects::predict_response(model_sti_quad_bin, terms = c("STi [all]", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2) +
  geom_line(aes(colour = group), linewidth = 1) +
  geom_point(data = data_analysis_parasites,
             aes(x = STi, y = InfectedBinary, colour = Predation),
             inherit.aes = FALSE,
             alpha = 0.4, 
             position = position_jitter(width = 0.1, height = 0.2)) +
  labs(
    x = "STi", # Or change to your actual variable name
    y = "Individual Infection incidence",
    fill = "Predation",
    colour = "Predation"
  )
```


### 3. Linear models for population level

Using Micro-Ar as the independent variable and MHC-Ar and MHC-STr as dependent variables at the population level. Using a mixed model to account for drainage differences. Run these with and without Turure (as it is an introduction site), however conclusions are the same across both so will report with complete dataset in manuscript. 

```{r microsat mixed model no turure, eval=FALSE}
model_micro_ar <- lmer(`MHC-Ar` ~ `Micro-Ar` + Predation + (1|Drainage), data = PopDF %>% filter(!Drainage == "Oropuche (Caroni type)"))

#check_model(model_micro_ar)
summary(model_micro_ar)

model_micro_str <- lmer(`MHC-STr` ~ `Micro-Ar` + Predation + (1|Drainage), data = PopDF %>% filter(!Drainage == "Oropuche (Caroni type)"))

#check_model(model_micro_str)
summary(model_micro_str)
```


```{r microsat mixed model}
#Predation regime may contribute variance but only has two levels and will be added as a fixed factor

model_micro_all_ar <- lmer(`MHC-Ar` ~ `Micro-Ar` + Predation + (1|Drainage), data = pop_level_dat)

#check_model(model_micro_all_ar)
summary(model_micro_all_ar)

model_micro_all_str <- lmer(`MHC-STr` ~ `Micro-Ar` + Predation + (1|Drainage), data = pop_level_dat)

#check_model(model_micro_all_str)
summary(model_micro_all_str)

merged_tbl2 <- tbl_merge(tbls = list(tbl_regression(model_micro_all_str),tbl_regression(model_micro_all_ar)),
                        tab_spanner = c("**ST Richness**", "**Allelic Richness**"))

#effect size, plots are in plots section
model_micro_all_ar_r2 <- r2beta(model_micro_all_ar) #marginal r2
model_micro_all_str_r2 <- r2beta(model_micro_all_str) #marginal r2
```
Null model comparison

Use microsatellite instead of MHC to see if that explains trends from MHC in colour, AIC modle comparison with MHC models?
Darinage removed due to high collinearity
```{r null model pop orange}
model_o_null_para <- glmmTMB(mean_o ~ Predation + gyro_prevalence + `MHC-Ar` + `MHC-STr` + `Micro-Ar`, dispformula = ~ 1, data = pop_level_dat, family = beta_family())
model_o_null_mhc<- glmmTMB(mean_o ~ Predation + `MHC-Ar` + `MHC-STr` + `Micro-Ar`, dispformula = ~ 1, data = pop_level_dat, family = beta_family())

check_model(model_o_null_para)
check_model(model_o_null_mhc)

summary(model_o_null_para)
summary(model_o_null_mhc)

compare_performance(model_o_null_para, model_o_null_mhc, rank = T)
```

```{r null model pop black}
model_b_null_para <- lm(mean_b ~ Predation + gyro_prevalence + `MHC-Ar` + `MHC-STr` + `Micro-Ar` , data = pop_level_dat)
model_b_null_mhc <- lm(mean_b ~ Predation + `MHC-Ar` + `MHC-STr` + `Micro-Ar` , data = pop_level_dat)

check_model(model_b_null_para)
check_model(model_b_null_mhc)

summary(model_b_null_para)
summary(model_b_null_mhc)
#car::Anova(model_b_null)
compare_performance(model_b_null_para, model_b_null_mhc, rank = T)
```

Export of table for all population model comparison with and without parasites for both colours.
```{r colour all pop tables}

# Run comparison for Black models
perf_black <- data.frame(compare_performance(model_b_null_para, model_b_null_mhc))

# Extract predictor variable names
perf_black$Variables <- sapply(list(model_b_null_para, model_b_null_mhc), function(mod) {
  paste(labels(terms(mod)), collapse = ", ")
})

# Filter and reorder columns
black_table <- perf_black[, c("Name", "Variables", "R2", "AIC_wt")]

# Run comparison for Orange models (swapping _b_ to _o_)
perf_orange <- data.frame(compare_performance(model_o_null_para, model_o_null_mhc))

# Extract predictor variable names
perf_orange$Variables <- sapply(list(model_o_null_para, model_o_null_mhc), function(mod) {
  paste(labels(terms(mod)), collapse = ", ")
})

# Filter and reorder columns
orange_table <- perf_orange[, c("Name", "Variables", "R2", "AIC_wt")]

# Create and format Black flextable
ft_black <- flextable(black_table)
ft_black <- theme_vanilla(ft_black)
ft_black <- autofit(ft_black)
ft_black <- bold(ft_black, part = "header")
ft_black <- align(ft_black, align = "center", part = "all")
ft_black <- bg(ft_black, bg = "#f2f2f2", i = ~ seq_len(nrow(black_table)) %% 2 == 0)

# Create and format Orange flextable
ft_orange <- flextable(orange_table)
ft_orange <- theme_vanilla(ft_orange)
ft_orange <- autofit(ft_orange)
ft_orange <- bold(ft_orange, part = "header")
ft_orange <- align(ft_orange, align = "center", part = "all")
ft_orange <- bg(ft_orange, bg = "#f2f2f2", i = ~ seq_len(nrow(orange_table)) %% 2 == 0)

# Initialize a new Word document object
doc <- read_docx()

# Add Black Table section
doc <- body_add_par(doc, "Black Population Models", style = "heading 1")
doc <- body_add_flextable(doc, value = ft_black)

# Add spacing between the tables
doc <- body_add_par(doc, "", style = "Normal") 

# Add Orange Table section
doc <- body_add_par(doc, "Orange Population Models", style = "heading 1")
doc <- body_add_flextable(doc, value = ft_orange)

# Save the final consolidated file
print(doc, target = "../output/colour_allpop_table.docx")

```


Comparison of models with neutral genetic diversity (microsatellites) and MHC diversity (allelic and supertype richness), This is completed at the population level only as we have only population averages for microsatellite diversity.
```{r null with parasites}
model_para_null <- lm(prevalence ~ Predation + `Micro-Ar` , data = PopDF_parasites) 
model_para_null2 <- lm(prevalence ~ Predation + mhc_Ai + mhc_STi + `Micro-Ar` , data = PopDF_parasites) 
model_para_null3 <- lm(prevalence ~ Predation + mhc_Ai + mhc_STi, data = PopDF_parasites) 

# model_para_null <- lm(gyro_prevalence ~ Predation + `Micro-Ar` , data = pop_level_dat) 
# model_para_null2 <- lm(gyro_prevalence ~ Predation + `MHC-Ar` + `MHC-STr` + `Micro-Ar` , data = pop_level_dat) 
# model_para_null3 <- lm(gyro_prevalence ~ Predation + `MHC-Ar` + `MHC-STr` , data = pop_level_dat) 

summary(model_para_null)
summary(model_para_null2)
summary(model_para_null3)

compare_performance(model_para_null, model_para_null2, model_para_null3, rank = T) # model without microsat is best
```

model comparison table export for null model with and without microsat to determine how much explanatory power MHC diversity adds for prevalence. 
```{r table for parasite comp}
performance_table <- data.frame(compare_performance(model_para_null, model_para_null2, model_para_null3))
performance_table$Variables <- sapply(list(model_para_null, model_para_null2, model_para_null3), function(mod) {
  # Get the formula terms, drop the response variable
  paste(labels(terms(mod)), collapse = ", ")
})

#select and reorder columns
final_table <- performance_table[, c("Name", "Variables", "R2", "AIC_wt")]

# create flextable object and do basic formatting
ft <- flextable(final_table)
ft <- theme_vanilla(ft)
ft <- autofit(ft)
ft <- bold(ft, part = "header")
ft <- align(ft, align = "center", part = "all")
ft <- bg(ft, bg = "#f2f2f2", i = ~ seq_len(nrow(final_table)) %% 2 == 0) # Alternating row colors

# Save as a word doc
doc <- read_docx()
doc <- body_add_flextable(doc, value = ft)
print(doc, target = "../output/parasite_nullmodel_table.docx")
```

### reviewer 3 comments

is gaussian distribution appropriate for colour models
```{r distribution comment}
# black is normally distributed
ggplot(data_analysis_parasites, aes(Black)) + 
  geom_histogram(colour = "grey25", binwidth = .03) + 
  geom_density(colour = "coral") +
  labs( x = "Relative Area Black", y = "Count")

#left skew for orange, models changed to beta-binomial distribution as suggested by reviewer
ggplot(data_analysis_parasites, aes(Orange)) + 
  geom_histogram(fill = "coral", colour = "grey25", binwidth = .03) + 
  geom_density(colour = "grey25") +
  labs( x = "Relative Area Orange", y = "Count")

```



### Save outputs for figures

```{r figure outputs}
save(model_b, model_o, data_analysis_parasites, file = "../output/fig2.RData")

save(PopDF_parasites, file = "../output/fig3.Rdata")

save(model_micro_all_str_r2, model_micro_all_ar_r2, model_micro_all_ar, model_micro_all_str, file = "../output/microsat_figs.RData")

save(coin_2, coin_9, coin_10, file = "../output/coia_supp_figs.RData")
```


### putative allele counts and summary

```{r total population set}
col_names <- c("Sample",	"Site",	"Drainage",	"a_1",	"a_2",	"a_3",	"a_4",	"a_5",	"a_6","a_7",	"a_8",	"a_9", "st_1",	"st_2",	"st_3",	"st_4",	"st_5",	"st_6",	"st_7",	"st_8",	"st_9","Ai","STi")
genotype_dat <- read.csv("../data/Appendix_5.csv", header = FALSE, skip = 5) %>% 
  select(1:12,14:22,24:25) %>% 
  rename_with(~ col_names, everything()) %>% 
  pivot_longer(
    cols = matches("^(a|st)_\\d+$"),
    names_to = c(".value", "pair_index"),
    names_pattern = "(a|st)_(\\d+)"
  ) %>% 
  rename(allele = a, supertype = st) %>% 
  select(-pair_index) %>% 
  filter(allele != "")

genotype_dat$Site <- str_replace(genotype_dat$Site, ' ', '') # standardise site names
genotype_dat$Site <- str_replace(genotype_dat$Site, 'Elcedro ', 'ElCedro') 

good_sites <- pop_level_dat$Site

allele_counts <- genotype_dat %>% 
  filter(Site %in% good_sites) %>% 
  group_by(allele) %>% 
  summarize(num_individuals = n_distinct(Sample), num_populations = n_distinct(Site)) %>% 
  # Group by whether they are private or shared
  mutate(status = case_when(
    num_populations == 1 ~ "private", #unique to 1 site
    num_populations >= 2 ~ "shared" #2 or more sites
  ))

# find total number of alleles
nrow(allele_counts) # 177

#rare putative alleles (in one individual)
allele_counts %>% 
  filter(num_individuals == 1) %>% 
  nrow() #32

# private vs shared alleles
allele_counts %>% 
  count(status) # 99 private and 78 shared
```


```{r only parasite pops}
#only parasite populations
gyro_sites <- pop_level_dat %>% 
  filter(gyro_prevalence > 0) %>% 
  pull(Site)

allele_counts_gyro <- genotype_dat %>% 
  filter(Site %in% gyro_sites) %>% 
  group_by(allele) %>% 
  summarize(num_individuals = n_distinct(Sample), num_populations = n_distinct(Site)) %>% 
  # Group by whether they are private or shared
  mutate(status = case_when(
    num_populations == 1 ~ "private", #unique to 1 site
    num_populations >= 2 ~ "shared" #2 or more sites
  ))

# find total number of alleles
nrow(allele_counts_gyro) # 137

#rare putative alleles (in one individual)
allele_counts_gyro %>% 
  filter(num_individuals == 1) %>% 
  nrow() #28

# private vs shared alleles
allele_counts_gyro %>% 
  count(status) # 78 private and 59 shared

```

