---
title: "02-data-analysis"
author: "Allegra Love"
date: "2025-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(gtsummary)
library(ade4)
library(corrr)
library(lme4)
library(r2glmm)
```

### 1. Co-inertia analyses

```{r coia data include = FALSE}
load("../output/analysis_1.RData")
```


**COIA(1)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-ST frequencies for each population

**COIA(2)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-Ar, and MHC-STr

Permutation test with 1000 bootstrap replications to test significance of co-variation in COIA

To address variation from drainage I added a within-class analysis at the drainage level.

#### At the individual level...

```{r coia individual level}
# pcas, currently keeping all factors
pca_colour <- dudi.pca(colour_matrix, scale = TRUE, scannf = FALSE, nf = 3) # because parasite counts and relative area have different scales
pca_mhc_f <- dudi.pca(mhc_f_matrix, scale = FALSE, scannf = FALSE, nf = 10) #same scale
pca_mhc_r <- dudi.pca(mhc_r_matrix, scale = FALSE, scannf = FALSE, nf = 2) #same scale

# for both COIN analyses the screeplot indicates only the first factor is significantly contributing, should we only keep 1?

#COIN 1 - colour and MHC ST frequency
coin_1 <- coinertia(dudiY = pca_colour, dudiX = pca_mhc_f, scannf = FALSE) #manually put axes to keep

#COIN 2 - colour and allelic richness
coin_2 <- coinertia(dudiY = pca_colour, dudiX = pca_mhc_r, scannf = FALSE) #manually put axes to keep

#permutation test
rv1 <- RV.rtest(pca_colour$tab, pca_mhc_f$tab, 1000)
plot(rv1) # simulated p = 0.001, RV = 0.11

rv2 <- RV.rtest(pca_colour$tab, pca_mhc_r$tab, 1000)
plot(rv2) # simulated p = 0.374, RV = 0.014

plot(coin_1)
plot(coin_2)
```


"Within-class analyses take into account a partition of individuals into groups and focus on structures which are common to all groups." Dray & Dufour, 2007

Analyze this data and account for drainage differences, covariation is no longer significant. Correlations between colour and parasite load are likely explained by drainage level processes
```{r within group}
#doing a within group analysis for drainage level, this only looks at common processes across groups?

wca_colour <- wca(pca_colour, fac = data_matrices$Drainage, scannf = FALSE)
wca_mhc_f <- wca(pca_mhc_f, fac = data_matrices$Drainage, scannf = FALSE) 

plot(wca_mhc_f)
plot(wca_colour)

coin_3 <- coinertia(dudiY = wca_colour, dudiX = wca_mhc_f, scannf = FALSE)
plot(coin_3)

#permutation test
rv3 <- RV.rtest(wca_colour$tab, wca_mhc_f$tab, 1000)
plot(rv3) # p-value = 0.26, rv = 0.035

#with site? Same interpretation
wca_colour_2 <- wca(pca_colour, fac = data_matrices$Site, scannf = FALSE)
wca_mhc_f_2 <- wca(pca_mhc_f, fac = data_matrices$Site, scannf = FALSE) 

plot(wca_mhc_f_2)
plot(wca_colour_2)

coin_4 <- coinertia(dudiY = wca_colour_2, dudiX = wca_mhc_f_2, scannf = FALSE)
plot(coin_4) #saved as coia_plot_2.png, 590 x 640

#permutation test
rv4 <- RV.rtest(wca_colour_2$tab, wca_mhc_f_2$tab, 1000)
plot(rv4) # rv = 0.023, p-value = 0.69


#for the other coia (2)
wca_mhc_r <- wca(pca_mhc_r, fac = data_matrices$Drainage, scannf = FALSE) 

plot(wca_mhc_r)
plot(wca_colour)

coin_6 <- coinertia(dudiY = wca_colour, dudiX = wca_mhc_r, scannf = FALSE)
plot(coin_6) #saved as coia_plot_2.png, 590 x 640

#permutation test
rv6 <- RV.rtest(wca_colour$tab, wca_mhc_r$tab, 1000)

plot(rv6) # rv = 0.61, p-value = 0.66
```

#### At the population level...

Confirmed that the same trend exists at the population level (prevalence), significant relationship when not using the within class analysis
```{r coia population level}
# pcas, currently keeping all factors
pca_colour_p <- dudi.pca(colour_matrix_pop, scale = TRUE, scannf = FALSE) # because parasite counts and relative area have different scales
pca_mhc_f_p <- dudi.pca(mhc_f_matrix_pop, scale = FALSE, scannf = FALSE)
pca_mhc_r_p <- dudi.pca(mhc_r_matrix_pop, scale = FALSE, scannf = FALSE) 

#COIN 1 - colour and MHC ST frequency
coin_7 <- coinertia(dudiY = pca_colour_p, dudiX = pca_mhc_f_p, scannf = FALSE)

#COIN 2 - colour and allelic richness
coin_8 <- coinertia(dudiY = pca_colour_p, dudiX = pca_mhc_r_p, scannf = FALSE) 

#permutation test
rv7 <- RV.rtest(pca_colour_p$tab, pca_mhc_f_p$tab, 1000)
plot(rv7) # simulated p = 0.05, RV = 0.386

rv8 <- RV.rtest(pca_colour_p$tab, pca_mhc_r_p$tab, 1000)
plot(rv8) # simulated p = 0.002, RV = 0.564


## WITHIN CLASS
wca_colour_p <- wca(pca_colour_p, fac = PopDF_parasites$Drainage_simple, scannf = FALSE)
wca_mhc_f_p <- wca(pca_mhc_f_p, fac = PopDF_parasites$Drainage_simple, scannf = FALSE) 
wca_mhc_r_p <- wca(pca_mhc_r_p, fac = PopDF_parasites$Drainage_simple, scannf = FALSE) 

plot(wca_mhc_f_p)
plot(wca_mhc_r_p)
plot(wca_colour_p)

coin_9 <- coinertia(dudiY = wca_colour_p, dudiX = wca_mhc_f_p, scannf = FALSE)
plot(coin_9)

coin_10 <- coinertia(dudiY = wca_colour_p, dudiX = wca_mhc_r_p, scannf = FALSE)
plot(coin_10)

#permutation test
rv9 <- RV.rtest(wca_colour_p$tab, wca_mhc_f_p$tab, 1000)
plot(rv9) # p-value = 0.24, rv = 0.26
rv10 <- RV.rtest(wca_colour_p$tab, wca_mhc_r_p$tab, 1000)
plot(rv10) # p-value = 0.065, rv = 0.30
```

#### Pearson's Correlations on key interactions at population level
```{r PCC}
#correlation plot
PopDF_parasites %>% 
  correlate() %>% 
  focus(prevalence)


cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST1, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST2, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST3, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST4, method = "pearson") 
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST5, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST6, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST7, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST8, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST9, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST10, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST11, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$mean_o, method = "pearson") # strong but insig p = 0.09

PopDF_parasites %>% 
  correlate() %>% 
  focus(mean_o)

cor.test(PopDF_parasites$mean_o, PopDF_parasites$mhc_STi, method = "pearson") #sig
cor.test(PopDF_parasites$mean_o, PopDF_parasites$mhc_Ai, method = "pearson") #sig

cor.test(PopDF_parasites$prevalence, PopDF_parasites$mhc_STi, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$mhc_Ai, method = "pearson") #sig

#of the ST none correlated with orange
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST1, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST2, method = "pearson") 
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST3, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST4, method = "pearson") 
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST5, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST6, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST7, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST8, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST9, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST10, method = "pearson") 
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST11, method = "pearson") 

```


### 2. individual level mixed models

Linear mixed models that control for both predation and drainage

```{r indiv lmms data include = FALSE}
load("../output/analysis_2.RData")
```


```{r individual level models colour}
model_o <- lmer(Orange ~ Tparasites + Ai + STi + Predation + Drainage + Tparasites:Drainage + (1|Site), data = data_analysis_parasites) #north drainage and low pred almost sig
model_b <- lmer(Black ~ Tparasites + Ai + STi + Predation + Drainage + Tparasites:Drainage + (1|Site), data = data_analysis_parasites) #oropouche drainage sig

#check_model(model_o)
#check_model(model_b)

#currently presenting summary statistics in text
summary(model_o)
summary(model_b)
#car::Anova(model_o)
#car::Anova(model_b)

merged_tbl <- tbl_merge(tbls = list(tbl_regression(model_o),tbl_regression(model_b)),tab_spanner = c("**Orange**", "**Black**"))

model_o_r2 <- r2beta(model_o)
model_b_r2 <- r2beta(model_b)
plot(model_o_r2)
plot(model_b_r2)

```


individual level parasite model, no colour.
```{r individual models parasites}
model_para <- glmer(Tparasites ~ Ai + STi + Predation + Drainage + (1|Site), data = data_analysis_parasites, family = "poisson")

summary(model_para)
#car::Anova(model_para)

tbl_regression(model_para, exponentiate = TRUE)

#visualization, not in MS
data.frame(ggeffects::ggpredict(model_para, terms = c("Ai", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_point(inherit.aes = FALSE,
             data = data_analysis_parasites,
             aes(x = Ai, y = Tparasites, colour = Predation),
             alpha = 0.5) +
  xlab("MHC Allelic Richness") +
  ylab("Individual Parasite Intensity")
```


### 3. Linear models for population level

Using Micro-Ar as the independent variable and MHC-Ar and MHC-STr as dependent variables at the population level. Using a mixed model to account for drainage differences. Run these with and without Turure (as it is an introduction site), however conclusions are the same across both so will report with complete dataset in manuscript. 

```{r pop lmms data include = FALSE}
load("../output/analysis_3.RData")
```


```{r microsat mixed model no turure, eval=FALSE}
model_micro_ar <- lmer(`MHC-Ar` ~ `Micro-Ar` + Predation + (1|Drainage), data = PopDF %>% filter(!Drainage == "Oropuche (Caroni type)"))

#check_model(model_micro_ar)
summary(model_micro_ar)

model_micro_str <- lmer(`MHC-STr` ~ `Micro-Ar` + Predation + (1|Drainage), data = PopDF %>% filter(!Drainage == "Oropuche (Caroni type)"))

#check_model(model_micro_str)
summary(model_micro_str)
```


```{r microsat mixed model}
#Predation regime may contribute variance but only has two levels and will be added as a fixed factor

model_micro_all_ar <- lmer(`MHC-Ar` ~ `Micro-Ar` + Predation + (1|Drainage), data = pop_level_dat)

#check_model(model_micro_all_ar)
summary(model_micro_all_ar)

model_micro_all_str <- lmer(`MHC-STr` ~ `Micro-Ar` + Predation + (1|Drainage), data = pop_level_dat)

#check_model(model_micro_all_str)
summary(model_micro_all_str)

merged_tbl2 <- tbl_merge(tbls = list(tbl_regression(model_micro_all_str),tbl_regression(model_micro_all_ar)),
                        tab_spanner = c("**ST Richness**", "**Allelic Richness**"))

#effect size, plots are in plots section
model_micro_all_ar_r2 <- r2beta(model_micro_all_ar) #marginal r2
model_micro_all_str_r2 <- r2beta(model_micro_all_str) #marginal r2
```

### Save outputs for figures

```{r figure outputs}
save(model_b, model_o, data_analysis_parasites, file = "../output/fig2.RData")

save(PopDF_parasites, file = "../output/fig3.Rdata")

save(model_micro_all_str_r2, model_micro_all_ar_r2, model_micro_all_ar, model_micro_all_str, file = "../output/microsat_figs.RData")

save(coin_2, coin_9, coin_10, file = "../output/coia_supp_figs.RData")
```

