---
title: "02-data-analysis"
author: "Allegra Love"
date: "2025-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(gtsummary)
library(ade4)
library(corrr)
library(lme4)
library(r2glmm)
library(performance)
library(flextable)
library(officer)
library(glmmTMB)
library(lmerTest)
library(gtsummary)
library(gt)
library(purrr)

#read in cleaned data
#analysis 1
load(file = "../output/analysis_1.RData")
#analysis 2
load(file = "../output/analysis_2.RData")
#analysis 3
load(file = "../output/analysis_3.RData")

pop_level_dat <- pop_level_dat %>% 
  mutate(gyro_present = ifelse(gyro_prevalence > 0.1, "Yes", "No"))
```





### 1. Co-inertia analyses

**COIA(1)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-ST frequencies for each population

**COIA(2)**

    Data matrix (1) contained colour means and Gyrodactylus prevalence for each population
    Data matrix (2) contained MHC-Ar, and MHC-STr

Permutation test with 1000 bootstrap replications to test significance of co-variation in COIA

To address variation from drainage I added a within-class analysis at the drainage level.

#### At the individual level...

```{r coia individual level}
# pcas, currently keeping all factors
pca_colour <- dudi.pca(colour_matrix, scale = TRUE, scannf = FALSE, nf = 3) # because parasite counts and relative area have different scales
pca_mhc_f <- dudi.pca(mhc_f_matrix, scale = FALSE, scannf = FALSE, nf = 10) #same scale
pca_mhc_r <- dudi.pca(mhc_r_matrix, scale = FALSE, scannf = FALSE, nf = 2) #same scale

# for both COIN analyses the screeplot indicates only the first factor is significantly contributing, should we only keep 1?

#COIN 1 - colour and MHC ST frequency
coin_1 <- coinertia(dudiY = pca_colour, dudiX = pca_mhc_f, scannf = FALSE) #manually put axes to keep

#COIN 2 - colour and allelic richness
coin_2 <- coinertia(dudiY = pca_colour, dudiX = pca_mhc_r, scannf = FALSE) #manually put axes to keep

#permutation test
rv1 <- RV.rtest(pca_colour$tab, pca_mhc_f$tab, 1000)
plot(rv1) # simulated p = 0.001, RV = 0.11

rv2 <- RV.rtest(pca_colour$tab, pca_mhc_r$tab, 1000)
plot(rv2) # simulated p = 0.374, RV = 0.014

plot(coin_1)
plot(coin_2)
```


"Within-class analyses take into account a partition of individuals into groups and focus on structures which are common to all groups." Dray & Dufour, 2007

Analyze this data and account for drainage differences, covariation is no longer significant. Correlations between colour and parasite load are likely explained by drainage level processes
```{r within group}
#doing a within group analysis for drainage level, this only looks at common processes across groups?

wca_colour <- wca(pca_colour, fac = data_matrices$Drainage, scannf = FALSE)
wca_mhc_f <- wca(pca_mhc_f, fac = data_matrices$Drainage, scannf = FALSE) 

plot(wca_mhc_f)
plot(wca_colour)

coin_3 <- coinertia(dudiY = wca_colour, dudiX = wca_mhc_f, scannf = FALSE)
plot(coin_3)

#permutation test
rv3 <- RV.rtest(wca_colour$tab, wca_mhc_f$tab, 1000)
plot(rv3) # p-value = 0.26, rv = 0.035

#with site? Same interpretation
wca_colour_2 <- wca(pca_colour, fac = data_matrices$Site, scannf = FALSE)
wca_mhc_f_2 <- wca(pca_mhc_f, fac = data_matrices$Site, scannf = FALSE) 

plot(wca_mhc_f_2)
plot(wca_colour_2)

coin_4 <- coinertia(dudiY = wca_colour_2, dudiX = wca_mhc_f_2, scannf = FALSE)
plot(coin_4) #saved as coia_plot_2.png, 590 x 640

#permutation test
rv4 <- RV.rtest(wca_colour_2$tab, wca_mhc_f_2$tab, 1000)
plot(rv4) # rv = 0.023, p-value = 0.69


#for the other coia (2)
wca_mhc_r <- wca(pca_mhc_r, fac = data_matrices$Drainage, scannf = FALSE) 

plot(wca_mhc_r)
plot(wca_colour)

coin_6 <- coinertia(dudiY = wca_colour, dudiX = wca_mhc_r, scannf = FALSE)
plot(coin_6) #saved as coia_plot_2.png, 590 x 640

#permutation test
rv6 <- RV.rtest(wca_colour$tab, wca_mhc_r$tab, 1000)

plot(rv6) # rv = 0.61, p-value = 0.66
```

#### At the population level...

Confirmed that the same trend exists at the population level (prevalence), significant relationship when not using the within class analysis
```{r coia population level}
# pcas, currently keeping all factors
pca_colour_p <- dudi.pca(colour_matrix_pop, scale = TRUE, scannf = FALSE) # because parasite counts and relative area have different scales
pca_mhc_f_p <- dudi.pca(mhc_f_matrix_pop, scale = FALSE, scannf = FALSE)
pca_mhc_r_p <- dudi.pca(mhc_r_matrix_pop, scale = FALSE, scannf = FALSE) 

#COIN 1 - colour and MHC ST frequency
coin_7 <- coinertia(dudiY = pca_colour_p, dudiX = pca_mhc_f_p, scannf = FALSE)

#COIN 2 - colour and allelic richness
coin_8 <- coinertia(dudiY = pca_colour_p, dudiX = pca_mhc_r_p, scannf = FALSE) 

#permutation test
rv7 <- RV.rtest(pca_colour_p$tab, pca_mhc_f_p$tab, 1000)
plot(rv7) # simulated p = 0.05, RV = 0.386

rv8 <- RV.rtest(pca_colour_p$tab, pca_mhc_r_p$tab, 1000)
plot(rv8) # simulated p = 0.002, RV = 0.564


## WITHIN CLASS
wca_colour_p <- wca(pca_colour_p, fac = PopDF_parasites$Drainage_simple, scannf = FALSE)
wca_mhc_f_p <- wca(pca_mhc_f_p, fac = PopDF_parasites$Drainage_simple, scannf = FALSE) 
wca_mhc_r_p <- wca(pca_mhc_r_p, fac = PopDF_parasites$Drainage_simple, scannf = FALSE) 

plot(wca_mhc_f_p)
plot(wca_mhc_r_p)
plot(wca_colour_p)

coin_9 <- coinertia(dudiY = wca_colour_p, dudiX = wca_mhc_f_p, scannf = FALSE)
plot(coin_9)

coin_10 <- coinertia(dudiY = wca_colour_p, dudiX = wca_mhc_r_p, scannf = FALSE)
plot(coin_10)

#permutation test
rv9 <- RV.rtest(wca_colour_p$tab, wca_mhc_f_p$tab, 1000)
plot(rv9) # p-value = 0.24, rv = 0.26
rv10 <- RV.rtest(wca_colour_p$tab, wca_mhc_r_p$tab, 1000)
plot(rv10) # p-value = 0.065, rv = 0.30
```

#### Pearson's Correlations on key interactions at population level
```{r PCC}
#correlation plot
PopDF_parasites %>% 
  correlate() %>% 
  focus(prevalence)


cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST1, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST2, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST3, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST4, method = "pearson") 
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST5, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST6, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST7, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST8, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST9, method = "pearson")
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST10, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$ST11, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$mean_o, method = "pearson") # strong but insig p = 0.09

PopDF_parasites %>% 
  correlate() %>% 
  focus(mean_o)

cor.test(PopDF_parasites$mean_o, PopDF_parasites$MHC_STr, method = "pearson") #sig
cor.test(PopDF_parasites$mean_o, PopDF_parasites$MHC_Ar, method = "pearson") #sig

cor.test(PopDF_parasites$prevalence, PopDF_parasites$MHC_STr, method = "pearson") #sig
cor.test(PopDF_parasites$prevalence, PopDF_parasites$MHC_Ar, method = "pearson") #sig

#of the ST none correlated with orange
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST1, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST2, method = "pearson") 
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST3, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST4, method = "pearson") 
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST5, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST6, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST7, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST8, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST9, method = "pearson")
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST10, method = "pearson") 
cor.test(PopDF_parasites$mean_o, PopDF_parasites$ST11, method = "pearson") 

```


### 2. individual level mixed models

Microsatellite models across entire population
```{r microsat indiv models orange}
microsat_o_model_para <- glmmTMB(Orange_squeezed ~ Tparasites + Ai + STi + microsat_rich + length_sc + Predation + Drainage + (1|Site), data = data_analysis, family = beta_family())
microsat_o_model_inf <- glmmTMB(Orange_squeezed ~ Infected + Ai + STi + microsat_rich + length_sc + Predation + Drainage + (1|Site), data = data_analysis, family = beta_family())
microsat_o_model_null<- glmmTMB(Orange_squeezed ~  Ai + STi + microsat_rich + length_sc + Predation + Drainage + (1|Site), data = data_analysis, family = beta_family())

check_model(microsat_o_model_para)
check_model(microsat_o_model_inf)
check_model(microsat_o_model_null)

summary(microsat_o_model_para)
summary(microsat_o_model_inf)
summary(microsat_o_model_null)

compare_performance(microsat_o_model_para, microsat_o_model_inf, microsat_o_model_null, rank = T)
AIC(microsat_o_model_para, microsat_o_model_inf, microsat_o_model_null)
```


```{r microsat indiv models black}
microsat_b_model_para <- lmer(Black ~ Tparasites + Ai + STi + microsat_rich + length_sc +  Predation + Drainage + (1|Site), data = data_analysis)
microsat_b_model_inf <- lmer(Black ~ Infected + Ai + STi + microsat_rich + length_sc + Predation + Drainage + (1|Site), data = data_analysis)
microsat_b_model_null<- lmer(Black ~  Ai + STi + microsat_rich + length_sc + Predation + Drainage + (1|Site), data = data_analysis)

check_model(microsat_b_model_para)
check_model(microsat_b_model_inf)
check_model(microsat_b_model_null)

summary(microsat_b_model_para)
summary(microsat_b_model_inf)
summary(microsat_b_model_null)

compare_performance(microsat_b_model_para,  microsat_b_model_inf, microsat_b_model_null, rank = T)
AIC(microsat_b_model_para,  microsat_b_model_inf, microsat_b_model_null)
```


```{r tables for individual all pop colour}

colour_ind_all_pop_models <- list("Black_All_Pop" = microsat_b_model_null, "Orange_all_pop" = microsat_o_model_inf)

iwalk(colour_ind_all_pop_models, function(model, name) {
  tbl_regression(
    model, 
    component = "cond", 
    estimate_fun = function(x) style_sigfig(x, digits = 3), 
    pvalue_fun = function(x) style_pvalue(x, digits = 3), 
    label = list(
      Ai ~ "MHC Allelic Richness",
      STi ~ "MHC Supertype Richness", 
      microsat_rich ~ "Microsatellite Richness",
      length_sc ~ "Standard length (scaled)",
      Predation ~ "Predation",
      Drainage ~ "Drainage",
      Infected = "Infection Status"
    )) %>% 
    modify_table_styling(
      columns = everything(), 
      rows = p.value < 0.05, 
      text_format = "bold"
    ) %>% 
    as_gt() %>% 
    gtsave(filename = paste0("../output/Table_S7_", name, ".docx"))})
```

Linear mixed models that control for both predation and drainage

is gaussian distribution appropriate for colour models? Black yes, orange no.
```{r distribution of colour}
# black is normally distributed
ggplot(data_analysis_parasites, aes(Black)) + 
  geom_histogram(colour = "grey25", binwidth = .03) + 
  geom_density(colour = "coral") +
  labs( x = "Relative Area Black", y = "Count")

#left skew for orange, models will use a beta-binomial distribution
ggplot(data_analysis_parasites, aes(Orange)) + 
  geom_histogram(fill = "coral", colour = "grey25", binwidth = .03) + 
  geom_density(colour = "grey25") +
  labs( x = "Relative Area Orange", y = "Count")
```

we have one 0 value in the orange area data which cannot be tolerated by a betabinomial distribution, use a squeeze transformation (completed in 01-data-prep.rmd) rather than omitting that singular data point (see Smithson & Verkuilen, 2007).

```{r individual level models colour}
model_o <- glmmTMB(Orange_squeezed ~ Infected + Tparasites + Ai + STi + microsat_rich + length_sc + Predation + Drainage + (1|Site), data = data_analysis_parasites, family = beta_family())
#infection, drainage, and predation sig
model_b <- glmmTMB(Black ~ Infected + Tparasites + Ai + STi + microsat_rich + length_sc + Predation + Drainage + (1|Site), data = data_analysis_parasites, family = gaussian()) #oropouche drainage and length sig

check_model(model_o)
check_model(model_b)

summary(model_o)
summary(model_b)
```

Move to figures.rmd?

```{r Table 1}

colour_models <- list( "Orange Model" = model_o, "Black Model" = model_b)

colour_table <- imap(colour_models, function(model, name) {
  tbl_regression(
    model,
    component = "cond", 
    estimate_fun = function(x) style_sigfig(x, digits = 3), 
    pvalue_fun = function(x) style_pvalue(x, digits = 3),
    label = list(
      Infected   ~ "Infected Status",
      Tparasites ~ "Gyrodactylus Abundance",
      Ai         ~ "MHC-Ar",
      STi        ~ "MHC-STr",
      microsat_rich ~ "Microsatellite Richness",
      length_sc  ~ "Standard Length (Scaled)",
      Predation  ~ "Predation",
      Drainage   ~ "Drainage"
    )
  ) %>%
    modify_column_hide(columns = groupname_col) %>% 
    modify_table_styling(
      columns = everything(), 
      rows = p.value < 0.05,  
      text_format = "bold"
    )
})

tbl_merge(tbls = colour_table, tab_spanner = names(colour_table)) %>% 
  as_gt() %>%
  gtsave(filename = "../output/Table_1_colour.docx")
```


individual level parasite model, no colour.

Parasite load model with orthogonal polynomial term to assess the relevance of the MHC quadratic term.
```{r quad model}
#high collinearity between polynomial terms for Ai and STI so used model selection to proceed with the better model
model_both_quad <- glmer(Tparasites ~ poly(Ai, 2, raw = FALSE) + poly(STi, 2, raw = FALSE) + microsat_rich + length_sc + Predation + Drainage + (1|Site), data = data_analysis_parasites, family = "poisson", control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5)))

model_ai_quad <- glmer(Tparasites ~ poly(Ai, 2, raw = FALSE) + STi + microsat_rich + length_sc + Predation + Drainage + (1|Site), 
                       data = data_analysis_parasites, family = "poisson", control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5))) #convergence warning

model_sti_quad <- glmer(Tparasites ~ Ai + poly(STi, 2, raw = FALSE) + microsat_rich + length_sc + Predation + Drainage + (1|Site), 
                       data = data_analysis_parasites, family = "poisson", control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5))) 

# Compare AIC scores
compare_performance(model_both_quad, model_ai_quad, model_sti_quad, rank = TRUE) # model_sti_quad has lowest AIC
AIC(model_both_quad, model_ai_quad, model_sti_quad) # AIC within 2 units but STi quad model is lower

check_model(model_sti_quad)
summary(model_sti_quad)
#car::Anova(model_sti_quad)
```


```{r quad model visualization}
tbl_regression(model_sti_quad, exponentiate = TRUE)

data.frame(ggeffects::predict_response(model_sti_quad, terms = c("STi [all]", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2) +
  geom_line(aes(colour = group), linewidth = 1) +
  geom_point(data = data_analysis_parasites,
             aes(x = STi, y = Tparasites, colour = Predation),
             inherit.aes = FALSE,
             alpha = 0.4, 
             position = position_jitter(width = 0.1, height = 0.2)) +
  labs(
    x = "STi", # Or change to your actual variable name
    y = "Individual Parasite Load",
    fill = "Predation",
    colour = "Predation"
  )
```

### 2.2 individual models binary infection variable

individual level parasite (infected binary) model, no colour quadratic term.
Singular fit warnings as random factor Site is not explaining any variation (explained by drainage?)
```{r quad binary parasite model}
model_both_quad_bin <- glmer(Infected ~ poly(Ai, 2, raw = FALSE) + poly(STi, 2, raw = FALSE) + microsat_rich + length_sc + Predation + Drainage + (1|Site), data = data_analysis_parasites, family = "binomial") #high VIF may be a bug in performance package from using polynomials

model_ai_quad_bin <- glmer(Infected ~ poly(Ai, 2, raw = FALSE) + STi + microsat_rich + length_sc + Predation + Drainage + (1|Site), 
                       data = data_analysis_parasites, family = "binomial")

model_sti_quad_bin <- glmer(Infected ~ Ai + poly(STi, 2, raw = FALSE) + microsat_rich + length_sc + Predation + Drainage + (1|Site), 
                       data = data_analysis_parasites, family = "binomial") 

# Compare AIC scores
compare_performance(model_both_quad_bin, model_ai_quad_bin, model_sti_quad_bin, rank = TRUE)
AIC(model_both_quad_bin, model_ai_quad_bin, model_sti_quad_bin) # model_sti_quad has similar AIC so use this to parallel intensity model

check_model(model_sti_quad_bin)
summary(model_sti_quad_bin) # singular fit, Site not explaining any variation, nothing sig
```


```{r binary quad model visualization}
tbl_regression(model_sti_quad_bin, exponentiate = TRUE)

data.frame(ggeffects::predict_response(model_sti_quad_bin, terms = c("STi [all]", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2) +
  geom_line(aes(colour = group), linewidth = 1) +
  geom_point(data = data_analysis_parasites,
             aes(x = STi, y = InfectedBinary, colour = Predation),
             inherit.aes = FALSE,
             alpha = 0.4, 
             position = position_jitter(width = 0.1, height = 0.2)) +
  labs(
    x = "STi", # Or change to your actual variable name
    y = "Individual Infection incidence",
    fill = "Predation",
    colour = "Predation"
  )
```


```{r Table 2}
parasite_models <- list( "Abundance Model" = model_sti_quad, "Infection Status Model" = model_sti_quad_bin)

parasite_table <- imap(parasite_models, function(model, name) {
  tbl_regression(
    model,
    component = "cond", 
    estimate_fun = function(x) style_sigfig(x, digits = 3), 
    pvalue_fun = function(x) style_pvalue(x, digits = 3),
    label = list(
      Ai         ~ "MHC-Ar",
      STi  ~ "MHC-STr (poly)",
      microsat_rich ~ "Microsatellite richness",
      length_sc  ~ "Standard Length",
      Predation  ~ "Predation",
      Drainage   ~ "Drainage"
    )
  ) %>%
    #modify_column_hide(columns = groupname_col) %>% 
    modify_table_styling(
      columns = everything(), 
      rows = p.value < 0.05,  
      text_format = "bold"    
    )
})

tbl_merge(tbls = parasite_table, tab_spanner = names(parasite_table)) %>% 
  as_gt() %>%
  gtsave(filename = "../output/Table_2_parasite.docx")
```


### 3. Linear models for population level

Using Micro-Ar as the independent variable and MHC-Ar and MHC-STr as dependent variables at the population level. Using a mixed model to account for drainage differences. Run these with and without Turure (as it is an introduction site), however conclusions are the same across both so will report with complete dataset in manuscript. 

```{r microsat pop mixed model}
model_micro_all_ar <- lmer(mhc_Ai ~ micro_Ai + (1|Predation) + (1|Drainage_simple), data = pop_level_dat)
r2(model_micro_all_ar)

#check_model(model_micro_all_ar)
summary(model_micro_all_ar)

model_micro_all_str <- lmer(mhc_STi ~ micro_Ai + (1|Predation) + (1|Drainage_simple), data = pop_level_dat)
r2(model_micro_all_str)
#check_model(model_micro_all_str)
summary(model_micro_all_str)
```


```{r microsat mixed model individual level}
# out of curiosity at the individual level now that we have that data
model_micro_ar_ind <- glmmTMB(Ai ~ microsat_rich + (1|Predation) + (1|Drainage/Site), data = data_analysis, family = Gamma(link = "log"))
r2(model_micro_ar_ind)
summary(model_micro_ar_ind)

model_micro_str_ind <- glmmTMB(STi ~ microsat_rich + (1|Predation) + (1|Drainage/Site), data = data_analysis, family = Gamma(link = "log"))
r2(model_micro_str_ind)
summary(model_micro_str_ind)
```


```{r Table S5}
microsat_models <- list( "MHC Allelic Richness" = model_micro_all_ar, "MHC Supertype Richness" = model_micro_all_str)

microsat_table <- imap(microsat_models, function(model, name) {
  tbl_regression(
    model,
    component = "cond", 
    estimate_fun = function(x) style_sigfig(x, digits = 3), 
    pvalue_fun = function(x) style_pvalue(x, digits = 3),
    label = list(micro_Ai        ~ "Microsatellite Richness")) %>%
    #modify_column_hide(columns = groupname_col) %>% 
    modify_table_styling(
      columns = everything(), 
      rows = p.value < 0.05,  
      text_format = "bold"    
    )
})

tbl_merge(tbls = microsat_table, tab_spanner = names(microsat_table)) %>% 
  as_gt() %>%
  gtsave(filename = "../output/Table_S6_microsat.docx")

microsat_models_ind <- list("MHC Allelic Richness individual" = model_micro_ar_ind, "MHC Supertype Richness individual" = model_micro_str_ind)

microsat_table_ind <- imap(microsat_models_ind, function(model, name) {
  tbl_regression(
    model,
    component = "cond", 
    estimate_fun = function(x) style_sigfig(x, digits = 3), 
    pvalue_fun = function(x) style_pvalue(x, digits = 3),
    label = list(microsat_rich        ~ "Microsatellite Richness")) %>%
    #modify_column_hide(columns = groupname_col) %>% 
    modify_table_styling(
      columns = everything(), 
      rows = p.value < 0.05,  
      text_format = "bold"    
    )
})

tbl_merge(tbls = microsat_table_ind, tab_spanner = names(microsat_table_ind)) %>% 
  as_gt() %>%
  gtsave(filename = "../output/Table_S7_microsat_ind.docx") #merged manually with S6 so there's one table


```

Null model comparison

Use microsatellite information in addition to MHC to see if that explains trends in colour, use a simple AIC comparison with MHC models to show MHC information improves predictive power of model
Drainage removed due to high collinearity
```{r null model pop orange}
model_o_null_para <- glmmTMB(mean_o ~ Predation + gyro_prevalence + mhc_Ai + mhc_STi + micro_Ai, dispformula = ~ 1, data = pop_level_dat, family = beta_family())
model_o_null_mhc<- glmmTMB(mean_o ~ Predation + mhc_Ai + mhc_STi + micro_Ai, dispformula = ~ 1, data = pop_level_dat, family = beta_family())

check_model(model_o_null_para)
check_model(model_o_null_mhc)

summary(model_o_null_para)
summary(model_o_null_mhc)

compare_performance(model_o_null_para, model_o_null_mhc, rank = T)
AIC(model_o_null_para, model_o_null_mhc)
```


```{r null model pop black}
model_b_null_para <- lm(mean_b ~ Predation + gyro_prevalence + mhc_Ai + mhc_STi + micro_Ai , data = pop_level_dat)
model_b_null_mhc <- lm(mean_b ~ Predation + mhc_Ai + mhc_STi + micro_Ai , data = pop_level_dat)

check_model(model_b_null_para)
check_model(model_b_null_mhc)

summary(model_b_null_para)
summary(model_b_null_mhc)
#car::Anova(model_b_null)
compare_performance(model_b_null_para, model_b_null_mhc, rank = T)
AIC(model_b_null_para, model_b_null_mhc)
```

Export of table for all population model comparison with and without parasites for both colours.
```{r colour all pop tables}

colour_all_pop_models <- list("Orange Model" = model_o_null_para, "Black Model" = model_b_null_para)

colour_all_pop_table <- imap(colour_all_pop_models, function(model, name) {
  tbl_regression(
    model,
    component = "cond", 
    estimate_fun = function(x) style_sigfig(x, digits = 3), 
    pvalue_fun = function(x) style_pvalue(x, digits = 3),
    label = list(Predation ~ "Predation",
                 gyro_prevalence ~ "Gyrodactylus Prevalence",
                 mhc_Ai ~ "MHC Allelic Richness (pop)",
                 mhc_STi ~ "MHC Supertype Richness (pop)",
                 micro_Ai        ~ "Microsatellite Richness (pop)")) %>%
    #modify_column_hide(columns = groupname_col) %>% 
    modify_table_styling(
      columns = everything(), 
      rows = p.value < 0.05,  
      text_format = "bold"    
    )
})

tbl_merge(tbls = colour_all_pop_table, tab_spanner = names(colour_all_pop_table)) %>% 
  as_gt() %>%
  gtsave(filename = "../output/colour_allpop_table.docx") 

```


Comparison of models with neutral genetic diversity (microsatellites) and MHC diversity (allelic and supertype richness), This is completed at the population level only as we have only population averages for microsatellite diversity.
```{r null with parasites}
model_para_null <- lm(prevalence ~ Predation + micro_Ai , data = PopDF_parasites) 
model_para_null2 <- lm(prevalence ~ Predation + mhc_Ai + mhc_STi + micro_Ai , data = PopDF_parasites) 
model_para_null3 <- lm(prevalence ~ Predation + mhc_Ai + mhc_STi, data = PopDF_parasites) 

# model_para_null <- lm(gyro_prevalence ~ Predation + Micro_Ar , data = pop_level_dat) 
# model_para_null2 <- lm(gyro_prevalence ~ Predation + `MHC-Ar` + `MHC-STr` + Micro_Ar , data = pop_level_dat) 
# model_para_null3 <- lm(gyro_prevalence ~ Predation + `MHC-Ar` + `MHC-STr` , data = pop_level_dat) 

summary(model_para_null)
summary(model_para_null2)
summary(model_para_null3)

compare_performance(model_para_null, model_para_null2, model_para_null3, rank = T) # model without microsat is best
```


#### Delta AIC table for supplement (all model comparisons)

split into a pop level and individual level table? or an all pops and parasite only pops table? Update with changes needed though
```{r aic comparisons}
library(AICcmodavg)

prevalence_analysis <- list(model_para_null, model_para_null2, model_para_null3)
prevalence_names <- c("Prevalence ~ Predation + Micro-Ar", "Prevalence ~ Predation + MHC-Ar + MHC-STr + Micro-Ar", "Prevalence ~ Predation + MHC-Ar + MHC-STr")

aic_dat <- aictab(cand.set = prevalence_analysis, modnames = prevalence_names, second.ord = FALSE) %>% 
  mutate(analysis = "Prevalence, MHC and microsatellite diversity", R2_conditional = NA, R2_marginal = NA) %>% 
  cbind(compare_performance(prevalence_analysis, rank = T) %>%  select(R2))

#mixed effect models will have conditional and marginal r2 reported
orange_analysis <- list(model_o_null_para, model_o_null_mhc)
orange_names <- c("Orange ~ Predation + Prevalence + MHC-Ar + MHC-STr + Micro-Ar", "Orange ~ Predation + MHC-Ar + MHC-STr + Micro-Ar")


orange <- aictab(cand.set = orange_analysis, modnames = orange_names, second.ord = FALSE) %>% 
  mutate(analysis = "Impact of parasitism on orange", R2_conditional = NA, R2_marginal = NA) %>% 
  cbind(compare_performance(orange_analysis, rank = T) %>%  select(R2))

black_analysis <- list(model_b_null_para, model_b_null_mhc)
black_names <- c("Black ~ Predation + Prevalence + MHC-Ar + MHC-STr + Micro-Ar", "Black ~ Predation + MHC-Ar + MHC-STr + Micro-Ar")

black <- aictab(cand.set = black_analysis, modnames = black_names, second.ord = FALSE) %>% 
  mutate(analysis = "Impact of parasitism on black", R2_conditional = NA, R2_marginal = NA) %>% 
  cbind(compare_performance(black_analysis, rank = T) %>%  select(R2))

#indiviudal level models
orange_individual_analysis <- list(microsat_o_model_null, microsat_o_model_inf, microsat_o_model_para)
orange_individual_names <- c("Orange ~ MHC-Ar + MHC-STr + Micro-Ar + length +Predation + Drainage + (1|Site)",
                  "Orange ~ Infected + MHC-Ar + MHC-STr + Micro-Ar + length +Predation + Drainage + (1|Site)",
                  "Orange ~ Abundance + MHC-Ar + MHC-STr + Micro-Ar + length +Predation + Drainage + (1|Site)")


orange_individual <- aictab(cand.set = orange_individual_analysis, modnames = orange_individual_names, second.ord = FALSE) %>%
  mutate(analysis = "Impact of parasitism on orange (individual)",R2 = NA) %>% 
  cbind(compare_performance(orange_individual_analysis, rank = T) %>%  select(R2_conditional, R2_marginal))

black_individual_analysis <- list(microsat_b_model_null, microsat_b_model_inf, microsat_b_model_para)
black_individual_names <- c("Black ~ MHC-Ar + MHC-STr + Micro-Ar + length +Predation + Drainage + (1|Site)",
                  "Black ~ Infected + MHC-Ar + MHC-STr + Micro-Ar + length +Predation + Drainage + (1|Site)",
                  "Black ~ Abundance + MHC-Ar + MHC-STr + Micro-Ar + length +Predation + Drainage + (1|Site)")

black_individual <- aictab(cand.set = black_individual_analysis, modnames = black_individual_names, second.ord = FALSE) %>% 
  mutate(analysis = "Impact of parasitism on black (individual)", R2 = NA) %>% 
  cbind(compare_performance(black_individual_analysis, rank = T) %>%  select(R2_conditional, R2_marginal)) %>% 
  rename(LL = Res.LL)

quad_analysis <- list(model_sti_quad, model_ai_quad, model_both_quad)
quad_names <- c("Abundance ~ poly(STi, 2, raw = FALSE) + Length + Predation + Drainage + (1|Site)", "Abundance ~ poly(MHC-Ar, 2, raw = FALSE) + Length + Predation + Drainage + (1|Site)", "Abundance ~ poly(MHC-Ar, 2, raw = FALSE) + poly(STi, 2, raw = FALSE) + Length + Predation + Drainage + (1|Site)")

quad <- aictab(cand.set = quad_analysis, modnames = quad_names, second.ord = FALSE) %>% 
  mutate(analysis = "Polynomial selection for abundance", R2 = NA) %>% 
  cbind(compare_performance(quad_analysis, rank = T) %>%  select(R2_conditional, R2_marginal))

quad_analysis_bin <- list(model_sti_quad_bin, model_ai_quad_bin, model_both_quad_bin)
quad_names_bin <- c("Infected ~ poly(STi, 2, raw = FALSE) + Length + Predation + Drainage + (1|Site)", "Infected ~ poly(MHC-Ar, 2, raw = FALSE) + Length + Predation + Drainage + (1|Site)", "Infected ~ poly(MHC-Ar, 2, raw = FALSE) + poly(STi, 2, raw = FALSE) + Length + Predation + Drainage + (1|Site)")

quad_bin <- aictab(cand.set = quad_analysis_bin, modnames = quad_names_bin, second.ord = FALSE) %>% 
  mutate(analysis = "Polynomial selection for infection status", R2 = NA, R2_conditional = NA) %>% 
  cbind(compare_performance(quad_analysis_bin, rank = T) %>%  select(R2_marginal))

aic_dat_all<- rbind(aic_dat, orange, black, orange_individual, black_individual, quad, quad_bin) %>% 
  select(analysis, Modnames, K, Delta_AIC, AICWt, R2, R2_conditional, R2_marginal) %>% 
  rename("Analysis" = analysis, "Model Name" = Modnames, "Delta AIC" = Delta_AIC, "AIC Weight" = AICWt) %>% 
  mutate(across(where(is.numeric), ~signif(., 2))) %>% 
  mutate(across(where(is.numeric), ~if_else(. == 0, "0", sprintf("%.2f", .)))) %>% 
  as.data.frame() %>% 
  gt() %>%
  gtsave(filename = "../output/Table_S4_aic.docx")
```


### Save outputs for figures

```{r figure outputs}
save(model_b, model_o, data_analysis_parasites, file = "../output/fig2.RData")

save(PopDF_parasites, pop_level_dat, file = "../output/fig3.Rdata")

save(model_micro_all_str_r2, model_micro_all_ar_r2, model_micro_all_ar, model_micro_all_str, file = "../output/microsat_figs.RData")

save(model_o_null_para, model_b_null_para, pop_level_dat, file = "../output/supp_fig3.RData")

save(model_sti_quad, data_analysis_parasites, file = "../output/fig4.RData")

save(coin_2, coin_9, coin_10, file = "../output/coia_supp_figs.RData")
```


