---
title: "03-figures"
author: "Allegra Love"
date: "2025-09-27"
output: html_document
---


```{r setup include = FALSE}
library(car)
library(ggplot2)
library(gridExtra)
library(grid)
library(tidyverse)
library(ggeffects)
```


### Figures

Note: drainage specific hex values for figures
Fill (map) (too light for figure): 
Northern - #b9c1ef
Oropouche - #f0e8bb
Caroni - #eecdcd

Fill (figures):
Northern - #8292ed
Oropouche - #ebdb81
Caroni - #ed8a8a

Lines and points (darker):
Northern - #425ced
Oropouche - #edd442
Caroni - #ed3e3e

Figure 1 was generated in QGIS

### Figure 2: Drainage specific relationship between individual parasite load and relative area black and orange.
```{r fig 2 data include = FALSE}
load("../output/fig2.RData")
#function to sahre a legend between a multi-plot arrangement (from: https://cran.r-project.org/web/packages/egg/vignettes/Ecosystem.html)
grid_arrange_shared_legend <-
  function(...,
           ncol = length(list(...)),
           nrow = 1,
           position = c("bottom", "right")) {
    
    plots <- list(...)
    position <- match.arg(position)
    g <-
      ggplotGrob(plots[[1]] + theme(legend.position = position))$grobs
    legend <- g[[which(sapply(g, function(x)
      x$name) == "guide-box")]]
    lheight <- sum(legend$height)
    lwidth <- sum(legend$width)
    gl <- lapply(plots, function(x)
      x + theme(legend.position = "none"))
    gl <- c(gl, ncol = ncol, nrow = nrow)
    
    combined <- switch(
      position,
      "bottom" = arrangeGrob(
        do.call(arrangeGrob, gl),
        legend,
        ncol = 1,
        heights = unit.c(unit(1, "npc") - lheight, lheight)
      ),
      "right" = arrangeGrob(
        do.call(arrangeGrob, gl),
        legend,
        ncol = 2,
        widths = unit.c(unit(1, "npc") - lwidth, lwidth)
      )
    )
    
    grid.newpage()
    grid.draw(combined)
    
    # return gtable invisibly
    invisible(combined)
    
  }
```


```{r Fig 2}
#ggplot black
bplot <- data.frame(ggeffects::ggpredict(model_b, terms = c("Tparasites [0:12]", "Drainage"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group), show.legend = F) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_jitter(inherit.aes = FALSE,
             data = data_analysis_parasites,
             aes(x = Tparasites, y = Black, colour = Drainage, shape = Predation),
             width = 0.1, height = 0.0,
             alpha = 0.5) +
  scale_color_manual(values = c("#ed3e3e","#425ced","#edd442")) +
  scale_fill_manual(values = c("#ed8a8a","#8292ed","#ebdb81")) +
  scale_x_continuous(breaks = seq(0, 12, by = 1)) +
  xlab("Individual Parasite Load") +
  ylab("Relative Area Black") +
  labs(colour = "Drainage", fill = "Drainage") +
  theme_classic() +
  theme(legend.position = "none")

#ggplot orange
oplot <- data.frame(ggeffects::ggpredict(model_o, terms = c("Tparasites [0:12]", "Drainage"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_jitter(inherit.aes = FALSE,
             data = data_analysis_parasites,
             aes(x = Tparasites, y = Orange_squeezed, colour = Drainage, shape = Predation),
             width = 0.1, height = 0.0,
             alpha = 0.5) +
  scale_color_manual(values = c("#ed3e3e","#425ced","#edd442")) +
  scale_fill_manual(values = c("#ed8a8a","#8292ed","#ebdb81")) +
  scale_x_continuous(breaks = seq(0, 12, by = 1)) +
  xlab("Individual Parasite Load") +
  ylab("Relative Area Orange") +
  labs(colour = "Drainage", fill = "Drainage") +
  theme_classic() +
  theme(legend.position = "bottom")


png(filename = "../Figures/ColourParasite 2026-09-11.png", width = 8, height = 4, units = "in", bg = "white", res = 320)

grid_arrange_shared_legend(oplot, bplot)
dev.off()

```


### Figure 3 Two panels, MicroAR on Y axis, one plot with MHC Ar, and one with MHC STr. different symbols for predation.
```{r fig 3 data include = FALSE}
load("../output/fig3.Rdata")
```


```{r Fig 3}
#use output from models in text
micro_ar_STr <- ggpredict(model_micro_all_str, terms = "micro_Ai")
micro_ar_Ar <- ggpredict(model_micro_all_ar, terms = "micro_Ai")
#triangles for low predation (17), coloured by drainage similar to map
drainage_colours <- c(  "Caroni" = "#ed3e3e", 
  "North Slope" = "#425ced",
  "Oropuche" = "#edd442",
  "Oropuche (Caroni Type)" = "#edd442")

png(filename =  "../Figures/fig_3_2026-09-11.png", width = 10, height = 5, units = "in", bg = "white", res = 320)
op <- par(mfrow = c(1,2), 
          mar = c(5.1,5.1,2.1,2.1))
plot(x = pop_level_dat$micro_Ai, y = pop_level_dat$mhc_Ai, 
     xlab = expression(Microsatellite ~ Richness ~ (Micro-italic(Ar[pop]))), 
     ylab = (MHC ~ Allelic ~ Richness ~ (MHC-italic(Ar[pop]))), 
     pch = ifelse(pop_level_dat$Predation == "High", 19, 17), 
     cex = 2, 
     col = drainage_colours, 
     xlim = c(6.5,9.0)
     )
#regLine(lm(PopDF_parasites$MHC_Ar ~ PopDF_parasites$Micro_Ar), col = "black")
lines(micro_ar_Ar$x, micro_ar_Ar$predicted, col = "black", lwd = 2)
legend(x = 6.5, y = 8, 
       c("High Predation", "Low Predation"), 
       pch = c(19,17), 
       bty = "n", 
       pt.bg = c("black"), pt.cex = 2, cex = 1.05)
legend(x = 6.5, y = 7.3, 
       legend = names(drainage_colours)[1:3], 
       col = drainage_colours[1:3], 
       pch = 18, 
       bty = "n", 
       pt.cex = 2, 
       cex = 1.05)
plot(x = pop_level_dat$micro_Ai, y = pop_level_dat$mhc_STi, 
     xlab = expression(Microsatellite ~ Richness ~ (Micro-italic(Ar[pop]))), 
     ylab = expression(MHC ~ Supertype ~ Richness ~ (MHC-italic(STr[pop]))), 
     pch = ifelse(pop_level_dat$Predation == "High", 19, 17), 
     cex = 2, 
     col = drainage_colours, 
     xlim = c(6.5,9.0)
     )
#regLine(lm(PopDF_parasites$MHC_STr ~ PopDF_parasites$Micro_Ar), col = "black")
lines(micro_ar_STr$x, micro_ar_STr$predicted, col = "black", lwd = 2)
par(op)
dev.off()
```

### Microsat figures

```{r microsat fig data include = FALSE}
load("../output/microsat_figs.RData")
```

Effect size visualiztion (lines 339-345) in ind. models manuscript.
```{r effect size visualization}
#allelic richness
# R2 plot
plot(model_micro_all_ar_r2)

#supertype richness
#R2 plot
plot(model_micro_all_str_r2)
#
```


Making other plots for reference. Manuscript has a summary table with both results (Merged_tbl2).
```{r microsat mixed model plots not in MS, eval=FALSE}

#model plot
data.frame(ggeffects::ggpredict(model_micro_all_ar, terms = c("Micro-Ar", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_point(inherit.aes = FALSE,
             data = PopDF,
             aes(y = MHC_Ar, x = Micro_Ar, colour = Predation),
             alpha = 0.5) +
  xlab(expression(Microsatellite ~ Richness ~ (Micro-italic(Ar[pop])))) +
  ylab(expression(MHC ~ Supertype ~ Richness ~ (Micro-italic(Ar[pop]))))



data.frame(ggeffects::ggpredict(model_micro_all_str, terms = c("Micro-Ar", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_point(inherit.aes = FALSE,
             data = PopDF,
             aes(y = MHC_STr, x = Micro_Ar, colour = Predation),
             alpha = 0.5) +
  xlab(expression(Microsatellite ~ Richness ~ (Micro-italic(Ar[pop])))) +
  ylab(expression(MHC ~ Supertype ~ Richness ~ (Micro-italic(STr[pop]))))

```

### Supplemental Figure 3

```{r Supp fig 3}
load("../output/supp_fig3.Rdata")
#use output from models in text
pred_o_STi <- ggpredict(model_o_null_para, terms = "mhc_STi")
pred_b_STi <- ggpredict(model_b_null_para, terms = "mhc_STi")
pred_o_Ai  <- ggpredict(model_o_null_para, terms = "mhc_Ai")
pred_b_Ai  <- ggpredict(model_b_null_para, terms = "mhc_Ai")

#Supplemental Figure 3
png(filename =  "../Figures/Non_Effect Colour_2026-09-1png", width = 8, height = 8, units = "in", bg = "white", res = 300)
op <- par(mfrow = c(2,2),
          mar = c(6.5, 4.1, 1.1, 2.1),
          cex.lab = 1.5)
#orange colours
plot(x = pop_level_dat$mhc_STi, y = pop_level_dat$mean_o
     , xlab = "", ylab = "Relative Orange Area"
     , pch = ifelse(PopDF_parasites$Predation == "High", 19, 17), cex = 2
     , col = drainage_colours
     )
title(xlab = expression(atop(MHC ~ Supertype ~ Richness,(MHC-italic(STr[pop])))), 
      mgp = c(5, 1, 0))
legend(x = 4.1, y = 0.17
       , c("High Predation", "Low Predation")
       , pch = c(19,17)
       , bty = "n"
       , pt.bg = c("black"), pt.cex = 2, cex = 1.05)
legend(x = 4.1, y = 0.15
       , legend = names(drainage_colours)[1:3]
       , col = drainage_colours[1:3]
       , pch = 18
       , bty = "n"
       , pt.cex = 2
       , cex = 1.05)
#regLine(lm(PopDF_parasites$mean_o ~ PopDF_parasites$MHC_STr), col = "black")
lines(pred_o_STi$x, pred_o_STi$predicted, col = "black", lwd = 2)
plot(x = pop_level_dat$mhc_STi, y = pop_level_dat$mean_b
     , xlab = "", ylab = "Relative Black Area"
     , pch = ifelse(PopDF_parasites$Predation == "High", 19, 17), cex = 2
     , col = drainage_colours
     )
title(xlab = expression(atop(MHC ~ Supertype ~ Richness, (MHC-italic(STr[pop])))), 
      mgp = c(5, 1, 0))
#regLine(lm(PopDF_parasites$mean_b ~ PopDF_parasites$MHC_STr), col = "black")
lines(pred_b_STi$x, pred_b_STi$predicted, col = "grey60", lwd = 2, lty = "dashed")

plot(x = pop_level_dat$mhc_Ai, y = pop_level_dat$mean_o
     , xlab = "", ylab = "Relative Orange Area"
     , pch = ifelse(PopDF_parasites$Predation == "High", 19, 17), cex = 2
     , col = drainage_colours
     )
title(xlab = expression(atop(MHC ~ Allelic ~ Richness,(MHC-italic(Ar[pop])))), 
      mgp = c(5, 1, 0))
#regLine(lm(PopDF_parasites$mean_o ~ PopDF_parasites$MHC_Ar), col = "black")
lines(pred_o_Ai$x, pred_o_Ai$predicted, col = "grey60", lwd = 2, lty = "dashed")

plot(x = pop_level_dat$mhc_Ai, y = pop_level_dat$mean_b
     , xlab = "", ylab = "Relative Black Area"
     , pch = ifelse(PopDF_parasites$Predation == "High", 19, 17), cex = 2
     , col = drainage_colours
     )
title(xlab = expression(atop(MHC ~ Allelic ~ Richness,(MHC-italic(Ar[pop])))), 
      mgp = c(5, 1, 0))
#regLine(lm(PopDF_parasites$mean_b ~ PopDF_parasites$MHC_Ar), col = "black")
lines(pred_b_Ai$x, pred_b_Ai$predicted, col = "grey60", lwd = 2, lty = "dashed")
par(op)
dev.off()

```

###  quadratic supertype richness figure

```{r quad model figure prep}
load("../output/fig4.Rdata")


predicted_data <- data.frame(
  STi = seq(min(data_analysis_parasites$STi, na.rm = TRUE), 
            max(data_analysis_parasites$STi, na.rm = TRUE), 
            length.out = 100),
  Ai = mean(data_analysis_parasites$Ai, na.rm = TRUE),
  microsat_rich = mean(data_analysis_parasites$microsat_rich, na.rm = TRUE),
  length_sc = mean(data_analysis_parasites$length_sc, na.rm = TRUE),
  Predation = levels(factor(data_analysis_parasites$Predation))[1],
  Drainage = levels(factor(data_analysis_parasites$Drainage))[1]
)

#add CI ribbon
mod_terms <- delete.response(terms(model_sti_quad))
original_levels <- model_sti_quad@frame[, sapply(model_sti_quad@frame, is.factor), drop = FALSE]
factor_lists    <- lapply(original_levels, levels)
X <- model.matrix(mod_terms, data = predicted_data, xlev = factor_lists)
se_link <- sqrt(diag(X %*% as.matrix(vcov(model_sti_quad)) %*% t(X)))

predicted_data$Predicted_Link <- predict(model_sti_quad, newdata = predicted_data, re.form = NA, type = "link")
predicted_data$LL <- exp(predicted_data$Predicted_Link - 1.96 * se_link)
predicted_data$UL <- exp(predicted_data$Predicted_Link + 1.96 * se_link)

# re.form = NA ignores random effects to show the population-level trend
predicted_data$Predicted <- predict(model_sti_quad, newdata = predicted_data, re.form = NA, type = "response")

# separate linear and quad terms
betas <- fixef(model_sti_quad)
linear_col_idx <- grep("poly\\(STi.*\\)1", colnames(X))
quad_col_idx   <- grep("poly\\(STi.*\\)2", colnames(X))
full_link <- X %*% betas

linear_link <- full_link - (X[, quad_col_idx] * betas[quad_col_idx])
predicted_data$Linear_Component <- exp(linear_link)

quad_link <- full_link - (X[, linear_col_idx] * betas[linear_col_idx])
predicted_data$Quadratic_Component <- exp(quad_link)

# with CIs
# 1. Extract the model's variance-covariance matrix
vcov_mat <- as.matrix(vcov(model_sti_quad))
betas    <- fixef(model_sti_quad)

# 2. Identify the column indices for the polynomial terms
linear_col_idx <- grep("poly\\(STi.*\\)1", colnames(X))
quad_col_idx   <- grep("poly\\(STi.*\\)2", colnames(X))

# 3. ISOLATE LINEAR COMPONENT: Create a design matrix where the quadratic term is zeroed out
X_linear <- X
X_linear[, quad_col_idx] <- 0

linear_link <- X_linear %*% betas
se_linear_link <- sqrt(diag(X_linear %*% vcov_mat %*% t(X_linear)))

predicted_data$Linear_Component <- exp(linear_link)
predicted_data$Linear_LL        <- exp(linear_link - 1.96 * se_linear_link)
predicted_data$Linear_UL        <- exp(linear_link + 1.96 * se_linear_link)

# 4. ISOLATE QUADRATIC COMPONENT: Create a design matrix where the linear term is zeroed out
X_quad <- X
X_quad[, linear_col_idx] <- 0

quad_link <- X_quad %*% betas
se_quad_link <- sqrt(diag(X_quad %*% vcov_mat %*% t(X_quad)))

predicted_data$Quadratic_Component <- exp(quad_link)
predicted_data$Quadratic_LL        <- exp(quad_link - 1.96 * se_quad_link)
predicted_data$Quadratic_UL        <- exp(quad_link + 1.96 * se_quad_link)
```



```{r quad model figure plot}
# with both linear and quadratic compenents
png(filename = "../Figures/QuadraticParasite_2026-09-11.png", width = 5, height = 5, units = "in", bg = "white", res = 320)

ggplot(data_analysis_parasites, aes(x = STi)) +
  geom_point(aes(y = Tparasites), alpha = 0.25, color = "darkgray", position = position_jitter(height = 0.1, width = 0)) + 
  geom_ribbon(data = predicted_data, aes(ymin = Linear_LL, ymax = Linear_UL, fill = "Linear Component"), alpha = 0.12, show.legend = FALSE) +
  geom_ribbon(data = predicted_data, aes(ymin = Quadratic_LL, ymax = Quadratic_UL, fill = "Quadratic Component"), alpha = 0.12, show.legend = FALSE) +
  geom_line(data = predicted_data, aes(y = Linear_Component, color = "Linear Component"), size = 1.2, linetype = "dashed") +
  geom_line(data = predicted_data, aes(y = Quadratic_Component, color = "Quadratic Component"), size = 1.2, linetype = "solid") +
  scale_color_manual(name = "Model Components", values = c("Linear Component" = "cornflowerblue", "Quadratic Component" = "goldenrod")) +
  scale_fill_manual(values = c("Linear Component" = "cornflowerblue", "Quadratic Component" = "goldenrod")) +
  theme_classic() +
  theme(legend.position = "bottom", 
        panel.grid.minor = element_blank()) +
  labs(x = "MHC Supertype Richness",
    y = "Gyrodactylus Abundance")

dev.off()

```


### Coinertia Supplemental Figures

```{r coia fig data include = FALSE}
load("../output/coia_supp_figs.RData")
```


```{r Supplemental coia plots}
# coin 1
g1 <- ade4::s.arrow(coin_9$l1, label = c("Prevalence", "Orange", "Black"), boxes = TRUE)
g2 <- ade4::s.arrow(coin_9$c1)
g3 <- s.corcircle(coin_9$aX)
g4 <- s.corcircle(coin_9$aY)
g5 <- plot(coin_9)

svg(file="../Figures/coin1_a.svg",width=4, height=4)
ade4::s.arrow(coin_9$l1, label = c("Prevalence", "Orange", "Black"), boxes = TRUE)
dev.off()
svg(file="../Figures/coin1_b.svg",width=4, height=4)
 ade4::s.arrow(coin_9$c1)
dev.off()


# coin 2
g6 <- ade4::s.arrow(coin_10$l1, label = c("Prevalence", "Orange", "Black"), boxes = TRUE)
g7 <- ade4::s.arrow(coin_10$c1)
g10 <- plot(coin_2)

svg(file="../Figures/coin2_a.svg",
width=4, height=4)
ade4::s.arrow(coin_10$l1, label = c("Prevalence", "Orange", "Black"), boxes = TRUE)
dev.off()
svg(file="../Figures/coin2_b.svg",
width=4, height=4)
 ade4::s.arrow(coin_10$c1)
dev.off()
```


## Parasite figure

maybe a figure representing the difference between intensity and infection incidence?