---
title: "03-figures"
author: "Allegra Love"
date: "2025-09-27"
output: html_document
---


```{r setup include = FALSE}
library(car)
library(ggplot2)
library(gridExtra)
library(grid)
library(tidyverse)
```


### Figures

Note: drainage specific hex values for figures
Fill (map) (too light for figure): 
Northern - #b9c1ef
Oropouche - #f0e8bb
Caroni - #eecdcd

Fill (figures):
Northern - #8292ed
Oropouche - #ebdb81
Caroni - #ed8a8a

Lines and points (darker):
Northern - #425ced
Oropouche - #edd442
Caroni - #ed3e3e

Figure 1 was generated in QGIS

### Figure 2: Drainage specific relationship between individual parasite load and relative area black and orange.
```{r fig 2 data include = FALSE}
load("../output/fig2.RData")
#function to sahre a legend between a multi-plot arrangement (from: https://cran.r-project.org/web/packages/egg/vignettes/Ecosystem.html)
grid_arrange_shared_legend <-
  function(...,
           ncol = length(list(...)),
           nrow = 1,
           position = c("bottom", "right")) {
    
    plots <- list(...)
    position <- match.arg(position)
    g <-
      ggplotGrob(plots[[1]] + theme(legend.position = position))$grobs
    legend <- g[[which(sapply(g, function(x)
      x$name) == "guide-box")]]
    lheight <- sum(legend$height)
    lwidth <- sum(legend$width)
    gl <- lapply(plots, function(x)
      x + theme(legend.position = "none"))
    gl <- c(gl, ncol = ncol, nrow = nrow)
    
    combined <- switch(
      position,
      "bottom" = arrangeGrob(
        do.call(arrangeGrob, gl),
        legend,
        ncol = 1,
        heights = unit.c(unit(1, "npc") - lheight, lheight)
      ),
      "right" = arrangeGrob(
        do.call(arrangeGrob, gl),
        legend,
        ncol = 2,
        widths = unit.c(unit(1, "npc") - lwidth, lwidth)
      )
    )
    
    grid.newpage()
    grid.draw(combined)
    
    # return gtable invisibly
    invisible(combined)
    
  }
```


```{r Fig 2}
#ggplot black
bplot <- data.frame(ggeffects::ggpredict(model_b, terms = c("Tparasites [0:12]", "Drainage"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group), show.legend = F) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_jitter(inherit.aes = FALSE,
             data = data_analysis_parasites,
             aes(x = Tparasites, y = Black, colour = Drainage),
             width = 0.1, height = 0.0,
             alpha = 0.5) +
  scale_color_manual(values = c("#ed3e3e","#425ced","#edd442")) +
  scale_fill_manual(values = c("#ed8a8a","#8292ed","#ebdb81")) +
  scale_x_continuous(breaks = seq(0, 12, by = 1)) +
  xlab("Individual Parasite Load") +
  ylab("Relative Area Black") +
  labs(colour = "Drainage", fill = "Drainage") +
  theme_classic() +
  theme(legend.position = "none")

#ggplot orange
oplot <- data.frame(ggeffects::ggpredict(model_o, terms = c("Tparasites [0:12]", "Drainage"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_jitter(inherit.aes = FALSE,
             data = data_analysis_parasites,
             aes(x = Tparasites, y = Black, colour = Drainage),
             width = 0.1, height = 0.0,
             alpha = 0.5) +
  scale_color_manual(values = c("#ed3e3e","#425ced","#edd442")) +
  scale_fill_manual(values = c("#ed8a8a","#8292ed","#ebdb81")) +
  scale_x_continuous(breaks = seq(0, 12, by = 1)) +
  xlab("Individual Parasite Load") +
  ylab("Relative Area Orange") +
  labs(colour = "Drainage", fill = "Drainage") +
  theme_classic() +
  theme(legend.position = "bottom")


jpeg(filename =  "../Figures/ColourParasite 2025-04-06.jpeg", width = 8, height = 4, units = "in", quality = 75, bg = "white", res = 300)

grid_arrange_shared_legend(bplot, oplot)
dev.off()

```


### Figure 3 Two panels, MicroAR on Y axis, one plot with MHC Ar, and one with MHC STr. different symbols for predation.
```{r fig 3 data include = FALSE}
load("../output/fig3.Rdata")
```


```{r Fig 3}
#triangles for low predation (17), coloured by drainage similar to map
drainage_colours <- c(  "Caroni" = "#ed3e3e", 
  "North Slope" = "#425ced",
  "Oropuche" = "#edd442",
  "Oropuche (Caroni Type)" = "#edd442")

# jpeg(filename =  "../Figures/MS Alleles MHC 2025-04-06.jpeg", width = 10, height = 6, units = "in", quality = 75, bg = "white", res = 300)
op <- par(mfrow = c(1,2)
          , mar = c(5.1,5.1,4.1,2.1))
plot(x = PopDF_parasites$`Micro-Ar`, y = PopDF_parasites$mhc_Ai
     , xlab = "Number of Unique Microsattelite Alleles", ylab = "MHC Allelic Richness"
     , pch = ifelse(PopDF_parasites$Predation == "High", 19, 17)
     , cex = 2
     , col = drainage_colours
     , xlim = c(6.5,9.0)
     )
regLine(lm(PopDF_parasites$mhc_Ai ~ PopDF_parasites$`Micro-Ar`), col = "black")

legend(x = 6.5, y = 8
       , c("High Predation", "Low Predation")
       , pch = c(19,17)
       , bty = "n"
       , pt.bg = c("black"), pt.cex = 2, cex = 1.05)
legend(x = 6.5, y = 7.5
       , legend = names(drainage_colours)[1:3]
       , col = drainage_colours[1:3]
       , pch = 18
       , bty = "n"
       , pt.cex = 2
       , cex = 1.05)
plot(x = PopDF_parasites$`Micro-Ar`, y = PopDF_parasites$mhc_STi
     , xlab = "Number of Unique Microsattelite Alleles"
     , ylab = "Number of Unique MHC \nSupertypes Represented by Alleles"
     , pch = ifelse(PopDF_parasites$Predation == "High", 19, 17)
     , cex = 2
     , col = drainage_colours
     , xlim = c(6.5,9.0)
     )
regLine(lm(PopDF_parasites$mhc_STi ~ PopDF_parasites$`Micro-Ar`), col = "black")
par(op)
# dev.off()
```

### Microsat figures

```{r microsat fig data include = FALSE}
load("../output/microsat_figs.RData")
```

Effect size visualiztion (lines 339-345) in ind. models manuscript.
```{r effect size visualization}
#allelic richness
# R2 plot
plot(model_micro_all_ar_r2)

#supertype richness
#R2 plot
plot(model_micro_all_str_r2)
#
```


Making other plots for reference. Manuscript has a summary table with both results (Merged_tbl2).
```{r microsat mixed model plots not in MS, eval=FALSE}

#model plot
data.frame(ggeffects::ggpredict(model_micro_all_ar, terms = c("Micro-Ar", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_point(inherit.aes = FALSE,
             data = PopDF,
             aes(y = `MHC-Ar`, x = `Micro-Ar`, colour = Predation),
             alpha = 0.5) +
  xlab("Microsatellite Richness") +
  ylab("MHC Allelic Richness")



data.frame(ggeffects::ggpredict(model_micro_all_str, terms = c("Micro-Ar", "Predation"))) %>% 
  ggplot(aes(x = x, y = predicted, group = group)) +
  geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error, group = group, fill = group), alpha = 0.35) +
  geom_line(aes(x = x, y = predicted, group = group, colour = group)) +
  geom_point(inherit.aes = FALSE,
             data = PopDF,
             aes(y = `MHC-STr`, x = `Micro-Ar`, colour = Predation),
             alpha = 0.5) +
  xlab("Microsatellite Richness") +
  ylab("MHC Supertype Richness")

```

### Supplemental Figure 3

```{r Supp fig 3}
#Supplemental Figure 3
# jpeg(filename =  "../Figures/Non Effect Colour 2025-04-06.jpeg", width = 8, height = 8, units = "in", quality = 75, bg = "white", res = 300)
op <- par(mfrow = c(2,2)
          , cex.lab = 1.5)
#orange colours
plot(x = PopDF_parasites$mhc_STi, y = PopDF_parasites$mean_o
     , xlab = "Number of unique supertypes", ylab = "Relative Orange Area"
     , pch = ifelse(PopDF_parasites$Predation == "High", 19, 17), cex = 2
     , col = drainage_colours
     )
legend(x = 4.1, y = 0.17
       , c("High Predation", "Low Predation")
       , pch = c(19,17)
       , bty = "n"
       , pt.bg = c("black"), pt.cex = 2, cex = 1.05)
legend(x = 4.1, y = 0.15
       , legend = names(drainage_colours)[1:3]
       , col = drainage_colours[1:3]
       , pch = 18
       , bty = "n"
       , pt.cex = 2
       , cex = 1.05)
regLine(lm(PopDF_parasites$mean_o ~ PopDF_parasites$mhc_STi), col = "black")
plot(x = PopDF_parasites$mhc_STi, y = PopDF_parasites$mean_b
     , xlab = "Number of unique supertypes", ylab = "Relative Black Area"
     , pch = ifelse(PopDF_parasites$Predation == "High", 19, 17), cex = 2
     , col = drainage_colours
     )
regLine(lm(PopDF_parasites$mean_b ~ PopDF_parasites$mhc_STi), col = "black")
plot(x = PopDF_parasites$mhc_Ai, y = PopDF_parasites$mean_o
     , xlab = "MHC Allelic Richness", ylab = "Relative Orange Area"
     , pch = ifelse(PopDF_parasites$Predation == "High", 19, 17), cex = 2
     , col = drainage_colours
     )
regLine(lm(PopDF_parasites$mean_o ~ PopDF_parasites$mhc_Ai), col = "black")
plot(x = PopDF_parasites$mhc_Ai, y = PopDF_parasites$mean_b
     , xlab = "MHC Allelic Richness", ylab = "Relative Black Area"
     , pch = ifelse(PopDF_parasites$Predation == "High", 19, 17), cex = 2
     , col = drainage_colours
     )
regLine(lm(PopDF_parasites$mean_b ~ PopDF_parasites$mhc_Ai), col = "black")
par(op)
# dev.off()

```

### Coinertia Supplemental Figures

```{r coia fig data include = FALSE}
load("../output/coia_supp_figs.RData")
```


```{r Supplemental coia plots}
# coin 1
g1 <- ade4::s.arrow(coin_9$l1, label = c("Prevalence", "Orange", "Black"), boxes = TRUE)
g2 <- ade4::s.arrow(coin_9$c1)
g3 <- s.corcircle(coin_9$aX)
g4 <- s.corcircle(coin_9$aY)
g5 <- plot(coin_9)

svg(file="../Figures/coin1_a.svg",width=4, height=4)
ade4::s.arrow(coin_9$l1, label = c("Prevalence", "Orange", "Black"), boxes = TRUE)
dev.off()
svg(file="../Figures/coin1_b.svg",width=4, height=4)
 ade4::s.arrow(coin_9$c1)
dev.off()


# coin 2
g6 <- ade4::s.arrow(coin_10$l1, label = c("Prevalence", "Orange", "Black"), boxes = TRUE)
g7 <- ade4::s.arrow(coin_10$c1)
g10 <- plot(coin_2)

svg(file="../Figures/coin2_a.svg",
width=4, height=4)
ade4::s.arrow(coin_10$l1, label = c("Prevalence", "Orange", "Black"), boxes = TRUE)
dev.off()
svg(file="../Figures/coin2_b.svg",
width=4, height=4)
 ade4::s.arrow(coin_10$c1)
dev.off()
```
